---
title: "Prelim_Visualization"
output: html_document
date: "2025-04-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(readxl)
library(purrr)
library(lubridate)
library(lmtest)
library(lme4)
library(car)
library(bbmle)
```

### Read in pop data and look-up table

```{r}
#BH <- read_csv("../input/Cleaned/BH_cleaned_18_23.csv")
CP2 <- read_csv(
  "../input/Cleaned/CP2_cleaned_17_23.csv",
  col_types = cols(Longest_leaf_cm = col_character())
)  %>% #need to do this so <0.5 in Longest_leaf_cm is kept as.character when read in
  mutate(Num_flowers = as.character(Num_flowers)) %>%
  mutate(Num_fruits = as.character(Num_fruits))
WL1 <- read_csv("../input/Cleaned/WL1_cleaned_17_23.csv") %>%
  mutate(Transect_plot = as.character(Transect_plot)) %>%
  mutate(Quad = as.character(Quad)) %>%
  mutate(Height_cm = as.character(Height_cm)) %>%
  mutate(Num_flowers = as.character(Num_flowers)) %>%
  mutate(Num_fruits = as.character(Num_fruits))
WL2 <- read_csv("../input/Cleaned/WL2_cleaned_16_23.csv") %>%
  mutate(Height_cm = as.character(Height_cm)) %>%
  mutate(Longest_leaf_cm = as.character(Longest_leaf_cm)) %>%
  mutate(Num_fruits = as.character(Num_fruits))
WL3 <- read_csv("../input/Cleaned/WL3_cleaned_17_23.csv") %>%
  mutate(Transect_plot = as.character(Transect_plot)) %>%
  mutate(Height_cm = as.character(Height_cm)) %>%
  mutate(Num_flowers = as.character(Num_flowers)) %>%
  mutate(Num_fruits = as.character(Num_fruits))
TM2 <- read_csv("../input/Cleaned/TM2_cleaned_16_23.csv") %>%
  mutate(Longest_leaf_cm = as.character(Longest_leaf_cm))

lookuptable <- read_excel("../input/STTO germination and reproductive phenology lookup table.xlsx")
```

### Merge pops

```{r}
germ_repro_pops <- bind_rows(CP2, WL1, WL2, WL3, TM2)
```

### Prep look-up table to merge

```{r}
#rename cols for germ survey and update pick colors
germ_dates <- lookuptable %>%
  filter(Survey_type == "germination survey") %>%
  rename(Pick_color = toothpick_color) %>%
  rename(Germ_date = Date) %>%
  rename(Census_year = Growing_season) %>%
  mutate(Pick_color = ifelse(
    Pick_color == "large_blue", "blue", Pick_color
  )) %>%
  mutate(Pick_color = ifelse(
    Pick_color == "large_yellow", "yellow", Pick_color
  ))
#filter to keep germ related cols
germ_dates <- germ_dates[,c(1:3, 5, 7, 12)] %>%
  rename(germ_notes = notes)

#rename cols for flw survey and update sword colors
flw_dates <- lookuptable %>%
  filter(Survey_type == "flowering survey") %>%
  rename(Sword_color = sword_color) %>%
  rename(Flw_date = Date) %>%
  rename(Census_year = Growing_season) %>%
  mutate(
    notes = ifelse(
      Sword_color == "yellow for flower, orange if had fruit",
      ifelse(is.na(notes) | notes == "",
             "yellow for flower, orange if had fruit",
             paste(notes, "yellow for flower, orange if had fruit", sep = "; ")),
      notes
    ),
    Sword_color = ifelse(
      Sword_color == "yellow for flower, orange if had fruit",
      "yellow,orange",
      Sword_color
    )
  ) %>%
  separate_rows(Sword_color, sep = ",") %>%
  mutate(Sword_color = trimws(Sword_color)) %>%
  mutate(
    notes = ifelse(
      Sword_color == "green for new flower",
      ifelse(is.na(notes) | notes == "", "green for new flower",
             paste(notes, "green for new flower", sep = "; ")), notes),
    Sword_color = ifelse(
      Sword_color == "green for new flower", "green",Sword_color)) %>%
  mutate(
    notes = ifelse(
      Sword_color == "yellow for many flower",
      ifelse(is.na(notes) | notes == "", "yellow for many flower",
             paste(notes, "yellow for many flower", sep = "; ")), notes),
    Sword_color = ifelse(
      Sword_color == "yellow for many flower", "yellow",Sword_color)) %>%
  mutate(
    notes = ifelse(
      Sword_color == "orange if had fruit",
      ifelse(is.na(notes) | notes == "", "orange if had fruit",
             paste(notes, "orange if had fruit", sep = "; ")), notes),
    Sword_color = ifelse(
      Sword_color == "orange if had fruit", "orange",Sword_color)) %>%
  mutate(
    notes = ifelse(
      Sword_color == "blue for new flower and green if had fruit",
      ifelse(is.na(notes) | notes == "",
             "blue for new flower and green if had fruit",
             paste(notes, "blue for new flower and green if had fruit", sep = "; ")),
      notes
    ),
    Sword_color = ifelse(
      Sword_color == "blue for new flower and green if had fruit",
      "blue,green",
      Sword_color
    )
  ) %>%
  separate_rows(Sword_color, sep = ",") %>%
  mutate(Sword_color = trimws(Sword_color))
#filter to keep flw survey cols
flw_dates <- flw_dates[,c(1:3, 9, 12)] %>%
  rename(flw_notes = notes)
```

### Add missing years

```{r}
germ_dates <- germ_dates %>%
  mutate(Germ_date = as.Date(Germ_date)) %>%
  add_row(Germ_date = as.Date("2021-05-05"), Census_year = 2021, Site = "CP2", cohort="3", Pick_color="orange") %>%
  add_row(Germ_date = as.Date("2021-05-12"), Census_year = 2021, Site = "CP2", cohort="4", Pick_color="yellow") %>%
  add_row(Germ_date = as.Date("2021-05-18"), Census_year = 2021, Site = "CP2", cohort="5", Pick_color="blue") %>%
  add_row(Germ_date = as.Date("2021-05-24"), Census_year = 2021, Site = "CP2", cohort="6", Pick_color="pink") %>%
  add_row(Germ_date = as.Date("2021-06-05"), Census_year = 2021, Site = "CP2", cohort="7", Pick_color="white") %>%
  add_row(Germ_date = as.Date("2018-01-03"), Census_year = 2018, Site = "WL2", Pick_color="yellow") %>%
  add_row(Germ_date = as.Date("2020-05-10"), Census_year = 2020, Site = "CP2", Pick_color="red") %>%
  add_row(Germ_date = as.Date("2021-04-19"), Census_year = 2021, Site = "CP2", Pick_color="green") %>%
  add_row(Germ_date = as.Date("2019-01-23"), Census_year = 2019, Site = "TM2", Pick_color="red")
```

### Add new col for germ/flw timing - days since Sept 1

```{r}
#convert to days since Sept 1
germ_dates <- germ_dates %>%
  mutate(
    Census_year = as.numeric(Census_year),  # force to numeric
    Germ_date = as.Date(Germ_date),
    Germ_days_since_Sept1 = as.numeric(
      Germ_date - as.Date(paste0(Census_year - 1, "-09-01"))
    )
  )

flw_dates <- flw_dates %>%
  mutate(
    Census_year = as.numeric(Census_year),  # force to numeric
    Flw_date = as.Date(Flw_date),
    Flw_days_since_Sept1 = as.numeric(
      Flw_date - as.Date(paste0(Census_year - 1, "-09-01"))
    )
  )
```

### Update pick and sword color names

```{r}
germ_repro_pops <- germ_repro_pops %>%
  mutate(Pick_color = case_when(
    Pick_color == "blu" ~ "blue",
    TRUE ~ as.character(Pick_color)
  )) %>%
  mutate(Pick_color = case_when(
    Pick_color == "grn" ~ "green",
    TRUE ~ as.character(Pick_color)
  )) %>%
  mutate(Pick_color = case_when(
    Pick_color == "yelllow" ~ "yellow",
    TRUE ~ as.character(Pick_color)
  )) %>%
  mutate(Pick_color = case_when(
    Pick_color == "ylw" ~ "yellow",
    TRUE ~ as.character(Pick_color)
  )) %>%
  mutate(Pick_color = case_when(
    Pick_color == "Ylw" ~ "yellow",
    TRUE ~ as.character(Pick_color)
  )) %>%
  mutate(Pick_color = case_when(
    Pick_color == "ylw*" ~ "orange",
    TRUE ~ as.character(Pick_color)
  )) %>%
  mutate(Pick_color = case_when(
    Pick_color == "R" ~ "red",
    TRUE ~ as.character(Pick_color)
  )) %>%
  mutate(Pick_color = case_when(
    Pick_color == "O" ~ "orange",
    TRUE ~ as.character(Pick_color)
  )) %>%
  mutate(Pick_color = case_when(
    Pick_color == "o" ~ "orange",
    TRUE ~ as.character(Pick_color)
  )) %>%
  mutate(Pick_color = case_when(
    Pick_color == "r" ~ "red",
    TRUE ~ as.character(Pick_color)
  )) %>%
  mutate(Pick_color = case_when(
    Pick_color == "W" ~ "white",
    TRUE ~ as.character(Pick_color)
  )) %>%
  mutate(Pick_color = case_when(
    Pick_color == "*" ~ "unmarked",
    TRUE ~ as.character(Pick_color)
  )) %>%
  mutate(Pick_color = case_when(
    Pick_color == "none" ~ NA_character_,
    TRUE ~ as.character(Pick_color)
  )) %>%
  mutate(Pick_color = case_when(
    Pick_color == "unmarked" ~ NA_character_,
    TRUE ~ as.character(Pick_color)
  )) %>%
  mutate(Pick_color = case_when(
    Pick_color == "blk" ~ "black",
    TRUE ~ as.character(Pick_color)
  )) %>%
  mutate(Pick_color = case_when(
    Pick_color == "G" ~ "green",
    TRUE ~ as.character(Pick_color)
  )) %>%
  mutate(Sword_color = case_when(
    Sword_color == "ylw" ~ "yellow",
    TRUE ~ as.character(Sword_color)
  )) %>%
  mutate(Sword_color = case_when(
    Sword_color == "grn" ~ "green",
    TRUE ~ as.character(Sword_color)
  )) %>%
  mutate(Sword_color = case_when(
    Sword_color == "whie" ~ "white",
    TRUE ~ as.character(Sword_color)
  )) %>%
  mutate(Sword_color = case_when(
    Sword_color == "none" ~ NA_character_,
    TRUE ~ as.character(Sword_color)
  )) %>%
  mutate(Sword_color = case_when(
    Sword_color == "*" ~ "unmarked",
    TRUE ~ as.character(Sword_color)
  )) %>%
  mutate(Sword_color = case_when(
    Sword_color == "unmarked" ~ NA_character_,
    TRUE ~ as.character(Sword_color)
  )) %>%
  mutate(Sword_color = case_when(
    Sword_color == "no" ~ NA_character_,
    TRUE ~ as.character(Sword_color)
  )) %>%
  mutate(Sword_color = case_when(
    Sword_color == "none (green? One loose sword in the plot)" ~ NA_character_,
    TRUE ~ as.character(Sword_color)
  )) %>%
  mutate(Sword_color = case_when(
    Sword_color == "fallen out" ~ NA_character_,
    TRUE ~ as.character(Sword_color)
  )) %>%
  mutate(Sword_color = case_when(
    Sword_color == "N" ~ NA_character_,
    TRUE ~ as.character(Sword_color)
  )) %>%
  mutate(Sword_color = case_when(
    Sword_color == "blu" ~ "blue",
    TRUE ~ as.character(Sword_color)
  )) %>%
  mutate(Pick_color = case_when(
    Pick_color == "neon blu" ~ "blue",
    TRUE ~ as.character(Pick_color)
  )) %>%
  mutate(Pick_color = case_when(
    Pick_color == "neon grn" ~ "green",
    TRUE ~ as.character(Pick_color)
  )) %>%
  mutate(Pick_color = case_when(
    Pick_color == "NO" ~ NA_character_,
    TRUE ~ as.character(Pick_color)
  ))
```

### Join germ and flw dates with annual census data

```{r}
germ_repro_pops_dates <- germ_repro_pops %>%
  left_join(germ_dates, by=c("Site", "Census_year", "Pick_color")) %>%
  left_join(flw_dates, by=c("Site", "Census_year", "Sword_color"))
```

#### Check every row with a pick/sword color now has a corresponding date - still missing dates for some individuals from CP2, WL2, and WL3

```{r}
#number of pick/sword rows
sum(!is.na(germ_repro_pops_dates$Pick_color))
sum(!is.na(germ_repro_pops_dates$Sword_color))

#number of germ/flw date rows
sum(!is.na(germ_repro_pops_dates$Germ_days_since_Sept1))
sum(!is.na(germ_repro_pops_dates$Flw_days_since_Sept1))

#~300 rows missing germ date
#~300 rows missing flw date

bad_germ_merge <- germ_repro_pops_dates %>%
  filter(!is.na(Pick_color) & is.na(Germ_days_since_Sept1))
```

### Plot germ time vs flower time 2 ways

```{r}
ggplot(germ_repro_pops_dates, aes(x=Germ_days_since_Sept1, y=Flw_days_since_Sept1))+
  geom_point() +
  facet_wrap(~Site, scales="free")

ggplot(germ_repro_pops_dates, aes(x=Germ_days_since_Sept1, y=Flw_days_since_Sept1, color=Census_year))+
  geom_point() +
  facet_wrap(~Site, scales="free")

#TM2 has the best data - will run model only using TM2
```

### Probability of surviving to reproduction and probability of reproducing

```{r}
#need to clean entries in Phenology column

prob_surv_repro <- germ_repro_pops_dates %>%
  mutate(Phenology = if_else(Phenology == "PF", "P", Phenology)) %>%
  mutate(Phenology = if_else(Phenology == "b", "B", Phenology)) %>%
  mutate(Phenology = if_else(Phenology == "f", "F", Phenology)) %>%
  mutate(Phenology = if_else(Phenology == "s", "S", Phenology)) %>%
  mutate(Phenology = if_else(Phenology == "Na", NA_character_, Phenology)) %>%
  mutate(Phenology = if_else(Phenology == "LS", "V", Phenology)) %>%
  mutate(Phenology = if_else(Phenology == "M", NA_character_, Phenology)) %>%
  mutate(Phenology = if_else(Phenology == "x", "X", Phenology)) %>%
  mutate(Phenology = if_else(Phenology == "D", "X", Phenology)) %>%
  mutate(Phenology = if_else(Phenology == "blank (probably X)", "X", Phenology)) %>% #there are still other entries to address but this should be enough to determine if plants were dead or alive at the census
  mutate(prob_surv_to_repro = if_else(Phenology == "X",0, 1)) %>% #0 if pheno is X and 1 if pheno is anything else - I'm not sure I like this because there are cases where a plant reproduced but died before the census
  mutate(probRepro = if_else(is.na(Flw_date), 0, 1)) #0 if NA for flw data, 1 if there is a date 
```

### Models

#### Visualize germ and flw time distributions

```{r}
#visualize distribution
ggplot(prob_surv_repro, aes(x=Germ_days_since_Sept1)) +
  geom_histogram() #not the most normal distribution, right tail

ggplot(prob_surv_repro, aes(x=Flw_days_since_Sept1)) +
  geom_histogram() #slightly more normal but almost bimodal

#interesting there are inverse trends when only plotting TM2
```

#### Log transform germ and flw time

```{r}
#log transform
scaled_prob_surv_repro <- prob_surv_repro %>%
  mutate(log_flw_time = log(Flw_days_since_Sept1)) %>%
  mutate(log_germ_time = log(Germ_days_since_Sept1))
  
ggplot(scaled_prob_surv_repro, aes(x=log_germ_time)) +
  geom_histogram() #might be worse

ggplot(scaled_prob_surv_repro, aes(x=log_flw_time)) +
  geom_histogram() #slighlty better
```

### Filter to only include TM2 for Flw time \~ Germ time analyses since it is the only pop with sufficient sample size

```{r}
#filter for only TM2
TM2_prob_surv_repro <- scaled_prob_surv_repro %>%
  filter(Site == "TM2") %>%
  filter(!is.na(Flw_days_since_Sept1) & !is.na(Germ_days_since_Sept1))

ggplot(TM2_prob_surv_repro, aes(x=Germ_days_since_Sept1, y=Flw_days_since_Sept1))+
  geom_point() +
  facet_wrap(~Census_year)

ggplot(TM2_prob_surv_repro, aes(x=log_germ_time, y=log_flw_time)) +
  geom_point() +
  facet_wrap(~Census_year)
```

#### Flw time \~ Germ time Fixed effects models

```{r}
#only fixed effect of germ days - SIGNIFICANT slightly negative relationship
null <- glm(Flw_days_since_Sept1 ~ 1,
            data=TM2_prob_surv_repro,
            family=gaussian)
m1_Germ <- glm(Flw_days_since_Sept1 ~ Germ_days_since_Sept1,
            data=TM2_prob_surv_repro,
            family=gaussian)
summary(m1_Germ)
lrtest(null, m1_Germ)

plot(m1_Germ, which = 1)
qqnorm(resid(m1_Germ))
qqline(resid(m1_Germ))

m1.25_log_Germ <- glm(log_flw_time ~ log_germ_time,
            data=TM2_prob_surv_repro,
            family=gaussian)

plot(m1.25_log_Germ, which = 1)
qqnorm(resid(m1.25_log_Germ))
qqline(resid(m1.25_log_Germ)) #log transformed looks slightly better


m1.5_Germ_year_int <- m1_Germ <- glm(log_flw_time ~ log_germ_time*Census_year,
            data=TM2_prob_surv_repro,
            family=gaussian)
summary(m1.5_Germ_year_int)

plot(m1.5_Germ_year_int, which = 1)
qqnorm(resid(m1.5_Germ_year_int))
qqline(resid(m1.5_Germ_year_int))
```

#### Flw time \~ Germ time Random effects models

```{r}
#model with random effect of year
m2_Germ_year <- lmer(Flw_days_since_Sept1 ~ Germ_days_since_Sept1 + (1 |Census_year), data = TM2_prob_surv_repro)
summary(m2_Germ_year) #year explains some of the variance
Anova(m2_Germ_year) #Germ day is still significant but now slightly positive

#testing random effects of year, transect, and quad
m3_Germ_year_transect_plot <- lmer(Flw_days_since_Sept1 ~ Germ_days_since_Sept1 + (1|Census_year) + (1|Transect_plot) +(1|Quad), data = TM2_prob_surv_repro)
summary(m3_Germ_year_transect_plot) #year explains some of the variance

#compare random to fixed effect model
ICtab(m1.25_log_Germ, m1.5_Germ_year_int, m2_Germ_year, m3_Germ_year_transect_plot)
```

#### Predicted vs Observed Flw time \~ Germ time

```{r}
TM2_prob_surv_repro$germ_flw_model <- predict(m1.5_Germ_year_int, newdata = TM2_prob_surv_repro, type = "response")

ggplot(TM2_prob_surv_repro, aes(x = log_germ_time, y = log_flw_time)) +
  geom_point(alpha = 0.3) +
  geom_smooth(aes(y = germ_flw_model), method = "glm", se = FALSE, linetype = "solid") +
  labs(
    x = "Log Germination Time",
    y = "Log Flowering Time",) +
  theme_minimal()
```

#### Visualize germ time and prob of surviving to reproduction and prob of reproducing

```{r}
germ_prop_surv_repro <- scaled_prob_surv_repro %>%
  filter(!is.na(log_germ_time))

ggplot(germ_prop_surv_repro, aes(x=Germ_days_since_Sept1, y=prob_surv_to_repro)) +
  geom_point()

ggplot(germ_prop_surv_repro, aes(x=log_germ_time, y=prob_surv_to_repro)) +
  geom_point()

ggplot(germ_prop_surv_repro, aes(x=log_germ_time, y=probRepro)) +
  geom_point()
```

#### Survival to Repro \~ Germ time Fixed effects models

```{r}
null_str <- glm(prob_surv_to_repro ~ 1,
                data=germ_prop_surv_repro,
                family=binomial())
m4_Germ_str <- glm(prob_surv_to_repro ~ Germ_days_since_Sept1,
                      data=germ_prop_surv_repro,
                family=binomial())
summary(m4_Germ_str)


m4.5_logGerm_str <- glm(prob_surv_to_repro ~ log_germ_time,
                      data=germ_prop_surv_repro,
                family=binomial())
summary(m4.5_logGerm_str)

m5_logGerm_year_str <- glm(prob_surv_to_repro ~ log_germ_time*Census_year,
                      data=germ_prop_surv_repro,
                family=binomial())
summary(m5_logGerm_year_str)
lrtest(null_str, m4_Germ_str, m4.5_logGerm_str, m5_logGerm_year_str)

m6_site_str <- glm(prob_surv_to_repro ~ Site,
                      data=germ_prop_surv_repro,
                family=binomial())
summary(m6_site_str)
lrtest(null_str, m6_site_str)

m7_logGerm_site_str <- glm(prob_surv_to_repro ~ log_germ_time*Site,
                      data=germ_prop_surv_repro,
                family=binomial())
summary(m7_logGerm_site_str)

m8_logGerm_year_site_str <- glm(prob_surv_to_repro ~ log_germ_time*Census_year*Site,
                      data=germ_prop_surv_repro,
                family=binomial())
summary(m8_logGerm_year_site_str)

ICtab(m4.5_logGerm_str, m5_logGerm_year_str, m6_site_str, m7_logGerm_site_str, m8_logGerm_year_site_str) #three way interaction with log germ time, census year, and site was the best fitting model
```

#### Survival to Repro \~ Germ time Random effects models

```{r}
m9_logGerm_year_site_plot_quad <- glmer(prob_surv_to_repro ~ log_germ_time + (1|Census_year) + (1|Site) + (1|Transect_plot) +(1|Quad),
             data=germ_prop_surv_repro,
             family=binomial)
summary(m9_logGerm_year_site_plot_quad)

#m10_logGerm_Site_Year_plot_quad <- glmer(prob_surv_to_repro ~ log_germ_time*Census_year*Site + (1|Transect_plot) +(1|Quad),
           #  data=germ_prop_surv_repro,
           #  family=binomial) #not enough power for this
#summary(m10_logGerm_Site_Year_plot_quad)

ICtab(m8_logGerm_year_site_str, m9_logGerm_year_site_plot_quad) #the random effects model is a better fit than the 3 way interaction
```

#### Probability of Surviving to Reproduction

```{r}
m9.5_prob_surv_repro <- glmer(prob_surv_to_repro ~ 1 + (1|Census_year) + (1|Site) + (1|Transect_plot) +(1|Quad),
             data=germ_prop_surv_repro,
             family=binomial)
summary(m9.5_prob_surv_repro)
```

#### Predicted vs Observed Survival to Repro \~ Germ time

plot everything on a logit scale and see how it looks

or could calculate means and variances on a logit scale (grouping by germination survey date), back-transform

plot log(germ_time ) on the x

```{r}
germ_prop_surv_repro$prob_surv_model <- predict(m9_logGerm_year_site_plot_quad, newdata = germ_prop_surv_repro, type = "response", allow.new.levels = TRUE)

ggplot(germ_prop_surv_repro, aes(x = Germ_days_since_Sept1, y = prob_surv_to_repro)) +
  geom_point(alpha = 0.3) +
  geom_smooth(aes(y = prob_surv_model), method = "glm", se = FALSE, linetype = "solid") +
  facet_wrap(~Site) +
  labs(
    x = "Germination Time",
    y = "Probability of surviving to reproduction",) +
  theme_minimal()
```

#### Prob of Repro \~ Germ time Fixed effects models - only TM2 will have probRepro=1

```{r}
null_probRepro <- glm(probRepro ~ 1,
                data=germ_prop_surv_repro,
                family=binomial())
m11_Germ_probRepro <- glm(probRepro ~ Germ_days_since_Sept1,
                      data=germ_prop_surv_repro,
                family=binomial())
summary(m11_Germ_probRepro)

m11.5_logGerm_probRepro <- glm(probRepro ~ log_germ_time,
                      data=germ_prop_surv_repro,
                family=binomial())
summary(m11.5_logGerm_probRepro)

m12_logGerm_year_probRepro <- glm(probRepro ~ log_germ_time*Census_year,
                      data=germ_prop_surv_repro,
                family=binomial())
summary(m12_logGerm_year_probRepro)
lrtest(null_probRepro, m11_Germ_probRepro, m11.5_logGerm_probRepro, m12_logGerm_year_probRepro)

ICtab(m11.5_logGerm_probRepro, m12_logGerm_year_probRepro) #log germ time and year interaction is the best fit

#only TM2 has probRepro=1 so can't do this for now
#m13_site_probRepro <- glm(probRepro ~ Site,
        #              data=germ_prop_surv_repro,
         #       family=binomial())
#summary(m13_site_probRepro) 
#lrtest(null_probRepro, m13_site_probRepro)

#m14_logGerm_site_probRepro <- glm(probRepro ~ log_germ_time*Site,
  #                    data=germ_prop_surv_repro,
   #             family=binomial())
#summary(m14_logGerm_site_probRepro)

#m15_logGerm_year_site_probRepro <- glm(probRepro ~ log_germ_time*Census_year*Site,
   #                   data=germ_prop_surv_repro,
   #             family=binomial())
#summary(m15_logGerm_year_site_probRepro)
```

#### Prob of Repro \~ Germ time Random effects models - only TM2 will have probRepro=1

```{r}
m16_logGerm_year_plot_quad <- glmer(probRepro ~ log_germ_time + (1|Census_year) +  (1|Transect_plot) +(1|Quad),
             data=germ_prop_surv_repro,
             family=binomial)
summary(m16_logGerm_year_plot_quad)

ICtab(m12_logGerm_year_probRepro, m16_logGerm_year_plot_quad) #random effect model is the better fit
```

#### Predicted vs Observed Prob of Repro \~ Germ time

plot x=axis on logGermtime

```{r}
germ_prop_surv_repro$probRepro_model <- predict(m16_logGerm_year_plot_quad, newdata = germ_prop_surv_repro, type = "response", allow.new.levels = TRUE)

ggplot(germ_prop_surv_repro, aes(x = Germ_days_since_Sept1, y = probRepro)) +
  geom_point(alpha = 0.3) +
  geom_smooth(aes(y = probRepro_model), method = "glm", se = FALSE, linetype = "solid") +
  labs(
    x = "Germination Time",
    y = "Probability of reproducing",) +
  theme_minimal()

ggplot(germ_prop_surv_repro, aes(x = Germ_days_since_Sept1, y = probRepro)) +
  geom_point(alpha = 0.3) +
  geom_smooth(aes(y = probRepro_model), method = "glm", se = FALSE, linetype = "solid") +
  facet_wrap(~Site)+
  labs(
    x = "Germination Time",
    y = "Probability of reproducing",) +
  theme_minimal()
```
