---
title: "Germ_Cleaning"
output: html_document
date: "2026-02-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load in libraries

```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(readxl)
library(lubridate)
```

### Load in TM2 Germ and Mortality Data

```{r}
#TM2 2016-2017
TM2_germ_2016 <- read_excel("../input/Germination and Mortality data/TM2_Germ_Mort/Demog_TM2_germsurvey_fall2016winter2017_20170717_AJC.xlsx") #will need to load in each sheet seperately

#TM2 2017-2018
TM2_germ_20171205 <- read_excel("../input/Germination and Mortality data/TM2_Germ_Mort/Demog_TM2_germsurvey_fall2017winter2018_20171205_AJC.xlsx")
TM2_germ_20171214 <- read_excel("../input/Germination and Mortality data/TM2_Germ_Mort/Demog_TM2_germsurvey_fall2017winter2018_20171214_AJC.xlsx")
TM2_germ_20180117 <- read_excel("../input/Germination and Mortality data/TM2_Germ_Mort/Demog_TM2_germsurvey_winter2018spring2018_20180727_SLZ.xlsx")
TM2_germ_2018 <- read_excel("../input/Germination and Mortality data/TM2_Germ_Mort/TM2_germ_2018.csv")

#TM2 2018-2019
TM2_germ_2019 <- read_excel("../input/Germination and Mortality data/TM2_Germ_Mort/TM2_germ_2019.csv")
TM2_germ_20181214 <- read_excel("../input/Germination and Mortality data/TM2_Germ_Mort/TM2_GermSurvey_12.14.18_entered2.25.21ES.xlsx")
TM2_germ_20181231 <- read_excel("../input/Germination and Mortality data/TM2_Germ_Mort/TM2_GermSurvey_12.31.18_edited9.4.19MM.xlsx")
TM2_germ_20190123 <- read_excel("../input/Germination and Mortality data/TM2_Germ_Mort/TM2_GermSurvey_1.23.19_edited9.4.19MM.xlsx")
TM2_germ_20190313 <- read_excel("../input/Germination and Mortality data/TM2_Germ_Mort/TM2_GermSurvey_3.13.2019_edited9.3.19MM.xlsx")
TM2_germ_20190419 <- read_excel("../input/Germination and Mortality data/TM2_Germ_Mort/TM2_mortality_updated_20190419ES.xlsx")
TM2_germ_20190508 <- read_excel("../input/Germination and Mortality data/TM2_Germ_Mort/TM2_mortality_updated_20190508ES.xlsx")

#TM2 2019-2020
TM2_germ_2020 <- read_excel("../input/Germination and Mortality data/TM2_Germ_Mort/TM2_germ_2020_done.xlsx")
TM2_germ_20191211 <- read_excel("../input/Germination and Mortality data/TM2_Germ_Mort/Germination_survey_field_data_20200423_EEB.xlsx")

#TM2 2020-2021
TM2_germ_2021 <- read_excel("../input/Germination and Mortality data/TM2_Germ_Mort/TM2_germinationsurvey_2021_all_C.xlsx")
TM2_mort_20210226 <- read_excel("../input/Germination and Mortality data/TM2_Germ_Mort/TM2_Mortality_Survey_20210226_verified.xlsx")


#TM2 2021-2022
TM2_germ_2022 <- read_excel("../input/Germination and Mortality data/TM2_Germ_Mort/TM2_germinationsurvey_2022_all_finished.xlsx")
TM2_germ_20211210 <- read_excel ("../input/Germination and Mortality data/TM2_Germ_Mort/TM2_mortalitysurvey_20211210_finished.xlsx")
TM2_germ_20220105 <- read_excel("../input/Germination and Mortality data/TM2_Germ_Mort/TM2_mortalitysurvey_20220105_finished.xlsx")
TM2_mort_20220114 <- read_excel("../input/Germination and Mortality data/TM2_Germ_Mort/TM2_mortalitysurvey_20220114_verified.xlsx.xlsx")
TM2_mort_20220222 <- read_excel("../input/Germination and Mortality data/TM2_Germ_Mort/TM2_mortalitysurvey_20220222_verified.xlsx.xlsx")


#TM2 2022-2023
TM2_germ_20221014 <- read_excel("../input/Germination and Mortality data/TM2_Germ_Mort/TM2_germinationsurvey_20221014_verified_SRA.xlsx")

```

### TM2 2022-2023 Germination - only 1 survey

```{r}
TM2_germ_20221014 <- TM2_germ_20221014 %>%
  mutate(survey_date = ymd(survey_date)) %>%
  rename(germ_survey_date = survey_date,
         germ_notes = notes)

#write.csv(TM2_germ_20221014, "../output/all_TM2_germ_2022_2023.csv", row.names = FALSE)
```

### TM2 2021-2022 Germination and Mortality

```{r}
#TM2_germ_2022 has the 20211103, 20211117, 20211210, 20220105, and 20220422 surveys

#rename columns 
TM2_germ_2022 <- TM2_germ_2022 %>%
  rename(germ_survey_date = survey_date,
         germ_notes = notes)


TM2_22_germ1 <- TM2_germ_2022 %>%
  filter(pick_color == "red")


TM2_22_germ1_mort1 <- TM2_22_germ1 %>%
  left_join(TM2_mort_20220114[,c(4:9,16)] %>%
              rename(mort_survey_date = survey_date,
                     mort_notes = notes)) %>%
  mutate(sum_check  = red_living + red_dead == number_germ) #many do not match

TM2_22_germ1_mort1_updated <- TM2_22_germ1_mort1 %>%
  select(-sum_check) %>%
  mutate(missing_germs = 0) %>%
  mutate()


TM2_mort_20220114
TM2_mort_20220222
```

Germ Cohort 2

```{r}
TM2_22_germ2 <- TM2_germ_2022 %>%
  filter(pick_color == "yellow")

TM2_22_germ2_mort1 <- TM2_22_germ2 %>%
  left_join(TM2_mort_20220114[,c(4:7, 10:11,16)] %>%
              rename(mort_survey_date = survey_date,
                     mort_notes = notes)) %>%
  mutate(sum_check  = yellow_living + yellow_dead == number_germ)
```

Germ Cohort 3

```{r}
TM2_22_germ3 <- TM2_germ_2022 %>%
  filter(pick_color == "blue")

TM2_22_germ3_mort1 <- TM2_22_germ3 %>%
  left_join(TM2_mort_20220114[,c(4:7, 12:13,16)] %>%
              rename(mort_survey_date = survey_date,
                     mort_notes = notes)) %>%
  mutate(sum_check  = blue_living + blue_dead == number_germ)
```

Germ Cohort 4

```{r}
TM2_22_germ4 <- TM2_germ_2022 %>%
  filter(pick_color == "orange")

TM2_22_germ4_mort1 <- TM2_22_germ4 %>%
  left_join(TM2_mort_20220114[,c(4:7, 14:16)] %>%
              rename(mort_survey_date = survey_date,
                     mort_notes = notes)) %>%
  mutate(sum_check  = orange_living + orange_dead == number_germ)
```

Germ Cohort 5 - no mort survey following this cohort

```{r}
TM2_22_germ5 <- TM2_germ_2022 %>%
  filter(pick_color == "white")

```

### TM2 2020-2021 Germination and Mort

Germ cohort 1

```{r}
TM2_germ_2021 <- TM2_germ_2021 %>%
  rename(germ_survey_date = survey_date,
         germ_notes = notes)
  
TM2_mort_20210226 <- TM2_mort_20210226 %>%
  rename(mort_survey_date = survey_date,
         mort_notes = notes)

TM2_21_germ1 <- TM2_germ_2021 %>%
  filter(pick_color == "red")

TM2_21_germ1_mort <- TM2_21_germ1 %>%
  left_join(TM2_mort_20210226[,c(4:9, 18)]) %>%
  mutate(sum_check  = color_living_red + color_dead_red == number_germ)

TM2_21_germ1_mort_updated <- TM2_21_germ1_mort %>%
  select(-sum_check) %>%
  mutate(missing_germs = 0) %>%
  mutate(number_germ = ifelse(transect_plot == "lower" & quadrat == 7.5, 8, number_germ),
         number_germ = ifelse(transect_plot == "lower" & quadrat == 10.0, 2, number_germ)) %>%
  mutate(sum_check  = color_living_red + color_dead_red + missing_germs== number_germ)

TM2_mort_germ1 <- TM2_mort_20210226 %>%
  rename(num_dead = color_dead_red) %>%
        select(mort_survey_date, transect_plot, quadrat, num_dead, mort_notes)

TM2_21_germ1_mort_long <- TM2_21_germ1_mort_updated %>%
  select(site, survey_year, transect_plot, quadrat, germ_survey_date, pick_color, number_germ, missing_germs, germ_notes) %>%
  left_join(TM2_mort_germ1)

TM2_21_germ1_day_since_germ <- TM2_21_germ1_mort_long %>%
   mutate(germ_survey_date = as.Date(germ_survey_date),
         mort_survey_date = as.Date(mort_survey_date),
         days_since_germ = as.integer(mort_survey_date - germ_survey_date)) %>%
  arrange(transect_plot, quadrat, mort_survey_date) %>%
  group_by(transect_plot, quadrat) %>%
  mutate(cum_dead = cumsum(num_dead)) %>%
  ungroup() %>%
  mutate(count_check = cum_dead + missing_germs <= number_germ)

```

Germ cohort 2

```{r}
TM2_21_germ2 <- TM2_germ_2021 %>%
  filter(pick_color == "yellow")

TM2_21_germ2_mort <- TM2_21_germ2 %>%
  left_join(TM2_mort_20210226[,c(4:7,10:11, 18)]) %>%
  mutate(sum_check  = color_living_yellow + color_dead_yellow == number_germ)

TM2_21_germ2_mort_updated <- TM2_21_germ2_mort %>%
  select(-sum_check) %>%
  mutate(missing_germs = 0) %>%
  mutate(missing_germs = ifelse(transect_plot == "upper" & quadrat == 7.0, 1, missing_germs)) %>%
  mutate(sum_check  = color_living_yellow + color_dead_yellow + missing_germs== number_germ)

TM2_mort_germ2 <- TM2_mort_20210226 %>%
  rename(num_dead = color_dead_yellow) %>%
        select(mort_survey_date, transect_plot, quadrat, num_dead, mort_notes)

TM2_21_germ2_mort_long <- TM2_21_germ2_mort_updated %>%
  select(site, survey_year, transect_plot, quadrat, germ_survey_date, pick_color, number_germ, missing_germs, germ_notes) %>%
  left_join(TM2_mort_germ2)

TM2_21_germ2_day_since_germ <- TM2_21_germ2_mort_long %>%
   mutate(germ_survey_date = as.Date(germ_survey_date),
         mort_survey_date = as.Date(mort_survey_date),
         days_since_germ = as.integer(mort_survey_date - germ_survey_date)) %>%
  arrange(transect_plot, quadrat, mort_survey_date) %>%
  group_by(transect_plot, quadrat) %>%
  mutate(cum_dead = cumsum(num_dead)) %>%
  ungroup() %>%
  mutate(count_check = cum_dead + missing_germs <= number_germ)

```

Germ Cohort 3

```{r}
TM2_21_germ3 <- TM2_germ_2021 %>%
  filter(pick_color == "blue")

TM2_21_germ3_mort <- TM2_21_germ3 %>%
  left_join(TM2_mort_20210226[,c(4:7,12:13, 18)]) %>%
  mutate(sum_check  = color_living_blue + color_dead_blue == number_germ)

TM2_21_germ3_mort_updated <- TM2_21_germ3_mort %>%
  select(-sum_check) %>%
  mutate(missing_germs = 0) 

TM2_mort_germ3 <- TM2_mort_20210226 %>%
  rename(num_dead = color_dead_blue) %>%
        select(mort_survey_date, transect_plot, quadrat, num_dead, mort_notes)

TM2_21_germ3_mort_long <- TM2_21_germ3_mort_updated %>%
  select(site, survey_year, transect_plot, quadrat, germ_survey_date, pick_color, number_germ, missing_germs, germ_notes) %>%
  left_join(TM2_mort_germ3)

TM2_21_germ3_day_since_germ <- TM2_21_germ3_mort_long %>%
   mutate(germ_survey_date = as.Date(germ_survey_date),
         mort_survey_date = as.Date(mort_survey_date),
         days_since_germ = as.integer(mort_survey_date - germ_survey_date)) %>%
  arrange(transect_plot, quadrat, mort_survey_date) %>%
  group_by(transect_plot, quadrat) %>%
  mutate(cum_dead = cumsum(num_dead)) %>%
  ungroup() %>%
  mutate(count_check = cum_dead + missing_germs <= number_germ)

```

Germ Cohort 4

```{r}
TM2_21_germ4 <- TM2_germ_2021 %>%
  filter(pick_color == "orange")

TM2_21_germ4_mort <- TM2_21_germ4 %>%
  left_join(TM2_mort_20210226[,c(4:7,14:15, 18)]) %>%
  mutate(sum_check  = color_living_orange + color_dead_orange == number_germ)

TM2_21_germ4_mort_updated <- TM2_21_germ4_mort %>%
  select(-sum_check) %>%
  mutate(missing_germs = 0) 

TM2_mort_germ4 <- TM2_mort_20210226 %>%
  rename(num_dead = color_dead_orange) %>%
        select(mort_survey_date, transect_plot, quadrat, num_dead, mort_notes)

TM2_21_germ4_mort_long <- TM2_21_germ4_mort_updated %>%
  select(site, survey_year, transect_plot, quadrat, germ_survey_date, pick_color, number_germ, missing_germs, germ_notes) %>%
  left_join(TM2_mort_germ4)

TM2_21_germ4_day_since_germ <- TM2_21_germ4_mort_long %>%
   mutate(germ_survey_date = as.Date(germ_survey_date),
         mort_survey_date = as.Date(mort_survey_date),
         days_since_germ = as.integer(mort_survey_date - germ_survey_date)) %>%
  arrange(transect_plot, quadrat, mort_survey_date) %>%
  group_by(transect_plot, quadrat) %>%
  mutate(cum_dead = cumsum(num_dead)) %>%
  ungroup() %>%
  mutate(count_check = cum_dead + missing_germs <= number_germ)

```

Germ Cohort 5

```{r}
TM2_21_germ5 <- TM2_germ_2021 %>%
  filter(pick_color == "green")

TM2_21_germ5_mort <- TM2_21_germ5 %>%
  left_join(TM2_mort_20210226[,c(4:7,16:17,18)]) %>%
  mutate(sum_check  = color_living_green + color_dead_green == number_germ)

TM2_21_germ5_mort_updated <- TM2_21_germ5_mort %>%
  select(-sum_check) %>%
  mutate(missing_germs = 0) 

TM2_mort_germ5 <- TM2_mort_20210226 %>%
  rename(num_dead = color_dead_green) %>%
        select(mort_survey_date, transect_plot, quadrat, num_dead, mort_notes)

TM2_21_germ5_mort_long <- TM2_21_germ5_mort_updated %>%
  select(site, survey_year, transect_plot, quadrat, germ_survey_date, pick_color, number_germ, missing_germs, germ_notes) %>%
  left_join(TM2_mort_germ5)

TM2_21_germ5_day_since_germ <- TM2_21_germ5_mort_long %>%
   mutate(germ_survey_date = as.Date(germ_survey_date),
         mort_survey_date = as.Date(mort_survey_date),
         days_since_germ = as.integer(mort_survey_date - germ_survey_date)) %>%
  arrange(transect_plot, quadrat, mort_survey_date) %>%
  group_by(transect_plot, quadrat) %>%
  mutate(cum_dead = cumsum(num_dead)) %>%
  ungroup() %>%
  mutate(count_check = cum_dead + missing_germs <= number_germ)
```

#### Combine all germ surveys and create csv

```{r}
TM2_21_all_cohort_list <- list(TM2_21_germ1_day_since_germ, TM2_21_germ2_day_since_germ, TM2_21_germ3_day_since_germ, TM2_21_germ4_day_since_germ, TM2_21_germ5_day_since_germ)

TM2_21_all_germ_mort <- bind_rows(TM2_21_all_cohort_list)
#write.csv(TM2_21_all_germ_mort, "../output/TM2_21_all_germ_mort.csv", row.names = FALSE)
```

### TM2 2019-2020 Germination and Mortality

```{r}

```

### Load in WL and CP2 Germ and Mortality Data

```{r}
#WL and CP2 2021-2022 Surveys
all_WL1_germ_2022 <- read_excel("../input/Germination and Mortality data/WL1_germinationsurvey_2022_all.xlsx")
WL1_mort_20220401 <- read_excel("../input/Germination and Mortality data/WL1_mortality_20220401_completed.xlsx")
WL1_mort_20220427 <- read_excel("../input/Germination and Mortality data/WL1_mortalitysurvey_20220427_entered.xlsx.xlsx")
WL1_mort_20220518 <- read_excel("../input/Germination and Mortality data/WL1_mortality_20220518_finished_SHS.xlsx")
WL1_mort_20220701 <- read_excel("../input/Germination and Mortality data/WL1_mortalitysurvey_20220701_complete.xlsx")
all_WL2_WL3_germ_2022 <- read_excel("../input/Germination and Mortality data/WL2_3 germination 2022_all.xlsx")
all_CP2_germ_2022 <- read_excel("../input/Germination and Mortality data/CP2_germinationsurvey_2022_all_entered.xlsx")
CP2_mort_20220715 <- read_excel("../input/Germination and Mortality data/CP2_mortality_20220715_finished.xlsx")

#WL and CP2 2020-2021 Surveys
all_WL1_germ_2021 <- read_excel("../input/Germination and Mortality data/WL1_germinationsurvey_2021_all_C.xlsx")
WL1_mort_20210438 <- read_excel("../input/Germination and Mortality data/WL1_Mortality_20210428_verified.xlsx")
WL1_mort_20210513 <- read_excel("../input/Germination and Mortality data/WL1_Mortality_20210513_verified.xlsx")
WL1_mort_20210618 <- read_excel("../input/Germination and Mortality data/WL1_Mortality_20210618_verified.xlsx")
WL1_mort_20210702 <- read_excel("../input/Germination and Mortality data/WL1_mortality_20210702_verified.xlsx")
all_WL2_WL3_germ_2021 <- read_excel("../input/Germination and Mortality data/WL2and3_germinationsurvey_2021_all_finished_EC_SRA.xlsx")
WL2_WL3_mort_20210618 <- read_excel("../input/Germination and Mortality data/WL2_3_Mortality_20210618_verified.xlsx")
WL2_WL3_mort_20210702 <- read_excel("../input/Germination and Mortality data/WL2_WL3_Mortality_20210702_verified.xlsx")
WL2_WL3_mort_20210716 <- read_excel("../input/Germination and Mortality data/WL2_WL3_Mortality_20210716_verified.xlsx")

#WL and CP2 2017-2018 Surveys
WL1_germ_20180522 <- read_excel("../input/Germination and Mortality data/WL1_SpringGermSurvey_5.22.18_20180721_SLZ.xlsx")
WL1_germ_20180603 <- read_excel("../input/Germination and Mortality data/WL1_SpringGermSurvey_6.03.2018_fixed_EEB.xlsx")
WL_germ_20171102_20171118 <- read_excel("../input/Germination and Mortality data/Demog_WLCP_germ11.2.2017_20171219_AJC.xlsx") #Demog_WLCP_germ11.2.2017_20171106_AJC.xlsx contains just the 11/02/2017
WL1_CP2_germ_20180103 <- read_excel("../input/Germination and Mortality data/WLCP_FallGermSurvey_1.3.18_20180721_SLZ.xlsx") #contains WL1, WL2, WL3, and CP2 but WL2 and WL3 were not accessible so all NA. CP2 is all 0
WL1_WL2_WL3_CP2_germ_20180622 <- read_excel("../input/Germination and Mortality data/WLCP_SpringGermSurvey_6.22.18_20180721_SLZ.xlsx")
WL1_WL2_WL3_CP2_germ_20180613_20180604_20180608 <- read_excel("../input/Germination and Mortality data/WLCP_SpringGermSurvey_6.3.18_20180721_SLZ.xlsx")
WL2_WL3_mort_20180720 <- read_excel("../input/Germination and Mortality data/WLCP_SpringMortalitySurvey_7.20.18_SLZ.xlsx")
CP2_germ_20180603 <- read_excel("../input/Germination and Mortality data/CP2_germsurvey_6.03.2018_fixed_EEB.xlsx")
WL2_germ_20180603 <- read_excel("../input/Germination and Mortality data/WL2_germsurvey_6.03.2018_fixed_EEB.xlsx")
WL3_germ_20180603 <- read_excel("../input/Germination and Mortality data/WL3_germsurvey_6.03.2018_fixed_EEB.xlsx")

#WL and CP2 2018-2019 Surveys
WL1_germ_20190612 <- read_excel("../input/Germination and Mortality data/WL1_GermSurvey_6.12.19_edited9.4.19MM.xlsx")
WL1_germ_20190429 <- read_excel("../input/Germination and Mortality data/WL1_GermSurvey_4.29.19_edited9.3.19MM.xlsx")
WL2_WL3_germ_20190715 <- read_excel("../input/Germination and Mortality data/WL23_GermSurvey_7.15.19_edited9.4.19MM.xlsx")
WL2_WL3_CP2_germ_20190703 <- read_excel("../input/Germination and Mortality data/WL23+CP2_GermSurvey_7.3.19_edited9.3.19MM.xlsx")

#WL2 2016-2017 Survey
WL2_germ_20161027 <- read_excel("../input/Germination and Mortality data/Demog_WL2_fall_2016_germ_survey_10.27_20161026_AC.xlsx")
```

### WL1 2022 Germ and Mortality - need to rename mort survey notes column

#### Germ Cohort 1

```{r}
#pull out only the 1st germ cohort
WL1_22_germ1 <- all_WL1_germ_2022[,c("survey_year", "survey_date", "site", "transect_plot", "quadrat", "pick_color", "number_germ", "notes")] %>% 
  filter(pick_color == "green") %>% #filter for 1st germ cohort
  rename(germ_survey_date = survey_date,
         germ_notes = notes) #rename col names so we know what survey was done on that date

#merge in the 1st mort survey
WL1_22_germ1_mort1 <- WL1_22_germ1 %>%
  left_join(WL1_mort_20220401[,c("survey_date", "site", "transect_plot", "quadrat", "green_living", "green_dead")] %>% #join mort survey date, number living from 1st germ cohort, and number dead from 1st germ cohort
              rename(mort_survey_date = survey_date)) %>% #rename survey date to specify survey type
  mutate(sum_check = green_living + green_dead == number_germ) #add column to see if the sum of living and dead equals the number of germinates from germ survey

#there are 6 rows were the sum of living and dead doesn't equal the number of germinates
#how to resolve this? Look at the next mort survey to see what the sum of living and dead is and if it matches the living count from the 1st mort survey
#how to do this with code instead of just looking at the datasheets? I will just look and update for now because its faster
WL1_22_germ1_mort1_updated <- WL1_22_germ1_mort1 %>%
  select(-sum_check) %>%
  mutate(missing_germs = 0) %>%
  mutate(number_germ = ifelse(transect_plot == 1 & quadrat == 4.5, 9, number_germ), #looked at 20220427 mort survey which had 9 green living so update number_germ from 8 to 9
         number_germ = ifelse(transect_plot == 1 & quadrat == 8.0, 24, number_germ), #looked at 20220427 mort survey which had 24 green living, checked 20220518 mort survey which also had 24 green living, updated number_germ from 21 to 24 and green_living from 23 to 24
         green_living = ifelse(transect_plot == 1 & quadrat == 8.0, 24, green_living),
         number_germ = ifelse(transect_plot == 1 & quadrat == 8.5, 13, number_germ), #looked at 20220427 mort survey which had 12 green living so updated number_germ from 12 to 13
         number_germ = ifelse(transect_plot == 3 & quadrat == 1.0, 41, number_germ), #looked at 20220427 mort survey which had 41 total germ cohort 1 individuals so updated number_germ from 38 to 41
         green_living = ifelse(transect_plot == 3 & quadrat == 1.5, 66, green_living), #looked at 20220427 mort survey which had 66 total individuals so number_germ is correct and need to update green_living in the 1st mort survey from 67 to 66
         number_germ = ifelse(transect_plot == 3 & quadrat == 2.0, 69, number_germ), #looked at 20220427 mort survey which had 66 green living (1 more than the living count in the 20220401 mort survey), looked at the 20220518 mort survey which had 66 total individuals, update number_germ to 69 and green_living in 1st mort survey to 66, will also have to update green_living in 20220701 mort survey from 75 to 65
         green_living = ifelse(transect_plot == 3 & quadrat == 2.0, 66, green_living)) %>%
  mutate(sum_check  = green_living + green_dead + missing_germs == number_germ)

#Need to update mort survey to match
WL1_mort_20220701 <- WL1_mort_20220701 %>%
  mutate(living_green = ifelse(transect_plot == 3 & quadrat == 2.0, 65, green_living))

#create a list of all mort surveys
WL1_22_mort_list <- list(WL1_mort_20220401, WL1_mort_20220427, WL1_mort_20220518, WL1_mort_20220701)

WL1_22_all_mort_germ1 <- WL1_22_mort_list %>%
  map(~ .x %>%
        rename(mort_survey_date = survey_date,
               mort_notes = notes,
               num_dead = green_dead) %>%
        select(mort_survey_date, transect_plot, quadrat, num_dead, mort_notes)) %>%
  bind_rows() %>%
  mutate(quadrat = if_else(quadrat == 4.4, 4.5, quadrat)) #need to update 20220518 transect_plot: 1 quadrat: 4.4 to quadrat 4.5

WL1_22_germ1_long_prep <- WL1_22_germ1_mort1_updated %>%
  select(site, survey_year, transect_plot, quadrat, germ_survey_date, pick_color, number_germ, missing_germs, germ_notes)

WL1_22_germ1_all_mort_long <- WL1_22_germ1_long_prep %>%
  left_join(WL1_22_all_mort_germ1)

#only keep number_germ and num_dead but calculate cum_dead- could subtract cum_dead from number_germ to get remaining living if needed
WL1_22_germ1_day_since_germ <- WL1_22_germ1_all_mort_long %>%
  mutate(germ_survey_date = as.Date(germ_survey_date),
         mort_survey_date = as.Date(mort_survey_date),
         days_since_germ = as.integer(mort_survey_date - germ_survey_date)) %>%
  arrange(transect_plot, quadrat, mort_survey_date) %>%
  group_by(transect_plot, quadrat) %>%
  mutate(cum_dead = cumsum(num_dead)) %>%
  ungroup() %>%
  mutate(count_check = cum_dead + missing_germs <= number_germ)
```

#### Germ Cohort 2

```{r}
WL1_22_germ2 <- all_WL1_germ_2022[,3:ncol(all_WL1_germ_2022)] %>%
  filter(pick_color == "blue") %>%
  rename(germ_survey_date = survey_date,
         germ_notes = notes)

WL1_22_germ2_mort1 <- WL1_22_germ2 %>%
  left_join(WL1_mort_20220401[,c(4:7,10:11,14)] %>%
              rename(mort_survey_date = survey_date)) %>%
  mutate(sum_check  = blue_living + blue_dead == number_germ) #2 rows have "missing" germs but I'll leave those rows for now

WL1_22_germ2_mort1_updated <- WL1_22_germ2_mort1 %>%
  select(-sum_check) %>%
  mutate(missing_germs = 0) %>%
  mutate(missing_germs = ifelse(transect_plot == 1 & quadrat == 0.0, 2, missing_germs),
         missing_germs = ifelse(transect_plot == 1 & quadrat == 6.5, 1, missing_germs)) %>%
  mutate(sum_check  = blue_living + blue_dead + missing_germs == number_germ)

WL1_22_all_mort_germ2 <- WL1_22_mort_list %>%
  map(~ .x %>%
        rename(mort_survey_date = survey_date,
               mort_notes = notes,
               num_dead = blue_dead) %>%
        select(mort_survey_date, transect_plot, quadrat, num_dead, mort_notes)) %>%
  bind_rows() %>%
  mutate(quadrat = if_else(quadrat == 4.4, 4.5, quadrat)) #update 20220518 transect_plot: 1 quadrat:4.4 to quadrat 4.5

WL1_22_germ2_long_prep <- WL1_22_germ2_mort1_updated %>%
  select(site, survey_year, transect_plot, quadrat, germ_survey_date, pick_color, number_germ, missing_germs, germ_notes)

WL1_22_germ2_all_mort_long <- WL1_22_germ2_long_prep %>%
  left_join(WL1_22_all_mort_germ2)

WL1_22_germ2_day_since_germ <- WL1_22_germ2_all_mort_long %>%
  mutate(germ_survey_date = as.Date(germ_survey_date),
         mort_survey_date = as.Date(mort_survey_date),
         days_since_germ = as.integer(mort_survey_date - germ_survey_date)) %>%
  arrange(transect_plot, quadrat, mort_survey_date) %>%
  group_by(transect_plot, quadrat) %>%
  mutate(cum_dead = cumsum(num_dead)) %>%
  ungroup() %>%
  mutate(count_check = cum_dead + missing_germs <= number_germ)
```

#### Germ Cohort 3

```{r}
WL1_22_germ3 <- all_WL1_germ_2022[,3:ncol(all_WL1_germ_2022)] %>%
  filter(pick_color == "red") %>%
  rename(germ_survey_date = survey_date,
         germ_notes = notes)

WL1_22_germ3_mort1 <- WL1_22_germ3 %>%
  left_join(WL1_mort_20220401[,c(4:7,12:14)] %>%
              rename(mort_survey_date = survey_date)) %>%
  mutate(sum_check  = red_living + red_dead == number_germ) 

WL1_22_germ3_mort1_updated <- WL1_22_germ3_mort1 %>%
  select(-sum_check) %>%
  mutate(missing_germs = 0) %>%
  mutate(missing_germs = ifelse(transect_plot == 1 & quadrat == 0.5, 1, missing_germs),
         missing_germs = ifelse(transect_plot == 1 & quadrat == 2.5, 1, missing_germs),
         number_germ = ifelse(transect_plot == 2 & quadrat == 6.0, 4, number_germ),
         missing_germs = ifelse(transect_plot == 3 & quadrat == 3.0, 7, missing_germs),
         missing_germs = ifelse(transect_plot == 3 & quadrat == 4.0, 1, missing_germs),
         missing_germs = ifelse(transect_plot == 3 & quadrat == 4.5, 1, missing_germs)) %>%
   mutate(sum_check  = red_living + red_dead + missing_germs == number_germ)

WL1_22_all_mort_germ3 <- WL1_22_mort_list %>%
  map(~ .x %>%
        rename(mort_survey_date = survey_date,
               mort_notes = notes,
               num_dead = red_dead) %>%
        select(mort_survey_date, transect_plot, quadrat, num_dead, mort_notes)) %>%
  bind_rows() %>%
  mutate(quadrat = if_else(quadrat == 4.4, 4.5, quadrat)) #need to update 20220518 transect_plot: 1 quadrat:4.4 to quadrat 4.5

WL1_22_germ3_long_prep <- WL1_22_germ3_mort1_updated %>%
  select(site, survey_year, transect_plot, quadrat, germ_survey_date, pick_color, number_germ, missing_germs, germ_notes)

WL1_22_germ3_all_mort_long <- WL1_22_germ3_long_prep %>%
  left_join(WL1_22_all_mort_germ3)

WL1_22_germ3_day_since_germ <- WL1_22_germ3_all_mort_long %>%
  mutate(germ_survey_date = as.Date(germ_survey_date),
         mort_survey_date = as.Date(mort_survey_date),
         days_since_germ = as.integer(mort_survey_date - germ_survey_date)) %>%
  arrange(transect_plot, quadrat, mort_survey_date) %>%
  group_by(transect_plot, quadrat) %>%
  mutate(cum_dead = cumsum(num_dead)) %>%
  ungroup() %>%
  mutate(count_check = cum_dead + missing_germs <= number_germ)
```

#### Germ Cohort 4

```{r}
WL1_22_germ4 <- all_WL1_germ_2022[,3:ncol(all_WL1_germ_2022)] %>%
  filter(pick_color == "yellow") %>%
  rename(germ_survey_date = survey_date,
         germ_notes = notes)

WL1_22_germ4_mort1 <- WL1_22_germ4 %>%
  left_join(WL1_mort_20220427[,c(4:7,14:ncol(WL1_mort_20220427))] %>%
              rename(mort_survey_date = survey_date)) %>%
  mutate(sum_check  = yellow_living + yellow_dead == number_germ) %>% #everything matches - probably because there were few germs
  select(-sum_check) %>%
  mutate(missing_germs = 0)

#This germ survey took place after some of the mort surveys so I need to only include the mort surveys with this cohort
WL1_22_germ4_mort_list <- list(WL1_mort_20220427, WL1_mort_20220518, WL1_mort_20220701)

WL1_22_all_mort_germ4 <- WL1_22_germ4_mort_list %>%
  map(~ .x %>%
        rename(mort_survey_date = survey_date,
               mort_notes = notes,
               num_dead = yellow_dead) %>%
        select(mort_survey_date, transect_plot, quadrat, num_dead, mort_notes)) %>%
  bind_rows() %>%
  mutate(quadrat = if_else(quadrat == 4.4, 4.5, quadrat)) #need to update 20220518 transect_plot: 1 quadrat:4.4 to quadrat 4.5

WL1_22_germ4_long_prep <- WL1_22_germ4_mort1 %>%
  select(site, survey_year, transect_plot, quadrat, germ_survey_date, pick_color, number_germ, missing_germs, germ_notes)

WL1_22_germ4_all_mort_long <- WL1_22_germ4_long_prep %>%
  left_join(WL1_22_all_mort_germ4)

WL1_22_germ4_day_since_germ <- WL1_22_germ4_all_mort_long %>%
  mutate(germ_survey_date = as.Date(germ_survey_date),
         mort_survey_date = as.Date(mort_survey_date),
         days_since_germ = as.integer(mort_survey_date - germ_survey_date)) %>%
  arrange(transect_plot, quadrat, mort_survey_date) %>%
  group_by(transect_plot, quadrat) %>%
  mutate(cum_dead = cumsum(num_dead)) %>%
  ungroup()%>%
  mutate(count_check = cum_dead + missing_germs <= number_germ)
```

#### Germ Cohort 5

```{r}
WL1_22_germ5 <- all_WL1_germ_2022[,3:ncol(all_WL1_germ_2022)] %>%
  filter(pick_color == "orange") %>%
  rename(germ_survey_date = survey_date,
         germ_notes = notes)

WL1_22_germ5_mort1 <- WL1_22_germ5 %>%
  left_join(WL1_mort_20220518[, c(4:7, 16:ncol(WL1_mort_20220518))] %>%
      rename(mort_survey_date = survey_date) %>%
      mutate(quadrat = if_else(quadrat == 4.4, 4.5, quadrat))) %>% #need to update 20220518 transect_plot: 1 quadrat:4.4 to quadrat 4.5) 
  mutate(sum_check = orange_living + orange_dead == number_germ)

WL1_22_germ5_mort1_updated <- WL1_22_germ5_mort1 %>%
  select(-sum_check) %>%
  mutate(missing_germs = 0) %>%
  mutate(number_germ = ifelse(transect_plot == 2 & quadrat == 3.0, 0, number_germ), #the germ wasn't STTO
         number_germ = ifelse(transect_plot == 2 & quadrat == 4.0, 1, number_germ),
         number_germ = ifelse(transect_plot == 2 & quadrat == 4.5, 1, number_germ),
         missing_germs = ifelse(transect_plot == 2 & quadrat == 6.0, 1, missing_germs),
         missing_germs = ifelse(transect_plot == 2 & quadrat == 6.5, 1, missing_germs),
         number_germ = ifelse(transect_plot == 3 & quadrat == 1.0, 1, number_germ), # 1 germ wasn't STTO
         number_germ = ifelse(transect_plot == 3 & quadrat == 2.0, 0, number_germ), # 1 germ wasn't STTO
         number_germ = ifelse(transect_plot == 3 & quadrat == 3.5, 1, number_germ)) %>% # 1 germ wasn't STTO
  mutate(sum_check = orange_living + orange_dead + missing_germs == number_germ)

#This germ survey took place after some of the mort surveys so I need to only include the mort surveys with this cohort
WL1_22_germ5_mort_list <- list(WL1_mort_20220518, WL1_mort_20220701)

WL1_22_all_mort_germ5 <- WL1_22_germ5_mort_list %>%
  map(~ .x %>%
        rename(mort_survey_date = survey_date,
               mort_notes = notes,
               num_dead = orange_dead) %>%
        select(mort_survey_date, transect_plot, quadrat, num_dead, mort_notes)) %>%
  bind_rows() %>%
  mutate(quadrat = if_else(quadrat == 4.4, 4.5, quadrat)) #need to update 20220518 transect_plot: 1 quadrat:4.4 to quadrat 4.5)

WL1_22_germ5_long_prep <- WL1_22_germ5_mort1_updated %>%
  select(site, survey_year, transect_plot, quadrat, germ_survey_date, pick_color, number_germ, missing_germs, germ_notes)

WL1_22_germ5_all_mort_long <- WL1_22_germ5_long_prep %>%
  left_join(WL1_22_all_mort_germ5)

WL1_22_germ5_day_since_germ <- WL1_22_germ5_all_mort_long %>%
  mutate(germ_survey_date = as.Date(germ_survey_date),
         mort_survey_date = as.Date(mort_survey_date),
         days_since_germ = as.integer(mort_survey_date - germ_survey_date)) %>%
   mutate(num_dead = ifelse(transect_plot == 2 & quadrat == 6.5 & mort_survey_date == "2022-07-01", 0, num_dead),
          num_dead = ifelse(transect_plot == 2 & quadrat == 7.0 & mort_survey_date == "2022-07-01", 0, num_dead)) %>% # number_germ was 0 initially and previous mort survey didn't have any individuals so these will be changed to 0
  arrange(transect_plot, quadrat, mort_survey_date) %>%
  group_by(transect_plot, quadrat) %>%
  mutate(cum_dead = cumsum(num_dead)) %>%
  ungroup() %>%
  mutate(count_check = cum_dead + missing_germs <= number_germ)
```

#### Combine All Germ Cohorts and Save CSV

```{r}
WL1_22_all_cohort_list <- list(WL1_22_germ1_day_since_germ, WL1_22_germ2_day_since_germ, WL1_22_germ3_day_since_germ, WL1_22_germ4_day_since_germ, WL1_22_germ5_day_since_germ)

WL1_22_all_germ_mort <- bind_rows(WL1_22_all_cohort_list)
#write.csv(WL1_22_all_germ_mort, "../output/WL1_22_all_germ_mort.csv", row.names = FALSE)
```

### WL2 and WL3 2022 Germ (No Mortality Surveys)

```{r}
WL2_22_all_germ <- all_WL2_WL3_germ_2022 %>%
  filter(site == "WL2") %>%
  rename(germ_survey_date = survey_date,
         germ_notes = notes) %>%
  mutate(missing_germs = 0)

WL3_22_all_germ <- all_WL2_WL3_germ_2022 %>%
  filter(site == "WL3") %>%
  rename(germ_survey_date = survey_date,
         germ_notes = notes) %>%
  mutate(missing_germs = 0)

#write.csv(WL2_22_all_germ, "../output/WL2_22_all_germ_no_mort.csv")
#write.csv(WL3_22_all_germ, "../output/WL3_22_all_germ_no_mort.csv")
```

### CP2 2022 Germ and Mortality

#### Germ Cohort 1

```{r}
CP2_22_germ1 <- all_CP2_germ_2022 %>%
  filter(pick_color == "red") %>%
  rename(germ_survey_date = survey_date,
         germ_notes = notes)

CP2_22_germ1_mort1 <- CP2_22_germ1 %>%
  left_join(CP2_mort_20220715[,c(4:9,14)] %>%
              rename(mort_survey_date = survey_date,
                     mort_notes = notes)) %>%
  mutate(sum_check  = red_living + red_dead == number_germ)

CP2_22_germ1_mort1_updated <- CP2_22_germ1_mort1 %>%
  select(-sum_check)%>%
  mutate(missing_germs = 0) %>%
  mutate(number_germ = ifelse(transect_plot == "T4" & quadrat == "T4", 10, number_germ),
         number_germ = ifelse(transect_plot == "T9" & quadrat == "T9", 1, number_germ),
         number_germ = ifelse(transect_plot == "T12" & quadrat == "T12", 6, number_germ),
         number_germ = ifelse(transect_plot == "T13" & quadrat == "T13", 15, number_germ)) %>%
  mutate(sum_check = red_living + red_dead + missing_germs == number_germ)

CP2_22_all_mort_germ1 <- CP2_mort_20220715 %>%
  rename(mort_survey_date = survey_date,
         mort_notes = notes,
         num_dead = red_dead) %>%
  select(mort_survey_date, transect_plot, quadrat, num_dead, mort_notes)

CP2_22_germ1_long_prep <- CP2_22_germ1_mort1_updated %>%
  select(site, survey_year, transect_plot, quadrat, germ_survey_date, pick_color, number_germ, missing_germs, germ_notes)

CP2_22_germ1_mort_long <- CP2_22_germ1_long_prep %>%
  left_join(CP2_22_all_mort_germ1)

CP2_22_germ1_day_since_germ <- CP2_22_germ1_mort_long %>%
  mutate(germ_survey_date = as.Date(germ_survey_date),
         mort_survey_date = as.Date(mort_survey_date),
         days_since_germ = as.integer(mort_survey_date - germ_survey_date)) %>%
  arrange(transect_plot, quadrat, mort_survey_date) %>%
  group_by(transect_plot, quadrat) %>%
  mutate(cum_dead = cumsum(num_dead)) %>%
  ungroup()%>%
  mutate(count_check = cum_dead + missing_germs <= number_germ)
```

#### Germ Cohort 2

```{r}
CP2_22_germ2 <- all_CP2_germ_2022 %>%
  filter(pick_color == "blue")%>%
  rename(germ_survey_date = survey_date,
         germ_notes = notes)

CP2_22_germ2_mort1 <- CP2_22_germ2 %>%
  left_join(CP2_mort_20220715[,c(4:7, 10:11 ,14)] %>%
              rename(mort_survey_date = survey_date,
                     mort_notes = notes)) %>%
  mutate(sum_check  = blue_living + blue_dead == number_germ) %>% #all correct
  mutate(missing_germs = 0)

CP2_22_all_mort_germ2 <- CP2_mort_20220715 %>%
  rename(mort_survey_date = survey_date,
         mort_notes = notes,
         num_dead = blue_dead) %>%
  select(mort_survey_date, transect_plot, quadrat, num_dead, mort_notes)

CP2_22_germ2_long_prep <- CP2_22_germ2_mort1 %>%
  select(site, survey_year, transect_plot, quadrat, germ_survey_date, pick_color, number_germ, missing_germs, germ_notes)

CP2_22_germ2_mort_long <- CP2_22_germ2_long_prep %>%
  left_join(CP2_22_all_mort_germ2)

CP2_22_germ2_day_since_germ <- CP2_22_germ2_mort_long %>%
  mutate(germ_survey_date = as.Date(germ_survey_date),
         mort_survey_date = as.Date(mort_survey_date),
         days_since_germ = as.integer(mort_survey_date - germ_survey_date)) %>%
  arrange(transect_plot, quadrat, mort_survey_date) %>%
  group_by(transect_plot, quadrat) %>%
  mutate(cum_dead = cumsum(num_dead)) %>%
  ungroup()%>%
  mutate(count_check = cum_dead + missing_germs <= number_germ)
```

#### Germ Cohort 3

```{r}
CP2_22_germ3 <- all_CP2_germ_2022 %>%
  filter(pick_color == "white")%>%
  rename(germ_survey_date = survey_date,
         germ_notes = notes)

CP2_22_germ3_mort1 <- CP2_22_germ3 %>%
  left_join(CP2_mort_20220715[,c(4:7,12:14)] %>%
              rename(mort_survey_date = survey_date,
                     mort_notes = notes)) %>%
  mutate(sum_check  = white_living + white_dead == number_germ) %>% #all correct
  mutate(missing_germs = 0)

CP2_22_all_mort_germ3 <- CP2_mort_20220715 %>%
  rename(mort_survey_date = survey_date,
         mort_notes = notes,
         num_dead = white_dead) %>%
  select(mort_survey_date, transect_plot, quadrat, num_dead, mort_notes)

CP2_22_germ3_long_prep <- CP2_22_germ3_mort1 %>%
  select(site, survey_year, transect_plot, quadrat, germ_survey_date, pick_color, number_germ, missing_germs, germ_notes)

CP2_22_germ3_mort_long <- CP2_22_germ3_long_prep %>%
  left_join(CP2_22_all_mort_germ3)

CP2_22_germ3_day_since_germ <- CP2_22_germ3_mort_long %>%
  mutate(germ_survey_date = as.Date(germ_survey_date),
         mort_survey_date = as.Date(mort_survey_date),
         days_since_germ = as.integer(mort_survey_date - germ_survey_date)) %>%
  arrange(transect_plot, quadrat, mort_survey_date) %>%
  group_by(transect_plot, quadrat) %>%
  mutate(cum_dead = cumsum(num_dead)) %>%
  ungroup()%>%
  mutate(count_check = cum_dead + missing_germs <= number_germ)
```

#### Combine All Germ Cohorts and Save CSV

```{r}
CP2_22_all_cohort_list <- list(CP2_22_germ1_day_since_germ, CP2_22_germ2_day_since_germ, CP2_22_germ3_day_since_germ)
  
CP2_22_all_germ_mort <- bind_rows(CP2_22_all_cohort_list)
#write.csv(CP2_22_all_germ_mort, "../output/CP2_22_all_germ_mort.csv", row.names = FALSE)
```

### WL1 2021 Germ and Mortality - notes say plots are off so might not be able to use since it is hard to track where germs went

#### Germ Cohort 1

```{r}
WL1_21_germ1 <- all_WL1_germ_2021 %>%
  filter(pick_color == "red") %>%
  rename(germ_survey_date = survey_date,
         germ_notes = notes)

WL1_21_germ1_mort1 <- WL1_21_germ1 %>%
  left_join(WL1_mort_20210438[,c(4:9,18)] %>%
              rename(mort_survey_date = survey_date)) %>%
  mutate(sum_check  = color_living_RED + color_dead_RED == number_germ)

WL1_21_germ1_mort1_updated <- WL1_21_germ1_mort1 %>%
  select(-sum_check) %>%
  mutate(missing_germs = 0) %>%
  mutate(missing_germs = ifelse(transect_plot == 1 & quadrat == 5.5, 3, missing_germs),
         missing_germs = ifelse(transect_plot == 2 & quadrat == 1.5, 1, missing_germs),
         missing_germs = ifelse(transect_plot == 2 & quadrat == 2.5, 2, missing_germs),
         number_germ = ifelse(transect_plot == 2 & quadrat == 3.5, 8, number_germ),
         number_germ = ifelse(transect_plot == 2 & quadrat == 6.5, 10, number_germ),
         number_germ = ifelse(transect_plot == 2 & quadrat == 7.0, 7, number_germ), #notes say plots off for 6.5 and 7.0 but later mort surveys follow living counts so I will change number_germ to reflect this
         number_germ = ifelse(transect_plot == 3 & quadrat == 3.0, 14, number_germ),
         number_germ = ifelse(transect_plot == 3 & quadrat == 4.0, 8, number_germ)) %>%
  mutate(sum_check = color_living_RED + color_dead_RED + missing_germs == number_germ)

WL1_mort_20210513 <- WL1_mort_20210513 %>%
  rename(color_living_RED = color_living_red,
         color_dead_RED = color_dead_red)

WL1_mort_20210702 <- WL1_mort_20210702 %>%
  rename(color_living_RED = color_living_red,
         color_dead_RED = color_dead_red)

WL1_21_germ1_mort_list <- list(WL1_mort_20210438, WL1_mort_20210513, WL1_mort_20210618, WL1_mort_20210702)

WL1_21_all_mort_germ1 <- WL1_21_germ1_mort_list %>%
  map(~ .x %>%
        rename(mort_survey_date = survey_date,
               mort_notes = notes,
               num_dead = color_dead_RED) %>%
        select(mort_survey_date, transect_plot, quadrat, num_dead, mort_notes)) %>%
  bind_rows()

WL1_21_germ1_long_prep <- WL1_21_germ1_mort1_updated %>%
  select(site, survey_year, transect_plot, quadrat, germ_survey_date, pick_color, number_germ, missing_germs, germ_notes)

WL1_21_germ1_all_mort_long <- WL1_21_germ1_long_prep %>%
  left_join(WL1_21_all_mort_germ1)

WL1_21_germ1_day_since_germ <- WL1_21_germ1_all_mort_long %>%
  mutate(germ_survey_date = as.Date(germ_survey_date),
         mort_survey_date = as.Date(mort_survey_date),
         days_since_germ = as.integer(mort_survey_date - germ_survey_date)) %>%
  arrange(transect_plot, quadrat, mort_survey_date) %>%
  group_by(transect_plot, quadrat) %>%
  mutate(cum_dead = cumsum(num_dead)) %>%
  ungroup()%>%
  mutate(count_check = cum_dead + missing_germs <= number_germ)

WL1_mort_20210513 
WL1_mort_20210618
WL1_mort_20210702

```

#### Germ Cohort 2

```{r}
WL1_21_germ2 <- all_WL1_germ_2021 %>%
  filter(pick_color == "yellow") %>%
  rename(germ_survey_date = survey_date,
         germ_notes = notes)

WL1_21_germ2_mort1 <- WL1_21_germ2 %>%
  left_join(WL1_mort_20210438[,c(4:7,10:11,18)] %>%
              rename(mort_survey_date = survey_date)) %>%
  mutate(sum_check  = color_living_YELLOW + color_dead_YELLOW == number_germ)

WL1_21_germ2_mort1_updated <- WL1_21_germ2_mort1 %>%
  select(-sum_check) %>%
  mutate(missing_germs = 0) %>%
  mutate(number_germ = ifelse(transect_plot == 1 & quadrat == 0.0, 13, number_germ),
         number_germ = ifelse(transect_plot == 1 & quadrat == 0.5, 8, number_germ),
         number_germ = ifelse(transect_plot == 1 & quadrat == 2.0, 3, number_germ),
         missing_germs = ifelse(transect_plot == 1 & quadrat == 5.5, 1, missing_germs),
         number_germ = ifelse(transect_plot == 2 & quadrat == 0.0, 5, number_germ),
         number_germ = ifelse(transect_plot == 2 & quadrat == 4.0, 4, number_germ),
         missing_germs = ifelse(transect_plot == 2 & quadrat == 4.0, 1, missing_germs),
         number_germ = ifelse(transect_plot == 2 & quadrat == 5.0, 3, number_germ),
         number_germ = ifelse(transect_plot == 2 & quadrat == 6.0, 4, number_germ),
         number_germ = ifelse(transect_plot == 2 & quadrat == 7.0, 2, number_germ)) %>%
  mutate(sum_check = color_living_YELLOW + color_dead_YELLOW + missing_germs == number_germ)

WL1_mort_20210513 <- WL1_mort_20210513 %>%
  rename(color_living_YELLOW = color_living_yellow,
         color_dead_YELLOW = color_dead_yellow) %>%
  mutate(color_dead_YELLOW = ifelse(transect_plot == 3 & quadrat == 2.0, 11, color_dead_YELLOW),
         color_dead_YELLOW = ifelse(transect_plot == 3 & quadrat == 3.5, 1, color_dead_YELLOW))

WL1_mort_20210618 <- WL1_mort_20210618 %>%
  mutate(color_dead_YELLOW = ifelse(transect_plot == 2 & quadrat == 5.0 , 1, color_dead_YELLOW))

WL1_mort_20210702 <- WL1_mort_20210702 %>%
  rename(color_living_YELLOW = color_living_yellow,
         color_dead_YELLOW = color_dead_yellow)

WL1_21_germ2_mort_list <- list(WL1_mort_20210438, WL1_mort_20210513, WL1_mort_20210618, WL1_mort_20210702)

WL1_21_all_mort_germ2 <- WL1_21_germ2_mort_list %>%
  map(~ .x %>%
        rename(mort_survey_date = survey_date,
               mort_notes = notes,
               num_dead = color_dead_YELLOW) %>%
        select(mort_survey_date, transect_plot, quadrat, num_dead, mort_notes)) %>%
  bind_rows()

WL1_21_germ2_long_prep <- WL1_21_germ2_mort1_updated %>%
  select(site, survey_year, transect_plot, quadrat, germ_survey_date, pick_color, number_germ, missing_germs, germ_notes)

WL1_21_germ2_all_mort_long <- WL1_21_germ2_long_prep %>%
  left_join(WL1_21_all_mort_germ2)

WL1_21_germ2_day_since_germ <- WL1_21_germ2_all_mort_long %>%
  mutate(germ_survey_date = as.Date(germ_survey_date),
         mort_survey_date = as.Date(mort_survey_date),
         days_since_germ = as.integer(mort_survey_date - germ_survey_date)) %>%
  arrange(transect_plot, quadrat, mort_survey_date) %>%
  group_by(transect_plot, quadrat) %>%
  mutate(cum_dead = cumsum(num_dead)) %>%
  ungroup()%>%
  mutate(count_check = cum_dead + missing_germs <= number_germ) %>% #there are still two rows from transect_plot: 1, quadrat 0.0 where the counts across mortality surveys don't match/make sense, can filter them out for now
  filter(count_check == TRUE)
```

#### Germ Cohort 3 - there is no orange in the mortality surveys, also mort surveys have blue and green but germ surveys don't have those colors

```{r}
WL1_21_germ3 <- all_WL1_germ_2021 %>%
  filter(pick_color == "orange") %>%
  rename(germ_survey_date = survey_date,
         germ_notes = notes)
```

#### Combine Germ Cohorts 1 and 2 and Save CSV - still need to figure out what is happening with the orange, blue, and green cohorts

```{r}
WL1_21_all_cohort_list <- list(WL1_21_germ1_day_since_germ, WL1_21_germ2_day_since_germ)
  
WL1_21_all_germ_mort <- bind_rows(WL1_21_all_cohort_list)
#write.csv(WL1_21_all_germ_mort, "../output/WL1_21_all_germ_mort", row.names = FALSE)
```

### WL2 2021 Germ and Mortality

```{r}
#Need to correct transect plot names for 202C and 202D, also need to update site name to WL3 for those that say transect plot 1

all_WL2_WL3_germ_2021 <- all_WL2_WL3_germ_2021 %>%
  mutate(transect_plot = ifelse(transect_plot == "201C" & quadrat == 1.0 & number_germ == 2, "202C", transect_plot),
         transect_plot = ifelse(transect_plot == "201C" & quadrat == 2.0 & number_germ == 0, "202C", transect_plot),
         transect_plot = ifelse(transect_plot == "201C" & quadrat == 3.0 & number_germ == 3, "202C", transect_plot),
         transect_plot = ifelse(transect_plot == "201C" & quadrat == 4.0 & number_germ == 3, "202C", transect_plot),
        # transect_plot = replace(transect_plot, which(transect_plot == "201D" & quadrat == 0.0)[1], "202D"), #the number_germ and all other columns are the same for the two 201D 0.0 rows so I'm just selecting 1 of the rows and updating transect_plot to 202D - I'm not sure this is working
         transect_plot = ifelse(transect_plot == "201D" & quadrat == 1.0 & number_germ == 13, "202D", transect_plot),
         transect_plot = ifelse(transect_plot == "201D" & quadrat == 2.0 & number_germ == 0, "202D", transect_plot),
         transect_plot = ifelse(transect_plot == "201D" & quadrat == 3.0 & number_germ == 1, "202D", transect_plot),
         transect_plot = ifelse(transect_plot == "201D" & quadrat == 4.0 & number_germ == 0, "202D", transect_plot),
         site = ifelse(transect_plot == 1, "WL3", site))

```

#### Germ Cohort 0

```{r}
WL2_21_germ0 <- all_WL2_WL3_germ_2021 %>%
  filter(pick_color == "blue", #cohort 0 has true leaves
         site == "WL2") %>%  
  filter(!(transect_plot == "202D" & number_germ == 0)) %>% #2 202D 4.0 rows, deleted the one with the germ count that didn't match the mort survey counts
  rename(germ_survey_date = survey_date,
         germ_notes = notes) %>%
  mutate(quadrat = ifelse(transect_plot == "202A" & number_germ == 21, 4.5, quadrat)) %>%
  mutate(transect_plot = replace(transect_plot, 14, "201C"))

WL2_21_germ0_mort1 <- WL2_21_germ0 %>%
  left_join(WL2_WL3_mort_20210618[,c(4:9,18)] %>%
              rename(mort_survey_date = survey_date,
                     mort_notes = notes)) %>%
  mutate(sum_check  = color_living_blue + color_dead_blue == number_germ)

WL2_21_germ0_mort1_updated <- WL2_21_germ0_mort1 %>%
  select(-sum_check) %>%
  mutate(missing_germs = 0) %>%
  mutate(number_germ = ifelse(transect_plot == "201A" & quadrat == 0.0, 2, number_germ),
         number_germ = ifelse(transect_plot == "201B" & quadrat == 4.0, 2, number_germ),
         missing_germs = ifelse(transect_plot == "201C" & quadrat == 0.0, 1, missing_germs),
         number_germ = ifelse(transect_plot == "201D" & quadrat == 1.0, 9, number_germ),
         number_germ = ifelse(transect_plot == "202B" & quadrat == 3.0, 7, number_germ)) %>%
  mutate(sum_check  = color_living_blue + color_dead_blue + missing_germs == number_germ)

WL2_WL3_mort_20210716 <- WL2_WL3_mort_20210716 %>%
  mutate(color_dead_blue = ifelse(transect_plot == "202D" & quadrat == 0.0, 2, color_dead_blue))

WL2_21_mort_list <- list(WL2_WL3_mort_20210618, WL2_WL3_mort_20210702, WL2_WL3_mort_20210716)

WL2_21_all_mort_germ0 <- WL2_21_mort_list %>%
  map(~ .x %>%
        filter(site == "WL2") %>%
        rename(mort_survey_date = survey_date,
               mort_notes = notes,
               num_dead = color_dead_blue) %>%
        select(mort_survey_date, transect_plot, quadrat, num_dead, mort_notes)) %>%
  bind_rows() 

WL2_21_germ0_long_prep <- WL2_21_germ0_mort1_updated %>%
  select(site, survey_year, transect_plot, quadrat, germ_survey_date, pick_color, number_germ, missing_germs, germ_notes)

WL2_21_germ0_all_mort_long <- WL2_21_germ0_long_prep %>%
  left_join(WL2_21_all_mort_germ0)

WL2_21_germ0_day_since_germ <- WL2_21_germ0_all_mort_long %>%
  mutate(germ_survey_date = as.Date(germ_survey_date),
         mort_survey_date = as.Date(mort_survey_date),
         days_since_germ = as.integer(mort_survey_date - germ_survey_date)) %>%
  arrange(transect_plot, quadrat, mort_survey_date) %>%
  group_by(transect_plot, quadrat) %>%
  mutate(cum_dead = cumsum(num_dead)) %>%
  ungroup() %>%
  mutate(count_check = cum_dead + missing_germs <= number_germ)
```

#### Germ Cohort 1

```{r}
WL2_21_germ1 <- all_WL2_WL3_germ_2021 %>%
  filter(pick_color == "yellow", 
         site == "WL2") %>%  
  rename(germ_survey_date = survey_date,
         germ_notes = notes)

WL2_21_germ1_mort1 <- WL2_21_germ1 %>%
  left_join(WL2_WL3_mort_20210618[,c(4:7, 10:11, 18)] %>%
              rename(mort_survey_date = survey_date,
                     mort_notes = notes)) %>%
  mutate(sum_check  = color_living_yellow + color_dead_yellow == number_germ)

WL2_21_germ1_mort1_updated <- WL2_21_germ1_mort1 %>%
  select(-sum_check) %>%
  mutate(missing_germs = 0) %>%
  mutate(missing_germs = ifelse(transect_plot == "201C" & quadrat == 1.0, 1, missing_germs),
         number_germ = ifelse(transect_plot == "201C" & quadrat == 4.0, 11, number_germ),
         missing_germs = ifelse(transect_plot == "202A" & quadrat == 0.0, 1, missing_germs),
         missing_germs = ifelse(transect_plot == "202A" & quadrat == 4.0, 1, missing_germs), #notes say 2 missing but note for 202A 4.5 says 1 extra so assuming 1 of the missing germs from 202A 4.0 is in 4.5
         number_germ = ifelse(transect_plot == "202A" & quadrat == 4.0, 15, number_germ),
         number_germ = ifelse(transect_plot == "202A" & quadrat == 4.5, 18, number_germ),
         number_germ = ifelse(transect_plot == "202D" & quadrat == 0.0, 19, number_germ)) %>%
  mutate(sum_check  = color_living_yellow + color_dead_yellow + missing_germs == number_germ)

WL2_WL3_mort_20210716 <- WL2_WL3_mort_20210716 %>%
  mutate(color_dead_yellow = ifelse(transect_plot == "201C" & quadrat == 1.0, 4, color_dead_yellow))

WL2_WL3_mort_20210702 <- WL2_WL3_mort_20210702 %>%
  mutate(color_dead_yellow = ifelse(transect_plot == "202C" & quadrat == 4.0, 2,  color_dead_yellow))

WL2_21_mort_list <- list(WL2_WL3_mort_20210618, WL2_WL3_mort_20210702, WL2_WL3_mort_20210716)

WL2_21_all_mort_germ1 <- WL2_21_mort_list %>%
  map(~ .x %>%
        filter(site == "WL2") %>%
        rename(mort_survey_date = survey_date,
               mort_notes = notes,
               num_dead = color_dead_yellow) %>%
        select(mort_survey_date, transect_plot, quadrat, num_dead, mort_notes)) %>%
  bind_rows() 

WL2_21_germ1_long_prep <- WL2_21_germ1_mort1_updated %>%
  select(site, survey_year, transect_plot, quadrat, germ_survey_date, pick_color, number_germ, missing_germs, germ_notes)

WL2_21_germ1_all_mort_long <- WL2_21_germ1_long_prep %>%
  left_join(WL2_21_all_mort_germ1)

WL2_21_germ1_day_since_germ <- WL2_21_germ1_all_mort_long %>%
  mutate(germ_survey_date = as.Date(germ_survey_date),
         mort_survey_date = as.Date(mort_survey_date),
         days_since_germ = as.integer(mort_survey_date - germ_survey_date)) %>%
  arrange(transect_plot, quadrat, mort_survey_date) %>%
  group_by(transect_plot, quadrat) %>%
  mutate(cum_dead = cumsum(num_dead)) %>%
  ungroup() %>%
  mutate(count_check = cum_dead + missing_germs <= number_germ)
```

#### Germ Cohort 2

```{r}
WL2_21_germ2 <- all_WL2_WL3_germ_2021 %>%
  filter(pick_color == "red", 
         site == "WL2") %>%  
  mutate(transect_plot = ifelse(row_number() == 37, "202D", transect_plot)) %>% #the number_germ and all other columns are the same for the two 201D 0.0 rows so I'm manually doing it here
  rename(germ_survey_date = survey_date,
         germ_notes = notes)

WL2_21_germ2_mort1 <- WL2_21_germ2 %>%
  left_join(WL2_WL3_mort_20210618[,c(4:7, 12:13, 18)] %>%
              rename(mort_survey_date = survey_date,
                     mort_notes = notes)) %>%
  mutate(sum_check  = color_living_red + color_dead_red == number_germ)

WL2_21_germ2_mort1_updated <- WL2_21_germ2_mort1 %>%
  select(-sum_check) %>%
  mutate(missing_germs = 0) %>%
  mutate(number_germ = ifelse(transect_plot == "201B" & quadrat == 0.0, 17, number_germ),
         number_germ = ifelse(transect_plot == "201B" & quadrat == 1.0, 10, number_germ),
         number_germ = ifelse(transect_plot == "201B" & quadrat == 4.0, 4, number_germ),
         number_germ = ifelse(transect_plot == "201C" & quadrat == 4.0, 6, number_germ),
         number_germ = ifelse(transect_plot == "201D" & quadrat == 1.0, 4, number_germ), 
         number_germ = ifelse(transect_plot == "202C" & quadrat == 3.0, 4, number_germ)) %>%
  mutate(sum_check  = color_living_red + color_dead_red + missing_germs == number_germ)

WL2_21_mort_list <- list(WL2_WL3_mort_20210618, WL2_WL3_mort_20210702, WL2_WL3_mort_20210716)

WL2_21_all_mort_germ2 <- WL2_21_mort_list %>%
  map(~ .x %>%
        filter(site == "WL2") %>%
        rename(mort_survey_date = survey_date,
               mort_notes = notes,
               num_dead = color_dead_red) %>%
        select(mort_survey_date, transect_plot, quadrat, num_dead, mort_notes)) %>%
  bind_rows() 

WL2_21_germ2_long_prep <- WL2_21_germ2_mort1_updated %>%
  select(site, survey_year, transect_plot, quadrat, germ_survey_date, pick_color, number_germ, missing_germs, germ_notes)

WL2_21_germ2_all_mort_long <- WL2_21_germ2_long_prep %>%
  left_join(WL2_21_all_mort_germ2)

WL2_21_germ2_day_since_germ <- WL2_21_germ2_all_mort_long %>%
  mutate(germ_survey_date = as.Date(germ_survey_date),
         mort_survey_date = as.Date(mort_survey_date),
         days_since_germ = as.integer(mort_survey_date - germ_survey_date)) %>%
  arrange(transect_plot, quadrat, mort_survey_date) %>%
  group_by(transect_plot, quadrat) %>%
  mutate(cum_dead = cumsum(num_dead)) %>%
  ungroup() %>%
  mutate(count_check = cum_dead + missing_germs <= number_germ)
```

#### Combine All Germ Cohorts and Save CSV

```{r}
WL2_21_all_cohort_list <- list(WL2_21_germ0_day_since_germ, WL2_21_germ1_day_since_germ, WL2_21_germ2_day_since_germ)
  
WL2_21_all_germ_mort <- bind_rows(WL2_21_all_cohort_list)
#write.csv(WL2_21_all_germ_mort, "../output/WL2_21_all_germ_mort", row.names = FALSE)
```

### WL3 2021 Germ and Mortality - need to check to see what cohort orange is - flowering? (surveyed on the same day as blue)

#### Germ Cohort 0

```{r}
WL3_21_germ0 <- all_WL2_WL3_germ_2021 %>%
  filter(pick_color == "blue", #cohort 0 has true leaves
         site == "WL3") %>%  
  rename(germ_survey_date = survey_date,
         germ_notes = notes) 

WL3_21_germ0_mort1 <- WL3_21_germ0 %>%
  left_join(WL2_WL3_mort_20210618[,c(4:9,18)] %>%
              rename(mort_survey_date = survey_date,
                     mort_notes = notes)) %>%
  mutate(sum_check  = color_living_blue + color_dead_blue == number_germ)

WL3_21_germ0_mort1_updated <- WL3_21_germ0_mort1 %>%
  select(-sum_check) %>%
  mutate(missing_germs = 0) %>%
  mutate(number_germ = ifelse(transect_plot == 2 & quadrat == 0.0, 12, number_germ),
         number_germ = ifelse(transect_plot == 3 & quadrat == 0.0, 18, number_germ),
         missing_germs = ifelse(transect_plot == 3 & quadrat == 0.5, 1, missing_germs),
         number_germ = ifelse(transect_plot == 3 & quadrat == 1.5, 14, number_germ),
         missing_germs = ifelse(transect_plot == 3 & quadrat == 3.0, 1, missing_germs),
         missing_germs = ifelse(transect_plot == 3 & quadrat == 3.5, 3, missing_germs),
         number_germ = ifelse(transect_plot == 3 & quadrat == 4.5, 7, number_germ),
         missing_germs = ifelse(transect_plot == 4 & quadrat == 7.5, 1, missing_germs)) %>%
  mutate(sum_check  = color_living_blue + color_dead_blue + missing_germs == number_germ)

WL3_21_mort_list <- list(WL2_WL3_mort_20210618, WL2_WL3_mort_20210702, WL2_WL3_mort_20210716)

WL3_21_all_mort_germ0 <- WL3_21_mort_list %>%
  map(~ .x %>%
        filter(site == "WL3") %>%
        rename(mort_survey_date = survey_date,
               mort_notes = notes,
               num_dead = color_dead_blue) %>%
        select(mort_survey_date, transect_plot, quadrat, num_dead, mort_notes)) %>%
  bind_rows() 

WL3_21_germ0_long_prep <- WL3_21_germ0_mort1_updated %>%
  select(site, survey_year, transect_plot, quadrat, germ_survey_date, pick_color, number_germ, missing_germs, germ_notes)

WL3_21_germ0_all_mort_long <- WL3_21_germ0_long_prep %>%
  left_join(WL3_21_all_mort_germ0)

WL3_21_germ0_day_since_germ <- WL3_21_germ0_all_mort_long %>%
  mutate(germ_survey_date = as.Date(germ_survey_date),
         mort_survey_date = as.Date(mort_survey_date),
         days_since_germ = as.integer(mort_survey_date - germ_survey_date)) %>%
  arrange(transect_plot, quadrat, mort_survey_date) %>%
  group_by(transect_plot, quadrat) %>%
  mutate(cum_dead = cumsum(num_dead)) %>%
  ungroup() %>%
  mutate(count_check = cum_dead + missing_germs <= number_germ)
```

#### Germ Cohort 1

```{r}
WL3_21_germ1 <- all_WL2_WL3_germ_2021 %>%
  filter(pick_color == "yellow", 
         site == "WL3") %>%  
  rename(germ_survey_date = survey_date,
         germ_notes = notes) 

WL3_21_germ1_mort1 <- WL3_21_germ1 %>%
  left_join(WL2_WL3_mort_20210618[,c(4:7, 10:11,18)] %>%
              rename(mort_survey_date = survey_date,
                     mort_notes = notes)) %>%
  mutate(sum_check  = color_living_yellow + color_dead_yellow == number_germ)

WL3_21_germ1_mort1_updated <- WL3_21_germ1_mort1 %>%
  select(-sum_check) %>%
  mutate(missing_germs = 0) %>%
  mutate(number_germ = ifelse(transect_plot == 1 & quadrat == 0.0, 8, number_germ),
         missing_germs = ifelse(transect_plot == 1 & quadrat == 1.0, 1, missing_germs),
         number_germ = ifelse(transect_plot == 1 & quadrat == 5.0, 1, number_germ),
         number_germ = ifelse(transect_plot == 2 & quadrat == 0.0, 16, number_germ),
         number_germ = ifelse(transect_plot == 2 & quadrat == 1.0, 79, number_germ),
         number_germ = ifelse(transect_plot == 2 & quadrat == 1.5, 10, number_germ),
         missing_germs = ifelse(transect_plot == 2 & quadrat == 2.5, 1, missing_germs),
         number_germ = ifelse(transect_plot == 2 & quadrat == 3.5, 2, number_germ),
         missing_germs = ifelse(transect_plot == 3 & quadrat == 0.0, 1, missing_germs),
         missing_germs = ifelse(transect_plot == 3 & quadrat == 4.5, 1, missing_germs),
         missing_germs = ifelse(transect_plot == 3 & quadrat == 5.5, 1, missing_germs),
         number_germ = ifelse(transect_plot == 3 & quadrat == 6.0, 4, number_germ),
         number_germ = ifelse(transect_plot == 3 & quadrat == 6.5, 4, number_germ),
         missing_germs = ifelse(transect_plot == 4 & quadrat == 2.0, 2, missing_germs),
         number_germ = ifelse(transect_plot == 4 & quadrat == 5.5, 15, number_germ)) %>%
  mutate(sum_check  = color_living_yellow + color_dead_yellow + missing_germs == number_germ)

WL3_21_all_mort_germ1 <- WL3_21_mort_list %>%
  map(~ .x %>%
        filter(site == "WL3") %>%
        rename(mort_survey_date = survey_date,
               mort_notes = notes,
               num_dead = color_dead_yellow) %>%
        select(mort_survey_date, transect_plot, quadrat, num_dead, mort_notes)) %>%
  bind_rows() 

WL3_21_germ1_long_prep <- WL3_21_germ1_mort1_updated %>%
  select(site, survey_year, transect_plot, quadrat, germ_survey_date, pick_color, number_germ, missing_germs, germ_notes)

WL3_21_germ1_all_mort_long <- WL3_21_germ1_long_prep %>%
  left_join(WL3_21_all_mort_germ1)

WL3_21_germ1_day_since_germ <- WL3_21_germ1_all_mort_long %>%
  mutate(germ_survey_date = as.Date(germ_survey_date),
         mort_survey_date = as.Date(mort_survey_date),
         days_since_germ = as.integer(mort_survey_date - germ_survey_date)) %>%
  arrange(transect_plot, quadrat, mort_survey_date) %>%
  group_by(transect_plot, quadrat) %>%
  mutate(cum_dead = cumsum(num_dead)) %>%
  ungroup() %>%
  mutate(count_check = cum_dead + missing_germs <= number_germ)
```

#### Germ Cohort 2

```{r}
WL3_21_germ2 <- all_WL2_WL3_germ_2021 %>%
  filter(pick_color == "red", 
         site == "WL3") %>%  
  rename(germ_survey_date = survey_date,
         germ_notes = notes) 

WL3_21_germ2_mort1 <- WL3_21_germ2 %>%
  left_join(WL2_WL3_mort_20210618[,c(4:7, 12:13,18)] %>%
              rename(mort_survey_date = survey_date,
                     mort_notes = notes)) %>%
  mutate(sum_check  = color_living_red + color_dead_red == number_germ)

WL3_21_germ2_mort1_updated <- WL3_21_germ2_mort1 %>%
  select(-sum_check) %>%
  mutate(missing_germs = 0) %>%
  mutate(missing_germs = ifelse(transect_plot == 3 & quadrat == 5.0, 1, missing_germs)) %>%
  mutate(sum_check  = color_living_red + color_dead_red + missing_germs == number_germ)

WL3_21_all_mort_germ2 <- WL3_21_mort_list %>%
  map(~ .x %>%
        filter(site == "WL3") %>%
        rename(mort_survey_date = survey_date,
               mort_notes = notes,
               num_dead = color_dead_red) %>%
        select(mort_survey_date, transect_plot, quadrat, num_dead, mort_notes)) %>%
  bind_rows() 

WL3_21_germ2_long_prep <- WL3_21_germ2_mort1_updated %>%
  select(site, survey_year, transect_plot, quadrat, germ_survey_date, pick_color, number_germ, missing_germs, germ_notes)

WL3_21_germ2_all_mort_long <- WL3_21_germ2_long_prep %>%
  left_join(WL3_21_all_mort_germ2)

WL3_21_germ2_day_since_germ <- WL3_21_germ2_all_mort_long %>%
  mutate(germ_survey_date = as.Date(germ_survey_date),
         mort_survey_date = as.Date(mort_survey_date),
         days_since_germ = as.integer(mort_survey_date - germ_survey_date)) %>%
  arrange(transect_plot, quadrat, mort_survey_date) %>%
  group_by(transect_plot, quadrat) %>%
  mutate(cum_dead = cumsum(num_dead)) %>%
  ungroup() %>%
  mutate(count_check = cum_dead + missing_germs <= number_germ)
```

#### Combine All Germ Cohorts and Save CSV

```{r}
WL3_21_all_cohort_list <- list(WL3_21_germ0_day_since_germ, WL3_21_germ1_day_since_germ, WL3_21_germ2_day_since_germ)
  
WL3_21_all_germ_mort <- bind_rows(WL3_21_all_cohort_list)
write.csv(WL3_21_all_germ_mort, "../output/WL3_21_all_germ_mort", row.names = FALSE)
```

### WL1 2017-2018 Germination - this data is quite messy, there is no pick color and mort surveys are included in germ survey

```{r}
WL1_fall_2017_germ <- WL_germ_20171102_20171118 %>%
  filter(site == "WL1") %>%
  mutate(date = ymd(date)) %>%
  rename(germ_survey_date = date,
         transect_plot = 'plot/ transect',
         quadrat = quad,
         number_germ = '#germ',
         germ_notes = notes) %>%
  filter(germ_survey_date == "2017-11-18") %>%
  mutate(pick_color = "blue")

WL1_germ_20180103 <- WL1_CP2_germ_20180103 %>%
  filter(site == "WL1") %>%
  mutate(date = ymd(date)) %>%
  rename(germ_survey_date = date,
         transect_plot = 'plot/transect',
         quadrat = quad,
         number_germ = '#new_germ',
         germ_notes = notes) %>%
  mutate(pick_color = "yellow") %>%
  select(-'#blue_total') %>%
  select(-'#blue_dead') %>%
  select(-'#blue_misid')


#difficult to process later germ cohorts - probably need to look at physical datasheets

  
  
#Blue, yellow, red, green
WL_germ_20171102_20171118
WL1_CP2_germ_20180103
WL1_germ_20180522
WL1_germ_20180603
WL1_WL2_WL3_CP2_germ_20180613_20180604_20180608
WL1_WL2_WL3_CP2_germ_20180622
```

### WL2 2017-2018 Germination

```{r}
WL2_germ1_20180603 <- WL1_WL2_WL3_CP2_germ_20180613_20180604_20180608 %>%
  filter(site == "WL2") %>%
  mutate(date = ymd(date)) %>%
  rename(germ_survey_date = date,
         transect_plot = 'plot/transect',
         quadrat = quad,
         number_germ = '# old germ',
         germ_notes = notes) %>%
  mutate(pick_color = "orange") %>%
  select(-'# new germ') %>%
  select(-'# mortality')

WL2_germ2_20180603 <- WL1_WL2_WL3_CP2_germ_20180613_20180604_20180608 %>%
  filter(site == "WL2") %>%
  mutate(date = ymd(date)) %>%
  rename(germ_survey_date = date,
         transect_plot = 'plot/transect',
         quadrat = quad,
         number_germ = '# new germ',
         germ_notes = notes) %>%
  mutate(pick_color = "green") %>%
  select(-'# old germ') %>%
  select(-'# mortality')

WL2_germ_20180622 <- WL1_WL2_WL3_CP2_germ_20180622 %>%
  filter(site == "WL2") %>%
  mutate(date = ymd(date)) %>%
  rename(germ_survey_date = date,
         transect_plot = 'plot/transect',
         quadrat = quad,
         number_germ = '# germ',
         germ_notes = notes) %>%
  mutate(pick_color = "yellow") %>%
  select(-budding) %>%
  select(-'flowering (white pick)')

#Combine all germ surveys
WL2_18_all_germ_list <- list(WL2_germ1_20180603, WL2_germ2_20180603, WL2_germ_20180622)
  
all_WL2_germ_2018 <- bind_rows(WL2_18_all_germ_list)
#write.csv(all_WL2_germ_2018, "../output/all_WL2_germ_2018", row.names = FALSE)


#the mortality surveys
WL2_germ_20180603
WL2_WL3_mort_20180720
```

### WL3 2017-2018 Germination

```{r}
WL3_germ1_20180603 <- WL1_WL2_WL3_CP2_germ_20180613_20180604_20180608 %>%
  filter(site == "WL3") %>%
  mutate(date = ymd(date)) %>%
  rename(germ_survey_date = date,
         transect_plot = 'plot/transect',
         quadrat = quad,
         number_germ = '# old germ',
         germ_notes = notes) %>%
  mutate(pick_color = "orange") %>%
  select(-'# new germ') %>%
  select(-'# mortality')

WL3_germ2_20180603 <- WL1_WL2_WL3_CP2_germ_20180613_20180604_20180608 %>%
  filter(site == "WL3") %>%
  mutate(date = ymd(date)) %>%
  rename(germ_survey_date = date,
         transect_plot = 'plot/transect',
         quadrat = quad,
         number_germ = '# new germ',
         germ_notes = notes) %>%
  mutate(pick_color = "green") %>%
  select(-'# old germ') %>%
  select(-'# mortality')

WL3_germ_20180622 <- WL1_WL2_WL3_CP2_germ_20180622 %>%
  filter(site == "WL3") %>%
  mutate(date = ymd(date)) %>%
  rename(germ_survey_date = date,
         transect_plot = 'plot/transect',
         quadrat = quad,
         number_germ = '# germ',
         germ_notes = notes) %>%
  mutate(pick_color = "yellow") %>%
  select(-budding) %>%
  select(-'flowering (white pick)')

WL3_germ_20180603 #this says pink and green picks but look up table doesn't have pink
```

### CP2 2017-2018 Germiantion

```{r}
CP2_germ_20180608 <- WL1_WL2_WL3_CP2_germ_20180613_20180604_20180608 %>%
  filter(site == "CP2") %>%
  mutate(date = ymd(date)) %>%
  rename(germ_survey_date = date,
         transect_plot = 'plot/transect',
         quadrat = quad,
         number_germ = '# old germ',
         germ_notes = notes) %>%
  mutate(pick_color = "green") %>%
  select(-'# new germ') %>%
  select(-'# mortality')

CP2_germ_20180622 <- WL1_WL2_WL3_CP2_germ_20180622 %>%
  filter(site == "CP2") %>%
  mutate(date = ymd(date)) %>%
  rename(germ_survey_date = date,
         transect_plot = 'plot/transect',
         quadrat = quad,
         number_germ = '# germ',
         germ_notes = notes) %>%
  mutate(pick_color = "yellow") %>%
  select(-budding) %>%
  select(-'flowering (white pick)')

#Combine all germ surveys
CP2_18_all_germ_list <- list(CP2_germ_20180608, CP2_germ_20180622)
  
all_CP2_germ_2018 <- bind_rows(CP2_18_all_germ_list)
#write.csv(all_CP2_germ_2018, "../output/all_CP2_germ_2018", row.names = FALSE)

```

### WL1 2018-2019 Germination

```{r}
WL1_germ_20190612 <- read_excel("../input/Germination and Mortality data/WL1_GermSurvey_6.12.19_edited9.4.19MM.xlsx")
WL1_germ_20190429 <- read_excel("../input/Germination and Mortality data/WL1_GermSurvey_4.29.19_edited9.3.19MM.xlsx")
```

### WL2, WL3, CP2 2018-2019 Germination

```{r}
WL2_WL3_germ_20190715 <- read_excel("../input/Germination and Mortality data/WL23_GermSurvey_7.15.19_edited9.4.19MM.xlsx")
WL2_WL3_CP2_germ_20190703 <- read_excel("../input/Germination and Mortality data/WL23+CP2_GermSurvey_7.3.19_edited9.3.19MM.xlsx")
```

### WL2 2016-2017 Germination - only 1 germinate

```{r}
WL2_germ_20161027 <- read_excel("../input/Germination and Mortality data/Demog_WL2_fall_2016_germ_survey_10.27_20161026_AC.xlsx", sheet = "Sheet1")
```
