---
title: "Parity and survival_all sites_all years"
author: "JR Gremer"
date: "2026-09-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(lubridate)
```

Code for standard error of the mean 
NOTE: will need to update this to get SE on logit scale...
```{r}
sem <- function(x, na.rm=FALSE) {           #for calculating standard error
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} 
```

Load data, created in Prep_merging_allpopsandyears.Rmd
```{r}
alldat = read_csv("../input/Cleaned/AllSites_cleaned.csv")
summary(alldat)
```
data are already filtered for only years in which band number and color are recorded (see Prep script)

Score for survival
```{r}
all_survival <- alldat %>%
  mutate(plant_id = paste(Site, Transect_plot, Quad, original_census_year, Band_color, Band_num, sep = "_"), #need to update plant_id with original census year
         alive = as.integer(!is.na(Phenology) & Phenology != "X")) %>%
  arrange(plant_id, Census_year) %>%
  group_by(plant_id) %>%
mutate(age = row_number(),
       next_year = Census_year + 1,
       alive_next_year = lead(alive),
         survived_to_next = as.integer(next_year == Census_year + 1 & alive == 1 & alive_next_year == 1)) %>%
  ungroup()
```
Update any pheno = NA to X after confirming no missing plants come back to life - still need to deal with the WL2_201C plant (see survival_allsites_allyears.Rmd for details)

```{r}
unique(all_survival$Phenology)

all_survival %>%
  filter(Phenology %in% c("*", "A", "XX", "S", NA))
#Note for row with * = *found plant in this spot that looks like a perennial, the BB is lost, we tagged it #58 and took data
#Note for row with A = not found, disturbed by small mammals/gophers
#Notes for rows with NA = rattlesnake on this transcet not surveying, band missing/missing band, not found, missing, plant and band not found, Plant and tag NRL, No data taken, 
#No note for XX, I think this should just be X
#Protocol says S should only be for TM2, but this individual with S was measured at YOSE1

summary(all_survival)

```

Code to get number of times a plant flowered  
first - get column of whether a plant flowered in a given year
then group by unique ID and get number of times they flowered (sum of of flowered for each individual)

Also add chunk to get at probability of surviving|flowering (iteroparity)?  