---
title: "Parity and survival_all sites_all years"
author: "JR Gremer"
date: "2026-09-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(lubridate)
```

Code for standard error of the mean 
NOTE: will need to update this to get SE on logit scale...
```{r}
sem <- function(x, na.rm=FALSE) {           #for calculating standard error
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} 
```

Load data, created in Prep_merging_allpopsandyears.Rmd
```{r}
alldat = read_csv("../input/Cleaned/AllSites_cleaned_survival.csv")
#elevation order
alldat$siteorder <- factor(alldat$Site, levels = c("TM2", "BH", "WL1", "WL2", "WL3", "YOSE1", "CP2", "YOSE7", "YOSE8"))


summary(alldat)
```
data are already filtered for only years in which band number and color are recorded (see Prep script and survival_allsites_allyears.Rmd)


Score whether plant flowers and reproduces
```{r}
unique(alldat$Phenology)
#Note for row with * = *found plant in this spot that looks like a perennial, the BB is lost, we tagged it #58 and took data
#Note for row with A = not found, disturbed by small mammals/gophers
#Notes for rows with NA = rattlesnake on this transcet not surveying, band missing/missing band, not found, missing, plant and band not found, Plant and tag NRL, No data taken, 
#No note for XX, I think this should just be X
#Protocol says S should only be for TM2, but this individual with S was measured at YOSE1

alldat = alldat %>%
         mutate(flowered = if_else(Phenology %in% c("P", "F", "B"), 1, 0)) %>%
         mutate(reproduced = if_else(Num_fruits >0, 1, 0))

table(alldat$Phenology, alldat$flowered) #looks good

table(alldat$reproduced, alldat$Num_fruits) #looks good, why is there a 4.4?? check that

table(alldat$Phenology, alldat$reproduced) #uh oh.  check why 4 vegetative plants have fruits, and 1 senescing plant has them too

```

Code to get number of times a plant flowered or reproduced

```{r}
reprotimes = alldat %>%
             filter(original_census_year < 2024) %>%
             group_by(plant_id, Site, siteorder, Transect_plot, Quad) %>%
             summarize(timesflower = sum(flowered, na.rm=T), timesrepro = sum(reproduced, na.rm=T))

ggplot(reprotimes, aes(x=siteorder, y= timesrepro)) +
  geom_boxplot() +
  labs(
    x = "Site",
    y = "Number of reproductive events"
  ) +
 # scale_color_brewer(palette = "Dark2") +
  theme_classic() 

#probably need to filter somehow or summarize.  Lots of plants don't repro at all.  
```

first - get column of whether a plant flowered in a given year
then group by unique ID and get number of times they flowered (sum of of flowered for each individual)

Also add chunk to get at probability of surviving|flowering (iteroparity)?  

Also add in chunk to get age at first reproduction.  That will be trickier. 