---
title: "Prelim_Cleaning"
output: html_document
date: "2025-04-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(readxl)
library(purrr)
library(lubridate)
library(stringr)
```

### Merge all annual censuses for each pop (but keep 2024 separate but able to merge in)

```{r}
#create functions that convert the colnames to the right structure type
convert_Census_date <- function(df) {
  df %>% mutate(Census_date = lubridate::ymd(as.character(Census_date)))
}

convert_Site <- function(df) {
  df %>% mutate(Site = as.character(Site))
}

convert_Transect_plot <- function(df) {
  df %>% mutate(Transect_plot = as.character(Transect_plot))
}

convert_Quad <- function(df) {
  df %>% mutate(Quad = as.character(Quad)) #if I want to combine TM2, this need to be as a character - its fine to leave as alphanumeric entries for now
}

convert_Phenology <- function(df) {
  df %>% mutate(Phenology = as.character(Phenology))
}

convert_Xcoord <- function(df) {
  df %>% mutate(Xcoord = as.numeric(Xcoord))
}

convert_Ycoord <- function(df) {
  df %>% mutate(Ycoord = as.numeric(Ycoord))
}

convert_Stem_diam_mm <- function(df) {
  df %>% mutate(Stem_diam_mm = as.character(Stem_diam_mm)) #want to be numeric eventually but need to deal with <1 and ~1 entries
}

convert_Height_cm <- function(df) {
  df %>% mutate(Height_cm = as.character(Height_cm)) #want to be numeric eventually but need to deal with <1 and ~1 entries
}

convert_Num_total_branches <- function(df) {
  df %>% mutate(Num_total_branches = as.numeric(Num_total_branches))
}

convert_Longest_leaf_cm <- function(df) {
  df %>% mutate(Longest_leaf_cm = as.character(Longest_leaf_cm)) #want to be numeric enventually but need to deal with <0.5 in CP2_23 - leave as character for now
}

convert_Num_flowers <- function(df) {
  df %>% mutate(Num_flowers = as.character(Num_flowers)) #want to be numeric but need to deal with character entries in BH_18 and check the different NA in BH_20
}

convert_Num_fruits <- function(df) {
  df %>% mutate(Num_fruits = as.character(Num_fruits)) #want to be numeric but need to deal with character entries
} 

convert_Longest_fruit_cm <- function(df) {
  df %>% mutate(Longest_fruit_cm = as.numeric(Longest_fruit_cm))
}

convert_Num_repro_branches <- function(df) {
  df %>% mutate(Num_repro_branches = as.numeric(Num_repro_branches))
}

convert_Notes <- function(df) {
  df %>% mutate(Notes = as.character(Notes))
}

convert_Census_year <- function(df) {
  df %>% mutate(Census_year = as.numeric(Census_year))
}

convert_Year_marked <- function(df) {
  df %>% mutate(Year_marked = as.numeric(Year_marked))
}

convert_Collected_by <-  function(df) {
  df %>% mutate(Collected_by = as.character(Collected_by))
}

convert_Entered_by <- function(df) {
  df %>% mutate(Entered_by = as.character(Entered_by))
}

convert_Date_entered <- function(df) {
  df %>% mutate(Date_entered = as.Date(Date_entered))
}

convert_Band <- function(df) {
  df %>% mutate(Band = as.character(Band))
}

convert_Band_num <- function(df) {
  df %>% mutate(Band_num = as.numeric(Band_num))
}

convert_Band_color <- function(df) {
  df %>% mutate(Band_color = as.character(Band_color))
}

convert_Pick_color <- function(df) {
  df %>% mutate(Pick_color = as.character(Pick_color))
}

convert_Sword_color <- function(df) {
  df %>% mutate(Sword_color = as.character(Sword_color))
}

convert_Tag <- function(df) {
  df %>% mutate(Tag = as.character(Tag))
}

convert_2018Tag <- function(df) {
  df %>% mutate(`2018tag` = as.character(`2018tag`))
}

convert_2019Tag <- function(df) {
  df %>% mutate(`2019tag` = as.character(`2019tag`))
}

convert_Buds <- function(df) {
  df %>% mutate(Buds = as.character(Buds))
}

convert_Herbivory <- function(df) {
  df %>% mutate(Herbivory = as.character(Herbivory))
}

convert_Category_herbivory <- function(df) {
  df %>% mutate(Category_herbivory = as.character(Category_herbivory))
}

convert_Leaf_thickness_mm <- function(df) {
  df %>% mutate(Leaf_thickness_mm = as.numeric(Leaf_thickness_mm))
}

convert_Num_leaves <- function(df) {
  df %>% mutate(Num_leaves = as.numeric(Num_leaves))
}

convert_Canopy_width_cm <- function(df) {
  df %>% mutate(Canopy_width_cm = as.numeric(Canopy_width_cm))
}

convert_all_columns <- function(df) {
  df %>%
    convert_Census_date() %>%
    convert_Site() %>%
    convert_Transect_plot() %>%
    convert_Quad() %>%
    convert_Phenology() %>%
    convert_Xcoord() %>%
    convert_Ycoord() %>%
    convert_Stem_diam_mm() %>%
    convert_Height_cm() %>%
    convert_Num_total_branches() %>%
    convert_Longest_leaf_cm() %>%
    convert_Num_flowers() %>%
    convert_Num_fruits() %>%
    convert_Longest_fruit_cm() %>%
    convert_Num_repro_branches() %>%
    convert_Notes() %>%
    convert_Census_year() %>%
    convert_Year_marked() %>%
    convert_Collected_by() %>%
    convert_Entered_by() %>%
    convert_Date_entered() %>%
    convert_Band() %>%
    convert_Band_num() %>%
    convert_Band_color() %>%
    convert_Pick_color() %>%
    convert_Sword_color() %>%
    convert_Tag() %>%
    convert_2018Tag() %>%
    convert_2019Tag() %>%
    convert_Buds() %>%
    convert_Herbivory() %>%
    convert_Category_herbivory() %>%
    convert_Leaf_thickness_mm() %>%
    convert_Num_leaves() %>%
    convert_Canopy_width_cm()
}
```

#### BH 2018-2023, abv 2024

```{r}
BH_18 <- read_excel("../input/Annual census data/Demog_BHR_Summer2018_8.12.2017_SLZ_cleaned.xlsx")

BH_19 <- read_excel("../input/Annual census data/BH_AnnualCensus_6.17.19_edited9.16.2019_EEB_cleaned.xlsx")

BH_20 <- read_excel("../input/Annual census data/BH_annualcensus_census 2020_finished_EC_cleaned.xlsx")

BH_21 <- read_excel("../input/Annual census data/BH_annualcensus_2021_verified_cleaned.xlsx")

BH_22 <- read_excel("../input/Annual census data/BH_annualcensus_20220613_verified_cleaned.xlsx")

BH_23 <- read_excel("../input/Annual census data/BH_annualcensus_2023_finished_cleaned.xlsx")

BH_24 <- read_excel("../input/Annual census data/BH_annualcensus_2024_verified_cleaned.xlsx")
```

##### Check colnames are the same for all BH dataframes and update those that aren't - DONE

```{r}
#make a df of the colnames from each dataset, which datasets have those names, and a count 
BH_col_check <- list(
  BH_18 = colnames(BH_18),
  BH_19 = colnames(BH_19),
  BH_20 = colnames(BH_20),
  BH_21 = colnames(BH_21),
  BH_22 = colnames(BH_22),
  BH_23 = colnames(BH_23),
  BH_24 = colnames(BH_24)
)
BH_all_col <- unique(unlist(BH_col_check))
BH_col_presence <- map_dfr(BH_all_col, function(col) {
  tibble(
    column = col,
    dataframes = names(BH_col_check)[map_lgl(BH_col_check, ~ col %in% .x)]
  )
}, .id = "col_id") %>%
  group_by(column) %>%
  summarise(dataframes = paste(dataframes, collapse = ", "),
            count = n()) %>%
  arrange(count)
print(BH_col_presence)
```

##### Update colnames for those that are different

```{r}
#update BH_19 and BH_23 to have matching colnames
BH_19 <- BH_19 %>%
  rename(Num_total_branches='Num_total_ branches')
BH_23 <- BH_23 %>%
  rename(Num_total_branches=Num_total_branch)
BH_22 <- BH_22 %>%
  rename(Phenology=Pheno)
```

##### Address individual rows preventing datasets from merging

```{r}
#make sure All NA's are recognized - verified these should be NA 
BH_22 <- BH_22 %>% #change "Na" to NA
  mutate(Band_num = na_if(Band_num, "Na")) %>%
  filter(!is.na(Census_year)) #there is 1 row that is all empty so filter out using Census_year because that is NA

BH_21 <- BH_21 %>%
  mutate(Band_num = na_if(Band_num, "Na"))

#update Quad to NA for rows in BH_20 where Transect_plot is "DP"
BH_20 <- BH_20 %>%
  mutate(Quad = ifelse(Quad == "DP", NA, Quad))
```

##### Move notes from Buds to Notes column - there must be a better way to do this but some rows in the Buds column have NA's and I think they would be copied or pasted over the existing notes in teh Notes column if I did this another way

```{r}
BH_20 <- BH_20 %>%
  mutate(Notes = if_else(Transect_plot == "2.0" & Quad == "3.5" & Band_color == "red" & Band_num == "29.0", Buds, Notes),
         Buds = ifelse(Transect_plot == "2.0" & Quad == "3.5" & Band_color == "red" & Band_num == "29.0", NA_character_, Buds),
         Notes = if_else(Transect_plot == "DP" & Band_color == "blue" & Band_num == "58.0", Buds, Notes),
         Buds = ifelse(Transect_plot == "DP" & Band_color == "blue" & Band_num == "58.0", NA_character_, Buds),
         Notes = if_else(Transect_plot == "2.0" & Quad == "4.5" & Band_color == "blue" & Band_num == "44.0", Buds, Notes),
         Buds = ifelse(Transect_plot == "2.0" & Quad == "4.5" & Band_color == "blue" & Band_num == "44.0", NA_character_, Buds),
         Notes = if_else(Transect_plot == "DP" & Band_color == "blue" & Band_num == "80.0", Buds, Notes),
         Buds = ifelse(Transect_plot == "DP" & Band_color == "blue" & Band_num == "80.0", NA_character_, Buds),
         Notes = if_else(Transect_plot == "2.0" & Quad == "3.0" & Band_color == "blue" & Band_num == "26.0", Buds, Notes),
         Buds = ifelse(Transect_plot == "2.0" & Quad == "3.0" & Band_color == "blue" & Band_num == "26.0", NA_character_, Buds),
         Notes = if_else(Transect_plot == "DP" & Band_color == "blue" & Band_num == "5.0", Buds, Notes),
         Buds = ifelse(Transect_plot == "DP" & Band_color == "blue" & Band_num == "5.0", NA_character_, Buds),
         Notes = if_else(Transect_plot == "2.0" & Quad == "1.5" & Band_color == "blue" & Band_num == "8.0", Buds, Notes),
         Buds = ifelse(Transect_plot == "2.0" & Quad == "1.5" & Band_color == "blue" & Band_num == "8.0", NA_character_, Buds))
  
```

##### Add a new column Stem_diam_mm_UFA that is numeric - convert \<1 and \~1 to numerics

```{r}
#make a list with all of the BH datasets
BH_list <- list(BH_18, BH_19, BH_20, BH_21, BH_22, BH_23, BH_24)

#Loop over each dataset in BH_list to add a new column called Stem_diam_mm_UFA where the "<1", "~1", and also numeric values <1 are changed to 0.5 and 1 so they can be used for analysis
for(i in seq_along(BH_list)) {
  BH_list[[i]] <- BH_list[[i]] %>%
    mutate(
      Stem_diam_mm = as.character(Stem_diam_mm),
      Stem_diam_mm_UFA = case_when(
      Stem_diam_mm == "<1" ~ 0.5,
      Stem_diam_mm < 1 ~ 0.5,
      Stem_diam_mm == "~1" ~ 1,
      TRUE ~ as.numeric(Stem_diam_mm)
    ))
} #there are some NA coersion warnings but verified no new NAs were introduced unexpectedly
```

##### Convert all "NA" strings to NA's

```{r}
#convert all strings of NA to be recognized as NA's
BH_list <- map(BH_list, ~.x %>%
                 mutate(across(where(is.character), ~na_if(., "NA"))))
```

##### Combine all BH dataframes and write csv

```{r}
#convert all of the columns and bind all of the BH datasets together
BH_all <- BH_list %>%
  lapply(convert_all_columns) %>%
  bind_rows() 

#save cleaned dataset as a csv
write.csv(BH_all, "../input/Cleaned/BH_cleaned_all.csv", row.names = FALSE)
```

##### Filter out 2024 and write csv - this is because 2024 was an abbreviated census

```{r}
#create csv for BH years 2018-2023
BH_18_23 <- BH_all %>%
  filter(Census_year != 2024)

write.csv(BH_18_23, "../input/Cleaned/BH_cleaned_18_23.csv", row.names = FALSE)

#create csv for BH 2024
BH_cleaned_24 <- BH_all %>%
  filter(Census_year == 2024)

write.csv(BH_cleaned_24, "../input/Cleaned/BH_cleaned_24.csv", row.names = FALSE)
```

#### CP2 2017-2023, abv 2024

```{r}
CP2_17 <- read_excel("../input/Annual census data/Demog_CP_17_20170920_SLZ_cleaned.xlsx")

CP2_18 <- read_excel("../input/Annual census data/CP2_AnnualCensus_7.30.2018_9.7.2018_SLZ_cleaned.xlsx")

CP2_19 <- read_excel("../input/Annual census data/CP2_FullCensus_8.7.19_edited9.13.19MM_cleaned.xlsx")

CP2_20 <- read_excel("../input/Annual census data/CP2_annualcensus_2020_finished_cleaned.xlsx")

CP2_21 <- read_excel("../input/Annual census data/CP2_annualcensus_2021_verified_cleaned.xlsx")

CP2_22 <- read_excel("../input/Annual census data/CP2_annualcensus_20220729_verified_YHW_cleaned.xlsx")

CP2_23 <- read_excel("../input/Annual census data/CP2_annualcensus_2023_finished_cleaned.xlsx")

CP2_24 <- read_excel("../input/Annual census data/CP2_annualcensus_2024_verified_cleaned.xlsx")
```

##### Check colnames are the same for all CP2 dataframes

```{r}
#make a df of the colnames from each dataset, which datasets have those names, and a count 
CP2_col_check <- list(
  CP2_17 = colnames(CP2_17),
  CP2_18 = colnames(CP2_18),
  CP2_19 = colnames(CP2_19),
  CP2_20 = colnames(CP2_20),
  CP2_21 = colnames(CP2_21),
  CP2_22 = colnames(CP2_22),
  CP2_23 = colnames(CP2_23),
  CP2_24 = colnames(CP2_24)
)
CP2_all_col <- unique(unlist(CP2_col_check))
CP2_col_presence <- map_dfr(CP2_all_col, function(col) {
  tibble(
    column = col,
    dataframes = names(CP2_col_check)[map_lgl(CP2_col_check, ~ col %in% .x)]
  )
}, .id = "col_id") %>%
  group_by(column) %>%
  summarise(dataframes = paste(dataframes, collapse = ", "),
            count = n()) %>%
  arrange(count)
print(CP2_col_presence)
```

##### Update colnames for those that are different

```{r}
#update CP2_19 and CP2_24 to have matching colnames
CP2_19 <- CP2_19 %>%
  rename(Height_cm = Height) %>%
  rename(Num_fruits = Num_fruit)

CP2_24 <- CP2_24 %>%
  rename(Num_repro_branches=Num_repro_branch)
```

##### Address individual rows/cells preventing datasets from merging

```{r}
#CP2_18 has notes Pink TP for some rows - need to look at paper datsheets
#CP2_17 <- CP2_17 %>%
 # mutate(Height_cm = case_when(
  #  Height_cm == "<1" ~ 0.5,
   # Height_cm == "~1" ~ 1,
   # TRUE ~ as.numeric(Height_cm)))

CP2_18 <- CP2_18 %>% #one entry has Num_total_branches that is "2*" with note that says "*one branch broken, used to be 3". Decided to use intact branch count
  mutate(
    Notes = ifelse(Num_total_branches == "2*" & !is.na(Notes) & Notes != "",
                   paste(Notes, "4/29/25 MMA updated Num_total_branches to intact count of 2"),
                   ifelse(Num_total_branches == "2*",
                          "4/29 MMA updated Num_total_branches to intact count of 2",
                          Notes)),
    Num_total_branches = ifelse(Num_total_branches == "2*", "2", Num_total_branches)
  )
  
CP2_23 <- CP2_23 %>%
  mutate(Census_date = lubridate::mdy(Census_date)) %>% #change Census_date from a character to a date 
  mutate(Date_entered = lubridate::mdy(Date_entered)) %>%
  mutate(Stem_diam_mm = ifelse(
    Stem_diam_mm == "2/64", "2.64", Stem_diam_mm
  )) #%>%
 # mutate(Height_cm = case_when(
   # Height_cm == "<1" ~ 0.5,
   # Height_cm == "~1" ~ 1,
   # TRUE ~ as.numeric(Height_cm)))

#CP2_24 <- CP2_24 %>%
  #mutate(Height_cm = case_when(
  #  Height_cm == "<1" ~ 0.5,
   # Height_cm == "~1" ~ 1,
   # TRUE ~ as.numeric(Height_cm)))
```

##### Add a new column Stem_diam_mm_UFA that is numeric - convert \<1 and \~1 to numerics

```{r}
#make a list with all of the BH datasets
CP2_list <- list(CP2_17, CP2_18, CP2_19, CP2_20, CP2_21, CP2_22, CP2_23, CP2_24)

#Loop over each dataset in CP2_list to add a new column called Stem_diam_mm_UFA where the "<1", "~1", and also numeric values <1 are changed to 0.5 and 1 so they can be used for analysis
for(i in seq_along(CP2_list)) {
  CP2_list[[i]] <- CP2_list[[i]] %>%
    mutate(
      Stem_diam_mm = as.character(Stem_diam_mm),
      Stem_diam_mm_UFA = case_when(
      Stem_diam_mm == "<1" ~ 0.5,
      Stem_diam_mm < 1 ~ 0.5,
      Stem_diam_mm == "~1" ~ 1,
      TRUE ~ as.numeric(Stem_diam_mm)
    ))
} #there are some NA coersion warnings but verified no new NAs were introduced unexpectedly
```

##### Convert all "NA" strings to NA's

```{r}
#convert all strings of NA to be recognized as NA's
CP2_list <- map(CP2_list, ~.x %>%
                 mutate(across(where(is.character), ~na_if(., "NA"))))
```

##### Combine all CP2 dataframes and write csv

```{r}
#convert all of the columns and bind all of the CP2 datasets together
CP2_all <- CP2_list %>%
  lapply(convert_all_columns) %>%
  bind_rows()

#save cleaned dataset as a csv
write.csv(CP2_all, "../input/Cleaned/CP2_cleaned_all.csv", row.names = FALSE)
```

##### Filter out 2024 and write csv - this is because 2024 was an abbreviated census

```{r}
#create csv for CP2 years 2017-2023
CP2_17_23 <- CP2_all %>%
  filter(Census_year != 2024)

write.csv(CP2_17_23, "../input/Cleaned/CP2_cleaned_17_23.csv", row.names = FALSE)

#create csv for CP2 2024
CP2_cleaned_24 <- CP2_all %>%
  filter(Census_year == 2024)

write.csv(CP2_cleaned_24, "../input/Cleaned/CP2_cleaned_24.csv", row.names = FALSE)
```

#### TM2 2016-2023, abv 2024

```{r}
TM2_16 <- read_excel("../input/Annual census data/Demog_TM2_spring_2016_foranalysis_cleaned.xlsx")

TM2_17 <- read_excel("../input/Annual census data/TM2_annualcensus_2017_jrg_fixederrors_foranalysis_cleaned.xlsx")

TM2_18 <- read_excel("../input/Annual census data/TM2_annualcensus_data_2018_foranalysis_cleaned.xlsx") #Check col names

TM2_19 <- read_excel("../input/Annual census data/TM2_AnnualCensus_2019_all_edited6.30.21JRG_cleaned.xlsx")

TM2_20 <- read_excel("../input/Annual census data/TM2_AnnualCensus_2020_foranalysis_cleaned.xlsx") #Check col names

TM2_21 <- read_excel("../input/Annual census data/TM2_annualcensus_2021_finished_cleaned.xlsx")

TM2_22 <- read_excel("../input/Annual census data/TM2_annual_census_2022_complete_YHW_Verified_cleaned.xlsx")

TM2_23 <- read_excel("../input/Annual census data/TM2_annualcensus_2023_finished_cleaned.xlsx")

TM2_24 <- read_excel("../input/Annual census data/TM2_annualcensus_2024_verified_cleaned.xlsx")
```

##### Check colnames are the same for all TM2 dataframes and update those that aren't

```{r}
TM2_col_check <- list(
  TM2_16 = colnames(TM2_16),
  TM2_17 = colnames(TM2_17),
  TM2_18 = colnames(TM2_18),
  TM2_19 = colnames(TM2_19),
  TM2_20 = colnames(TM2_20),
  TM2_21 = colnames(TM2_21),
  TM2_22 = colnames(TM2_22),
  TM2_23 = colnames(TM2_23),
  TM2_24 = colnames(TM2_24)
)
TM2_all_col <- unique(unlist(TM2_col_check))
TM2_col_presence <- map_dfr(TM2_all_col, function(col) {
  tibble(
    column = col,
    dataframes = names(TM2_col_check)[map_lgl(TM2_col_check, ~ col %in% .x)]
  )
}, .id = "col_id") %>%
  group_by(column) %>%
  summarise(dataframes = paste(dataframes, collapse = ", "),
            count = n()) %>%
  arrange(count)
print(TM2_col_presence)
```

##### Update colnames for those that are different

```{r}
#update TM2_18, TM2_19, and TM2_20 to have matching colnames

#Pick_color...25 and Sword_color...24 to Num_repro_branches and Num_total_branches but need numbers for those cols
TM2_18 <- TM2_18 %>%
  #TM2_18 has 2 Pick_color columns but Pick_color...13 is the column filled with colors, Pick_color...25 is filled with NAs
  rename(Pick_color=Pick_color...13) %>% 
  #TM2_28 has 2 Sword_color columns but Sword_color...14 is the column filled with colors, Sword_color...24 is filled with NAs 
  rename(Sword_color=Sword_color...14) %>%
  #Changing pick_color...25 and Sword_color...24 to Num_repro_branches and Num_total_branches - verified using original data set that these cols were added in the initial cleaning and should be filled with NAs
  rename(Num_repro_branches=Pick_color...25) %>%
  rename(Num_total_branches=Sword_color...24)
  
TM2_19 <- TM2_19 %>%
  rename(Leaf_thickness_mm=Leaf_thickness)

TM2_20 <- TM2_20 %>%
  #TM2_20 has 2 Category_herbivory columns both filled with NAs so I'm renaming Category_herbivory...23
  rename(Category_herbivory=Category_herbivory...23) %>% 
  rename(Num_total_branches=Total_branches) %>%
  #Changing Category_herbivory...31 to Leaf_thickness_mm - verified using origninal data set that these cols were added in the initial cleaning and should be filled with NAs
  rename(Leaf_thickness_mm=Category_herbivory...31)
```

##### Address individual rows/cells preventing datasets from merging

```{r}
TM2_17 <- TM2_17 %>% #change dead in Xcoord and Ycoord to NA
  mutate(Xcoord = na_if(Xcoord, "dead")) %>%
  mutate(Ycoord = na_if(Ycoord, "dead"))

TM2_18 <- TM2_18 %>%
  mutate( #Longest_fruit_cm has numbers followed by "LL" or "*"
    move_leaf = grepl("\\d+\\s*(LL|\\*)", Longest_fruit_cm), #identify cells with LL or * next to longest fruit length
    Longest_leaf_cm = ifelse( #move just the numbers to Longest_leaf_cm
      move_leaf,
      as.numeric(str_extract(Longest_fruit_cm, "\\d+(\\.\\d+)?")),
      Longest_leaf_cm
    ),
    Longest_fruit_cm = ifelse(move_leaf, NA, Longest_fruit_cm) #change Longest_fruit_cm to NA if there was a value added to Longest_fruit_cm
  ) %>%
  select(-move_leaf) %>% #Delete the move_leaf column
  mutate(Longest_fruit_cm = case_when( #Longest_fruit_cm has one row with "<0.5", change to 0.4
    Longest_fruit_cm == "<0.5" ~ 0.4, #NA coersion warnings but verified there are the expected number of NA's still
  )) 

TM2_22 <- TM2_22 %>%
  mutate(Band_num = case_when( #change Band_name from a string to a numeric value that has not been used before
    Band_num == "O4" ~ 999
  )) %>% #NA coersion warning but this is changing "NA" to NAs
 mutate(#Num_flowers was not entered so columns to the right were shifted: updating row where Num_repro_branches was "N"
    Num_flowers = ifelse(Num_repro_branches == "N", 0, Num_flowers),
    Num_fruits = ifelse(Num_repro_branches == "N", 2, Num_fruits),
    Longest_fruit_cm = ifelse(Num_repro_branches == "N", 3.6, Longest_fruit_cm),
    Herbivory = ifelse(Num_repro_branches == "N", "N", Herbivory),
    Num_repro_branches = ifelse(Num_repro_branches == "N", 1, Num_repro_branches),
  ) %>% 
  mutate(#Num_flowers was not entered so columns to the right were shifted: updating rows where Num_repro_branches was "Y"
    Num_flowers = case_when(
      (is.na(Band_num) & Stem_diam_mm == "0.94" & Height_cm == "9.1") | 
      Band_num %in% c("805", "814") ~ "0",
      TRUE ~ Num_flowers
    ),
    Num_fruits = case_when(
      is.na(Band_num) & Stem_diam_mm == "0.94" & Height_cm == "9.1" ~ "1",
      Band_num == "805" ~ "5",
      Band_num == "814" ~ "2",
      TRUE ~ Num_fruits
    ),
    Longest_fruit_cm = case_when(
      is.na(Band_num) & Stem_diam_mm == "0.94" & Height_cm == "9.1" ~ "2.4",
      Band_num == "805" ~ "4.8",
      Band_num == "814" ~ "5.0",
      TRUE ~ Longest_fruit_cm
    ),
    Num_repro_branches = case_when(
      (is.na(Band_num) & Stem_diam_mm == "0.94" & Height_cm == "9.1") |
      Band_num %in% c("805", "814") ~ "1",
      TRUE ~ Num_repro_branches
    ),
    Herbivory = case_when(
      (is.na(Band_num) & Stem_diam_mm == "0.94" & Height_cm == "9.1") |
      Band_num %in% c("805", "814") ~ "Y",
      TRUE ~ Herbivory
    )
  ) %>%
  mutate(#data from paper datasheet missing in original dataframe for 1 row, adding missing data here
    Num_flowers = case_when(
      Stem_diam_mm == "1.7" & Height_cm == "15.4" ~ "0",
      TRUE ~ Num_flowers
    ),
    Num_fruits = case_when(
      Stem_diam_mm == "1.7" & Height_cm == "15.4" ~ "11",
      TRUE ~ Num_fruits
    ),
    Longest_fruit_cm = case_when(
      Stem_diam_mm == "1.7" & Height_cm == "15.4" ~ "4.6",
      TRUE ~ Longest_fruit_cm
    ),
    Num_repro_branches = case_when(
      Stem_diam_mm == "1.7" & Height_cm == "15.4" ~ "1",
      TRUE ~ Num_repro_branches
    ),
    Herbivory = case_when(
      Stem_diam_mm == "1.7" & Height_cm == "15.4" ~ "Y",
      TRUE ~ Herbivory
    )
  )

#add site to Site col
TM2_22 <- TM2_22 %>%
  mutate(Site = "TM2")

#change site name from TM to TM2
TM2_19 <- TM2_19 %>%
  mutate(Site = "TM2")

#STILL NEED TO FIGURE OUT HOW TO CONVERT QUAD TO NUMERIC GIVEN THERE ARE CHARACTERS
```

##### Add a new column Stem_diam_mm_UFA that is numeric - convert \<1 and \~1 to numerics

```{r}
TM2_list <- list(TM2_16, TM2_17, TM2_18, TM2_19, TM2_20, TM2_21, TM2_22, TM2_23, TM2_24)

#Loop over each dataset in TM2_list to add a new column called Stem_diam_mm_UFA where the "<1", "~1", and also numeric values <1 are changed to 0.5 and 1 so they can be used for analysis
for(i in seq_along(TM2_list)) {
  TM2_list[[i]] <- TM2_list[[i]] %>%
    mutate(Stem_diam_mm_UFA = case_when(
      Stem_diam_mm == "<1" ~ 0.5,
      Stem_diam_mm < 1 ~ 0.5,
      Stem_diam_mm == "~1" ~ 1,
      TRUE ~ as.numeric(Stem_diam_mm)
    ))
} #there are some NA coersion warnings but verified no new NAs were introduced unexpectedly
```

##### Convert all "NA" strings to NA's

```{r}
TM2_list <- map(TM2_list, ~.x %>%
                 mutate(across(where(is.character), ~na_if(., "NA"))))
```

##### Combine all TM2 dataframes and write csv

```{r}
TM2_all <- TM2_list %>%
  lapply(convert_all_columns) %>%
  bind_rows()

write.csv(TM2_all, "../input/Cleaned/TM2_cleaned_all.csv", row.names = FALSE)
```

##### Filter out 2024 and write csv - this is because 2024 was an abbreviated census

```{r}
#create csv for TM2 years 2016-2023
TM2_16_23 <- TM2_all %>%
  filter(Census_year != 2024)

write.csv(TM2_16_23, "../input/Cleaned/TM2_cleaned_16_23.csv", row.names = FALSE)

#create csv for TM2 2024
TM2_cleaned_24 <- TM2_all %>%
  filter(Census_year == 2024)

write.csv(TM2_cleaned_24, "../input/Cleaned/TM2_cleaned_24.csv", row.names = FALSE)
```

#### WL1 2017-2023, abv 2024

```{r}
WL1_17 <- read_excel("../input/Annual census data/Demog_WL1_17_20170920_SLZ_cleaned.xlsx")

WL1_18 <- read_excel("../input/Annual census data/WL1_AnnualCensus_7.06.2018_MM_cleaned.xlsx")

WL1_19 <- read_excel("../input/Annual census data/WL1_FullCensus_8.7.19_edited9.13.19MM_cleaned.xlsx")

WL1_20 <- read_excel("../input/Annual census data/WL1_annualcensus_2020_finished_cleaned.xlsx") 

WL1_21 <- read_excel("../input/Annual census data/WL1_annualcensus_2021_verified_cleaned.xlsx")

WL1_22 <- read_excel("../input/Annual census data/WL1_annualcensus_20220708_YHW_verified_cleaned.xlsx")

WL1_23 <- read_excel("../input/Annual census data/WL1_annualcensus_2023_finished_cleaned.xlsx")

WL1_24 <- read_excel("../input/Annual census data/WL1_annualcensus_2024_verified_cleaned.xlsx")
```

##### Check colnames are the same for all WL1 dataframes

```{r}
WL1_col_check <- list(
  WL1_17 = colnames(WL1_17),
  WL1_18 = colnames(WL1_18),
  WL1_19 = colnames(WL1_19),
  WL1_20 = colnames(WL1_20),
  WL1_21 = colnames(WL1_21),
  WL1_22 = colnames(WL1_22),
  WL1_23 = colnames(WL1_23),
  WL1_24 = colnames(WL1_24)
)
WL1_all_col <- unique(unlist(WL1_col_check))
WL1_col_presence <- map_dfr(WL1_all_col, function(col) {
  tibble(
    column = col,
    dataframes = names(WL1_col_check)[map_lgl(WL1_col_check, ~ col %in% .x)]
  )
}, .id = "col_id") %>%
  group_by(column) %>%
  summarise(dataframes = paste(dataframes, collapse = ", "),
            count = n()) %>%
  arrange(count)
print(WL1_col_presence)
```

##### Update colnames for those that are different

```{r}
#update WL1_18, WL1_20, WL1_21, WL1_22, WL1_23, WL1_24
WL1_20 <- WL1_20 %>%
  rename(Band_num=Band_number)

WL1_18 <- WL1_18 %>%
  rename(Canopy_width_cm=Canopy_width)

WL1_22 <- WL1_22 %>%
  rename(Stem_diam_mm=Stem_Diam_mm) %>%
  rename(Xcoord=X_coord) %>%
  rename(Ycoord=Y_coord)

WL1_24 <- WL1_24 %>%
  rename(Year_marked='Year _marked') %>%
  rename(Phenology=Pheno)

WL1_23 <- WL1_23 %>%
  rename(Phenology=Pheno)

WL1_21 <- WL1_21 %>%
  rename(Xcoord=X_coord) %>%
  rename(Ycoord=Y_coord)
```

##### Address individual rows/cells preventing datasets from merging

```{r}
WL1_23 <- WL1_23 %>%
  mutate(Date_entered = ifelse(Date_entered == "45315.0", as.character("01/24/2024"), Date_entered)) %>% #confirmed with original datasheet that there were exactly 60 rows where Date_entered was 01/24/2024 which matches the number of rows where Date_entered was 45315.0
  mutate(Date_entered = ifelse(Date_entered == "45322.0", as.character("01/31/2024"), Date_entered)) %>% #confirmed with original datasheet that there were exactly 97 rows where Date_entered was 01/31/2024 which matches the number of rows where Date_entered was 45322.0
  mutate(Date_entered = lubridate::mdy(Date_entered)) #change col from character to date

#add site info
WL1_22 <- WL1_22 %>%
  mutate(Site = "WL1") %>%
  mutate(Census_date = mdy("07/08/2022"))
```

##### Add a new column Stem_diam_mm_UFA that is numeric - convert \<1 and \~1 to numerics

```{r}
WL1_list <- list(WL1_17, WL1_18, WL1_19, WL1_20, WL1_21, WL1_22, WL1_23, WL1_24)

#Loop over each dataset in WL1_list to add a new column called Stem_diam_mm_UFA where the "<1", "~1", and also numeric values <1 are changed to 0.5 and 1 so they can be used for analysis
for(i in seq_along(WL1_list)) {
  WL1_list[[i]] <- WL1_list[[i]] %>%
    mutate(Stem_diam_mm_UFA = case_when(
      Stem_diam_mm == "<1" ~ 0.5,
      Stem_diam_mm < 1 ~ 0.5,
      Stem_diam_mm == "~1" ~ 1,
      TRUE ~ as.numeric(Stem_diam_mm)
    ))
} #there are some NA coersion warnings but verified no new NAs were introduced unexpectedly
```

##### Convert all "NA" strings to NA's

```{r}
WL1_list <- map(WL1_list, ~.x %>%
                 mutate(across(where(is.character), ~na_if(., "NA"))))
```

##### Combine all WL1 dataframes and write csv

```{r}
WL1_all <- WL1_list %>%
  lapply(convert_all_columns) %>%
  bind_rows()

write.csv(WL1_all, "../input/Cleaned/WL1_cleaned.csv", row.names = FALSE)
```

##### Filter out 2024 and write csv - this is because 2024 was an abbreviated census

```{r}
#create csv for WL1 years 2017-2023 
WL1_17_23 <- WL1_all %>%   filter(Census_year != 2024)  

write.csv(WL1_17_23, "../input/Cleaned/WL1_cleaned_17_23.csv", row.names = FALSE) 

#create csv for WL1 2024 
WL1_cleaned_24 <- WL1_all %>%   filter(Census_year == 2024)  

write.csv(WL1_cleaned_24, "../input/Cleaned/WL1_cleaned_24.csv", row.names = FALSE)
```

#### WL2 2016-2025

```{r}
WL2_16 <- read_excel("../input/Annual census data/Demog_WL2_summer_2016_20170622_AJC_cleaned.xlsx")

WL2_17 <- read_excel("../input/Annual census data/Demog_WL2_17_20170920_SLZ_cleaned.xlsx")

WL2_18 <- read_excel("../input/Annual census data/WL2_AnnualCensus_7.25.18_20180819_SLZ_cleaned.xlsx")

WL2_19 <- read_excel("../input/Annual census data/WL2_FullCensus_8.7.19_edited9.13.19MM_cleaned.xlsx")

WL2_20 <- read_excel("../input/Annual census data/WL2_annualcensus_2020_finished_cleaned.xlsx") #Check col names

WL2_21 <- read_excel("../input/Annual census data/WL2_annualcensus_2021_verified_cleaned.xlsx")

WL2_22 <- read_excel("../input/Annual census data/WL2_Annual_Census_2022_Verified_cleaned.xlsx")

WL2_23 <- read_excel("../input/Annual census data/WL2_annualcensus_2023_finished_cleaned.xlsx")

WL2_24 <- read_excel("../input/Annual census data/WL2_annualcensus_2024_verified_cleaned.xlsx")

WL2_25 <- read_csv("../input/Annual census data/WL2_annualcensus_20250731.csv")
```

##### Check colnames are the same for all WL2 dataframes

```{r}
WL2_col_check <- list(
  WL2_16 = colnames(WL2_16),
  WL2_17 = colnames(WL2_17),
  WL2_18 = colnames(WL2_18),
  WL2_19 = colnames(WL2_19),
  WL2_20 = colnames(WL2_20),
  WL2_21 = colnames(WL2_21),
  WL2_22 = colnames(WL2_22),
  WL2_23 = colnames(WL2_23),
  WL2_24 = colnames(WL2_24),
  WL2_25 = colnames(WL2_25)
)
WL2_all_col <- unique(unlist(WL2_col_check))
WL2_col_presence <- map_dfr(WL2_all_col, function(col) {
  tibble(
    column = col,
    dataframes = names(WL2_col_check)[map_lgl(WL2_col_check, ~ col %in% .x)]
  )
}, .id = "col_id") %>%
  group_by(column) %>%
  summarise(dataframes = paste(dataframes, collapse = ", "),
            count = n()) %>%
  arrange(count)
print(WL2_col_presence)
```

##### Update colnames for those that are different

```{r}
#update WL2_16, WL2_20, WL2_22
WL2_16 <- WL2_16 %>%
  rename(Height_cm=Plant_height_cm)

WL2_20 <- WL2_20 %>%
  rename(Band_num=Band_number) %>%
  rename(Num_leaves=Num_leaves...34) %>% #WL2_20 has 2 Num_leaves columns, both filled with NA so I will rename Num_leaves...34 
  select(-Num_leaves...35)

WL2_22 <- WL2_22 %>%
  rename(Phenology=Pheno)
```

##### Address individual rows/cells preventing datasets from merging

```{r}
#WL2_18 204D 1.0 23 34 Stem_diam_mm = HA Height = 2cm
#2641     8 a double           2 cm                              C:/Users/marik/R/Demography_Analyses/i…

#WL2_20 201B 3.5 15 40 <1 1.5 Num_fruits = A
#4206    13 a double           A                                 C:/Users/marik/R/Demography_Analyses/i…

#WL2_20 201D 1.5 <1 Height= 3.`
#4336     8 a double           3.`                               C:/Users/marik/R/Demography_Analyses/i…

#WL2_21 201C 0 Band_num = 60 Height_cm= >15
#4613     8 a double           >15                               C:/Users/marik/R/Demography_Analyses/i…

#this is the column for buds
#4663    24 1/0/T/F/TRUE/FALSE 2 extra blues; added at end       C:/Users/marik/R/Demography_Analyses/i…
#4666    24 1/0/T/F/TRUE/FALSE 1 fruit dehisced;dying; dried out C:/Users/marik/R/Demography_Analyses/i…
#4669    24 1/0/T/F/TRUE/FALSE dying; dried out                  C:/Users/marik/R/Demography_Analyses/i…

# WL2_22 201C 2 Height = .1.2
#5115     8 a double           .1.2                              C:/Users/marik/R/Demography_Analyses/i…

WL2_16 <- WL2_16 %>%
  mutate(Phenology = case_when(
    Phenology == "5.0" ~ "S",
    TRUE ~ as.character(Phenology)
  ))

WL2_17 <- WL2_17 %>%
  mutate(Phenology = case_when(
    Phenology == "none" ~ NA_character_,
    TRUE ~ as.character(Phenology)
  )) %>%
  mutate(Phenology = case_when(
    Phenology == "*F" ~ "F",
    TRUE ~ as.character(Phenology)
  )) %>%
   mutate(
    Notes = ifelse(Num_total_branches == "2*",
                   paste(Notes, "5/9/25 MMA updated Num_total_branches to intact count of 2"), Notes),
    Num_total_branches = ifelse(Num_total_branches == "2*", "2", Num_total_branches)
  ) %>%
  mutate(
    Notes = ifelse(
      Xcoord == "X",
      paste(Notes, "5/9/2025 MMA updated Xcoord to NA and Phenology to \"X\"", sep = " | "),
      Notes
    ),
    Phenology = ifelse(Xcoord == "X", "X", Phenology),
    Xcoord = ifelse(Xcoord == "X", NA, Xcoord)
  )

WL2_18 <- WL2_18 %>%
  mutate(Longest_fruit_cm = ifelse(Longest_fruit_cm == "3 cm", as.numeric(3.0), Longest_fruit_cm)) %>% #change row where Longest_fruit_cm is "3 cm"
  mutate(
    Notes = if_else(
      Num_repro_branches == "1?",
      paste(Notes, "5/9/2025 MMA changed Num_repro_branches 1 from 1?", sep = " | "),
      Notes
    )
  ) %>%
  mutate(Num_repro_branches = ifelse(Num_repro_branches == "1?", 1, Num_repro_branches))%>%
  filter(Phenology != "NO PLANTS") %>%
  mutate(Phenology = case_when(
    Phenology == "D" ~ "X",
    TRUE ~ as.character(Phenology)
  )) %>%
  mutate(Phenology = case_when(
    Phenology == "PF" ~ "P",
    TRUE ~ as.character(Phenology)
  )) %>%
  mutate(Height_cm = ifelse(Height_cm == "2 cm", 2, Height_cm)) %>%
  mutate(Stem_diam_mm = ifelse(Stem_diam_mm == "HA", "NA", Stem_diam_mm))

WL2_20 <- WL2_20 %>%
  mutate(Longest_fruit_cm = ifelse(Longest_fruit_cm == "2,6", "2.6", Longest_fruit_cm)) %>%
  mutate(Num_fruits = ifelse(Num_fruits == "A", "NA", Num_fruits)) %>%
  mutate(Height_cm = ifelse(Height_cm == "3.`", "3.1", Height_cm))

WL2_21 <- WL2_21 %>%
  mutate(
    Notes = ifelse(Height_cm == ">15",
                   paste(Notes, "5/12/25 MMA updated Height_cm from >15 to 15 so col is numeric"), Notes),
    Height_cm = ifelse(Height_cm == ">15", "15", Height_cm)
  ) %>%
  mutate(
  Notes = ifelse(
    Buds %in% c(
      "2 extra blues; added at end",
      "1 fruit dehisced;dying; dried out",
      "dying; dried out"),
    paste0("5/12/25 MMA moved from Num_buds: ", Buds),
    Notes),
  Buds = ifelse(
    Buds %in% c(
      "2 extra blues; added at end",
      "1 fruit dehisced;dying; dried out",
      "dying; dried out"), NA, Buds))

WL2_22 <- WL2_22 %>%
  mutate(Height_cm = ifelse(Height_cm == ".1.2", "1.2", Height_cm)) %>%
  mutate(Site = "WL2")

WL2_23 <- WL2_23 %>%
 mutate(Longest_fruit_cm = ifelse(Longest_fruit_cm == "3.83.9", "3.8", Longest_fruit_cm)) %>%
  mutate(Longest_fruit_cm = ifelse(Band_num == "513", "3.9", Longest_fruit_cm),) %>%
   mutate(Longest_fruit_cm = ifelse(Band_num == "51", "0.5", Longest_fruit_cm),) %>%
   mutate(Longest_fruit_cm = ifelse(Band_num == "848", "0.6", Longest_fruit_cm),) %>%
   mutate(Longest_fruit_cm = ifelse(Band_num == "164", "0.7", Longest_fruit_cm),) 
```

##### Add a new column Stem_diam_mm_UFA that is numeric - convert \<1 and \~1 to numerics

```{r}
WL2_list <- list(WL2_16, WL2_17, WL2_18, WL2_19, WL2_20, WL2_21, WL2_22, WL2_23, WL2_24, WL2_25)

#Loop over each dataset in WL2_list to add a new column called Stem_diam_mm_UFA where the "<1", "~1", and also numeric values <1 are changed to 0.5 and 1 so they can be used for analysis
for(i in seq_along(WL2_list)) {
  WL2_list[[i]] <- WL2_list[[i]] %>%
    mutate(Stem_diam_mm_UFA = case_when(
      Stem_diam_mm == "<1" ~ 0.5,
      Stem_diam_mm < 1 ~ 0.5,
      Stem_diam_mm == "~1" ~ 1,
      TRUE ~ as.numeric(Stem_diam_mm)
    ))
} #there are some NA coersion warnings but verified no new NAs were introduced unexpectedly
```

##### Convert all "NA" strings to NA's

```{r}
WL2_list <- map(WL2_list, ~.x %>%
                 mutate(across(where(is.character), ~na_if(., "NA"))))
```

##### Combine all WL2 dataframes and write csv

```{r}
WL2_all <- WL2_list %>%
  lapply(convert_all_columns) %>%
  bind_rows()

write.csv(WL2_all, "../input/Cleaned/WL2_cleaned_all.csv", row.names = FALSE)
```

##### Filter out 2024 and write csv - this is because 2024 was an abbreviated census

```{r}
#create csv for WL2 years 2016-2023 
WL2_16_23 <- WL2_all %>%   filter(Census_year != 2024)  

write.csv(WL2_16_23, "../input/Cleaned/WL2_cleaned_16_23.csv", row.names = FALSE) 

#create csv for WL2 2024 
WL2_cleaned_24 <- WL2_all %>%   filter(Census_year == 2024)  

write.csv(WL2_cleaned_24, "../input/Cleaned/WL2_cleaned_24.csv", row.names = FALSE)
```

#### WL3 2017-2023, abv 2024

```{r}
WL3_17 <- read_excel("../input/Annual census data/Demog_WL3_17_20170920_SLZ_cleaned.xlsx")

WL3_18 <- read_excel("../input/Annual census data/WL3_annualcensus_data_2018_cleaned.xlsx")

WL3_19 <- read_excel("../input/Annual census data/WL3_FullCensus_8.7.19_edited9.13.19MM_cleaned.xlsx")

WL3_20 <- read_excel("../input/Annual census data/WL3_annualcensus_2020_finished_cleaned.xlsx") 

WL3_21 <- read_excel("../input/Annual census data/WL3_annualcensus_2021_verified_cleaned.xlsx")

WL3_22 <- read_excel("../input/Annual census data/WL3_annualcensus_20220727_complete_YHW_verified_cleaned.xlsx")

WL3_23 <- read_excel("../input/Annual census data/WL3_annualcensus_2023_finished_cleaned.xlsx")

WL3_24 <- read_excel("../input/Annual census data/WL3_annualcensus_2024_verified_cleaned.xlsx")
```

##### Check colnames are the same for all WL3 dataframes

```{r}
WL3_col_check <- list(
  WL3_17 = colnames(WL3_17),
  WL3_18 = colnames(WL3_18),
  WL3_19 = colnames(WL3_19),
  WL3_20 = colnames(WL3_20),
  WL3_21 = colnames(WL3_21),
  WL3_22 = colnames(WL3_22),
  WL3_23 = colnames(WL3_23),
  WL3_24 = colnames(WL3_24)
)

WL3_all_col <- unique(unlist(WL3_col_check))

WL3_col_presence <- map_dfr(WL3_all_col, function(col) {
  tibble(
    column = col,
    dataframes = names(WL3_col_check)[map_lgl(WL3_col_check, ~ col %in% .x)]
  )
}, .id = "col_id") %>%
  group_by(column) %>%
  summarise(dataframes = paste(dataframes, collapse = ", "),
            count = n()) %>%
  arrange(count)

print(WL3_col_presence)
```

##### Update colnames that are different

```{r}
#update WL3_18, WL3_20, WL3_24
WL3_18 <- WL3_18 %>%
  rename(Longest_leaf_cm=Longest_leaf)

WL3_20 <- WL3_20 %>%
  rename(Band_num=Band_number)

WL3_24 <- WL3_24 %>%
  rename(Num_repro_branches=Num_repro_branch)
```

##### Address individual rows/cells preventing datasets from merging

```{r}
#WL3 has ranges in Quad (2017, 2018)

WL3_17 <- WL3_17 %>%
  mutate(Longest_fruit_cm = case_when(
    Longest_fruit_cm == "4.5cm" ~ 4.5,
    TRUE ~ as.numeric(Longest_fruit_cm)
  )) %>% #update row with Longest_fruit_cm "4.5cm" to 4.5
  mutate(
    Notes = ifelse(Num_repro_branches == "2*" & !is.na(Notes) & Notes != "",
                   paste(Notes, "4/29/25 MMA updated Num_repro_branches to intact count of 2"),
                   ifelse(Num_repro_branches == "2*",
                          "4/29/25 MMA updated Num_repro_branches to intact count of 2",
                          Notes)),
    Num_repro_branches = ifelse(Num_repro_branches == "2*", "2", Num_repro_branches)
  )

WL3_22 <- WL3_22 %>%
  mutate(Num_total_branches = case_when(
    Num_total_branches == "N" ~ NA_integer_,
    TRUE ~ as.numeric(Num_total_branches)
  ))

```

##### Add a new column Stem_diam_mm_UFA that is numeric - convert \<1 and \~1 to numerics

```{r}
WL3_list <- list(WL3_17, WL3_18, WL3_19, WL3_20, WL3_21, WL3_22, WL3_23, WL3_24)

#Loop over each dataset in WL3_list to add a new column called Stem_diam_mm_UFA where the "<1", "~1", and also numeric values <1 are changed to 0.5 and 1 so they can be used for analysis
for(i in seq_along(WL3_list)) {
  WL3_list[[i]] <- WL3_list[[i]] %>%
    mutate(Stem_diam_mm_UFA = case_when(
      Stem_diam_mm == "<1" ~ 0.5,
      Stem_diam_mm < 1 ~ 0.5,
      Stem_diam_mm == "~1" ~ 1,
      TRUE ~ as.numeric(Stem_diam_mm)
    ))
} #there are some NA coersion warnings but verified no new NAs were introduced unexpectedly
```

##### Convert all "NA" strings to NA's

```{r}
WL3_list <- map(WL3_list, ~.x %>%
                 mutate(across(where(is.character), ~na_if(., "NA"))))
```

##### Combine all WL3 dataframes and write csv

```{r}
WL3_all <- WL3_list %>%
  lapply(convert_all_columns) %>%
  bind_rows()

write.csv(WL3_all, "../input/Cleaned/WL3_cleaned_all.csv", row.names = FALSE)
```

##### Filter out 2024 and write csv - this is because 2024 was an abbreviated census

```{r}
#create csv for WL3 years 2017-2023 
WL3_17_23 <- WL3_all %>%   filter(Census_year != 2024)  

write.csv(WL3_17_23, "../input/Cleaned/WL3_cleaned_17_23.csv", row.names = FALSE) 

#create csv for WL3 2024 
WL3_cleaned_24 <- WL3_all %>%   filter(Census_year == 2024)  

write.csv(WL3_cleaned_24, "../input/Cleaned/WL3_cleaned_24.csv", row.names = FALSE)
```

#### YOSE1 2019-2023, abv 2024

```{r}
YOSE1_19 <- read_excel("../input/Annual census data/Yose1_AnnualCensus_7.29.19_edited9.10.19MM_cleaned.xlsx")

YOSE1_20 <- read_excel("../input/Annual census data/YOSE1_annualcensus_2020_finished_not_checked_cleaned.xlsx")

YOSE1_21 <- read_excel("../input/Annual census data/Yose1_annualcensus_2021_finished_cleaned.xlsx")

YOSE1_22 <- read_excel("../input/Annual census data/YOSE1_annualcensus_2022_completed_verified_cleaned.xlsx")

YOSE1_23 <- read_excel("../input/Annual census data/YOSE1_annualcensus_2023_finished_cleaned.xlsx")

YOSE1_24 <- read_excel("../input/Annual census data/YOSE1_annualcensus_2024_verified_cleaned.xlsx")

```

##### Check colnames are the same for all YOSE1 dataframes and update those that aren't - DONE

```{r}
YOSE1_col_check <- list(
  YOSE1_19 = colnames(YOSE1_19),
  YOSE1_20 = colnames(YOSE1_20),
  YOSE1_21 = colnames(YOSE1_21),
  YOSE1_22 = colnames(YOSE1_22),
  YOSE1_23 = colnames(YOSE1_23),
  YOSE1_24 = colnames(YOSE1_24)
)

YOSE1_all_col <- unique(unlist(YOSE1_col_check))

YOSE1_col_presence <- map_dfr(YOSE1_all_col, function(col) {
  tibble(
    column = col,
    dataframes = names(YOSE1_col_check)[map_lgl(YOSE1_col_check, ~ col %in% .x)]
  )
}, .id = "col_id") %>%
  group_by(column) %>%
  summarise(dataframes = paste(dataframes, collapse = ", "),
            count = n()) %>%
  arrange(count)

print(YOSE1_col_presence)

#update YOSE1_19
YOSE1_19 <- YOSE1_19 %>%
  rename(Num_fruits=Num_fruit)
```

```{r}
#YOSE1_20 doesn't have any Census_dates
#YOSE1_21 Census_date are messed up, Ycoord has missing data and "<1"
#YOSE1_22 has missing Ycoord data


YOSE1_list <- list(YOSE1_19, YOSE1_20, YOSE1_21, YOSE1_22, YOSE1_23, YOSE1_24)

YOSE1_list <- map(YOSE1_list, ~.x %>%
                 mutate(across(where(is.character), ~na_if(., "NA"))))

YOSE1 <- YOSE1_list %>%
  lapply(convert_all_columns) %>%
  bind_rows()
```

##### Add a new column Stem_diam_mm_UFA that is numeric - convert \<1 and \~1 to numerics

```{r}
#Loop over each dataset in YOSE1_list to add a new column called Stem_diam_mm_UFA where the "<1", "~1", and also numeric values <1 are changed to 0.5 and 1 so they can be used for analysis
for(i in seq_along(YOSE1_list)) {
  YOSE1_list[[i]] <- YOSE1_list[[i]] %>%
    mutate(Stem_diam_mm_UFA = case_when(
      Stem_diam_mm == "<1" ~ 0.5,
      Stem_diam_mm < 1 ~ 0.5,
      Stem_diam_mm == "~1" ~ 1,
      TRUE ~ as.numeric(Stem_diam_mm)
    ))
}
```

##### Convert all "NA" strings to NA's

```{r}
YOSE1_list <- map(YOSE1_list, ~.x %>%
                 mutate(across(where(is.character), ~na_if(., "NA"))))
```

##### Combine all YOSE1 censuses and write csv

```{r}
YOSE1_all <- YOSE1_list %>%
  lapply(convert_all_columns) %>%
  bind_rows()

write.csv(YOSE1_all, "../input/Cleaned/YOSE1_cleaned_all.csv", row.names = FALSE)
```

##### Filter out 2024 and write csv - this is because 2024 was an abbreviated census

```{r}
#create csv for YOSE1 years 2018-2023 
YOSE1_18_23 <- YOSE1_all %>%   filter(Census_year != 2024)  

write.csv(YOSE1_18_23, "../input/Cleaned/YOSE1_cleaned_18_23.csv", row.names = FALSE) 

#create csv for YOSE1 2024 
YOSE1_cleaned_24 <- YOSE1_all %>%   filter(Census_year == 2024)  

write.csv(YOSE1_cleaned_24, "../input/Cleaned/YOSE1_cleaned_24.csv", row.names = FALSE)
```

#### YOSE7 2018-2023, abv 2024

```{r}
YOSE7_18 <- read_excel("../input/Annual census data/Yose7_annual census_data_08.28.18_cleaned.xlsx")

YOSE7_19 <- read_excel("../input/Annual census data/Yose7_AnnualCensus_7.30.19_edited9.10.19MM_cleaned.xlsx")

YOSE7_20 <- read_excel("../input/Annual census data/YOSE7_annualcensus_2020_finished_cleaned.xlsx")

YOSE7_21 <- read_excel("../input/Annual census data/YOSE7_annualcensus_2021_finished_cleaned.xlsx") #Check col names

YOSE7_22 <- read_excel("../input/Annual census data/YOSE7_annualcensus_2022_completed_verified_cleaned.xlsx")

YOSE7_23 <- read_excel("../input/Annual census data/YOSE7_annualcensus_2023_finished_cleaned.xlsx")

YOSE7_24 <- read_excel("../input/Annual census data/YOSE7_annualcensus_2024_verified_cleaned.xlsx")
```

##### Check colnames are the same for all YOSE7 dataframes and update those that aren't - NEED TO COME BACK TO

```{r}
YOSE7_col_check <- list(
  YOSE7_18 = colnames(YOSE7_18),
  YOSE7_19 = colnames(YOSE7_19),
  YOSE7_20 = colnames(YOSE7_20),
  YOSE7_21 = colnames(YOSE7_21),
  YOSE7_22 = colnames(YOSE7_22),
  YOSE7_23 = colnames(YOSE7_23),
  YOSE7_24 = colnames(YOSE7_24)
)

YOSE7_all_col <- unique(unlist(YOSE7_col_check))

YOSE7_col_presence <- map_dfr(YOSE7_all_col, function(col) {
  tibble(
    column = col,
    dataframes = names(YOSE7_col_check)[map_lgl(YOSE7_col_check, ~ col %in% .x)]
  )
}, .id = "col_id") %>%
  group_by(column) %>%
  summarise(dataframes = paste(dataframes, collapse = ", "),
            count = n()) %>%
  arrange(count)

print(YOSE7_col_presence)

#update YOSE7_21
YOSE7_21 <- YOSE7_21 %>%
  rename(Num_total_branches=Total_num_branches) %>%
  #YOSE7_21 has 2 Leaf_thickness_mm columns, both filled with NAs
  rename(Leaf_thickness_mm=Leaf_thickness_mm...28) %>%
  #Changing Leaf_thickness_mm...35 to Year_marked - verified with the original dataset that the column should be filled with NAs
  rename(Year_marked=Leaf_thickness_mm...35)

#YOSE7_19 has Num_fruit instead of Num_fruits
YOSE7_19 <- YOSE7_19 %>%
  rename(Num_fruits = Num_fruit)
```

```{r}
#YOSE7_18 has "N/A" in Xcoord and Ycoord, Longest_fruit_cm, Num_repro_branches
#YOSE7_22 has "-" in Xcoord
#YOSE7_23 has missing data for Longest_fruit_cm
#YOSE7_21 has "500/607" for Band_num
#YOSE7_22 has "500/607" for Band_num


YOSE7_list <- list(YOSE7_18, YOSE7_19, YOSE7_20, YOSE7_21, YOSE7_22, YOSE7_23, YOSE7_24)

YOSE7_list <- map(YOSE7_list, ~.x %>%
                 mutate(across(where(is.character), ~na_if(., "NA"))))

YOSE7 <- YOSE7_list %>%
  lapply(convert_all_columns) %>%
  bind_rows()
```

##### Add a new column Stem_diam_mm_UFA that is numeric - convert \<1 and \~1 to numerics

```{r}
#Loop over each dataset in YOSE7_list to add a new column called Stem_diam_mm_UFA where the "<1", "~1", and also numeric values <1 are changed to 0.5 and 1 so they can be used for analysis
for(i in seq_along(YOSE7_list)) {
  YOSE7_list[[i]] <- YOSE7_list[[i]] %>%
    mutate(Stem_diam_mm_UFA = case_when(
      Stem_diam_mm == "<1" ~ 0.5,
      Stem_diam_mm < 1 ~ 0.5,
      Stem_diam_mm == "~1" ~ 1,
      TRUE ~ as.numeric(Stem_diam_mm)
    ))
}
```

##### Convert all "NA" strings to NA's

```{r}
YOSE7_list <- map(YOSE7_list, ~.x %>%
                 mutate(across(where(is.character), ~na_if(., "NA"))))
```

##### Combine all YOSE7 censuses and write csv

```{r}
YOSE7_all <- YOSE7_list %>%
  lapply(convert_all_columns) %>%
  bind_rows()  

write.csv(YOSE7_all, "../input/Cleaned/YOSE7_cleaned_all.csv", row.names = FALSE)
```

##### Filter out 2024 and write csv - this is because 2024 was an abbreviated census

```{r}
#create csv for YOSE7 years 2018-2023 
YOSE7_18_23 <- YOSE7_all %>%   filter(Census_year != 2024)  

write.csv(YOSE7_18_23, "../input/Cleaned/YOSE7_cleaned_18_23.csv", row.names = FALSE) 

#create csv for YOSE7 2024 
YOSE7_cleaned_24 <- YOSE7_all %>%   filter(Census_year == 2024)  

write.csv(YOSE7_cleaned_24, "../input/Cleaned/YOSE7_cleaned_24.csv", row.names = FALSE)
```

#### YOSE8 2018-2023, abv 2024

```{r}
YOSE8_18 <- read_excel("../input/Annual census data/Yose8_2018_annualcensus_data_cleaned.xlsx")

YOSE8_19 <- read_excel("../input/Annual census data/Yose8_AnnualCensus_7.30.19_edited9.10.19MM_cleaned.xlsx")

YOSE8_20 <- read_excel("../input/Annual census data/YOSE8_annualcensus_2020_finished_cleaned.xlsx")

YOSE8_21 <- read_excel("../input/Annual census data/YOSE8_annualcensus_2021_cleaned.xlsx") #Check col names

YOSE8_22 <- read_excel("../input/Annual census data/YOSE8_annualcensus_2022_completed_verified_cleaned.xlsx")

YOSE8_23 <- read_excel("../input/Annual census data/YOSE8_annualcensus_2023_finished_cleaned.xlsx")

YOSE8_24 <- read_excel("../input/Annual census data/YOSE8_annualcensus_2024_verified_cleaned.xlsx")
```

##### Check colnames are the same for all YOSE7 dataframes and update those that aren't - DONE

```{r}
YOSE8_col_check <- list(
  YOSE8_18 = colnames(YOSE8_18),
  YOSE8_19 = colnames(YOSE8_19),
  YOSE8_20 = colnames(YOSE8_20),
  YOSE8_21 = colnames(YOSE8_21),
  YOSE8_22 = colnames(YOSE8_22),
  YOSE8_23 = colnames(YOSE8_23),
  YOSE8_24 = colnames(YOSE8_24)
)

YOSE8_all_col <- unique(unlist(YOSE8_col_check))

YOSE8_col_presence <- map_dfr(YOSE8_all_col, function(col) {
  tibble(
    column = col,
    dataframes = names(YOSE8_col_check)[map_lgl(YOSE8_col_check, ~ col %in% .x)]
  )
}, .id = "col_id") %>%
  group_by(column) %>%
  summarise(dataframes = paste(dataframes, collapse = ", "),
            count = n()) %>%
  arrange(count)

print(YOSE8_col_presence)

#update YOSE8_24
YOSE8_24 <- YOSE8_24 %>%
  rename(Num_repro_branches=Num_repro_branch)
```

```{r}
#YOSE8_18 has "N/A" in Xcoord and Ycoord and Longest_fruit_cm and Num_repro_branches
#YOSE8_21 has missing data in Xcoord and Ycoord and Longest_fruit_cm and Num_repro_branches

YOSE8_list <- list(YOSE8_18, YOSE8_19, YOSE8_20, YOSE8_21, YOSE8_22, YOSE8_23, YOSE8_24)

YOSE8_list <- map(YOSE8_list, ~.x %>%
                 mutate(across(where(is.character), ~na_if(., "NA"))))

YOSE8 <- YOSE8_list %>%
  lapply(convert_all_columns) %>%
  bind_rows()
```

##### Add a new column Stem_diam_mm_UFA that is numeric - convert \<1 and \~1 to numerics

```{r}
#Loop over each dataset in YOSE8_list to add a new column called Stem_diam_mm_UFA where the "<1", "~1", and also numeric values <1 are changed to 0.5 and 1 so they can be used for analysis 

for(i in seq_along(YOSE8_list)) {   
  YOSE8_list[[i]] <- YOSE8_list[[i]] %>%     
    mutate(Stem_diam_mm_UFA = case_when(       
      Stem_diam_mm == "<1" ~ 0.5,       
      Stem_diam_mm < 1 ~ 0.5,       
      Stem_diam_mm == "~1" ~ 1,       
      TRUE ~ as.numeric(Stem_diam_mm)    
      )) }
```

##### Convert all "NA" strings to NA's

```{r}
YOSE8_list <- map(YOSE8_list, ~.x %>%                 
                    mutate(across(where(is.character), ~na_if(., "NA"))))
```

##### Combine all YOSE8 censuses and write csv

```{r}
YOSE8_all <- YOSE8_list %>%   
  lapply(convert_all_columns) %>%   
  bind_rows()    

write.csv(YOSE8_all, "../input/Cleaned/YOSE8_cleaned_all.csv", row.names = FALSE)
```

##### Filter out 2024 and write csv - this is because 2024 was an abbreviated census

```{r}
#create csv for YOSE8 years 2018-2023  
YOSE8_18_23 <- YOSE8_all %>%  
  filter(Census_year != 2024)    

write.csv(YOSE8_18_23, "../input/Cleaned/YOSE8_cleaned_18_23.csv", row.names = FALSE)   

#create csv for YOSE8 2024  
YOSE8_cleaned_24 <- YOSE8_all %>%  
  filter(Census_year == 2024)   

write.csv(YOSE8_cleaned_24, "../input/Cleaned/YOSE8_cleaned_24.csv", row.names = FALSE)
```

#### 
