---
title: "Prelim_Cleaning"
output: html_document
date: "2025-04-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(readxl)
library(purrr)
```

### Merge all annual censuses for each pop

#### BH 2018-2023, abv 2024

```{r}
BH_18 <- read_excel("../input/Annual census data/Demog_BHR_Summer2018_8.12.2017_SLZ_cleaned.xlsx")

BH_19 <- read_excel("../input/Annual census data/BH_AnnualCensus_6.17.19_edited9.16.2019_EEB_cleaned.xlsx")

BH_20 <- read_excel("../input/Annual census data/BH_annualcensus_census 2020_finished_EC_cleaned.xlsx")

BH_21 <- read_excel("../input/Annual census data/BH_annualcensus_2021_verified_cleaned.xlsx")

BH_22 <- read_excel("../input/Annual census data/BH_annualcensus_2021_verified_cleaned.xlsx")

BH_23 <- read_excel("../input/Annual census data/BH_annualcensus_2023_finished_cleaned.xlsx")

BH_24 <- read_excel("../input/Annual census data/BH_annualcensus_2024_verified_cleaned.xlsx")
```

##### Check colnames are the same for all BH dataframes and update those that aren't - DONE

```{r}
BH_col_check <- list(
  BH_18 = colnames(BH_18),
  BH_19 = colnames(BH_19),
  BH_20 = colnames(BH_20),
  BH_21 = colnames(BH_21),
  BH_22 = colnames(BH_22),
  BH_23 = colnames(BH_23),
  BH_24 = colnames(BH_24)
)

BH_all_col <- unique(unlist(BH_col_check))

BH_col_presence <- map_dfr(BH_all_col, function(col) {
  tibble(
    column = col,
    dataframes = names(BH_col_check)[map_lgl(BH_col_check, ~ col %in% .x)]
  )
}, .id = "col_id") %>%
  group_by(column) %>%
  summarise(dataframes = paste(dataframes, collapse = ", "),
            count = n()) %>%
  arrange(count)

print(BH_col_presence)

#update BH_19 and BH_23 to have matching colnames
BH_19 <- BH_19 %>%
  rename(Num_total_branches='Num_total_ branches')
BH_23 <- BH_23 %>%
  rename(Num_total_branches=Num_total_branch)
```

##### Combine all BH dataframes - WHAT IS A MORE EFFICIENT WAY TO DO THIS?

```{r}
#convert Quad column to character
convert_Quad <- function(df) {
  df %>% mutate(Quad = as.character(Quad))
}

convert_Date_entered <- function(df) {
  df %>% mutate(Date_entered = as.date(Date_entered))
}

BH_18$Quad <- as.character(BH_18$Quad)
BH_18$Date_entered <- as.Date(BH_18$Date_entered)
BH_19$Date_entered <- as.Date(BH_19$Date_entered)


BH <- bind_rows(BH_18, BH_19, BH_20, BH_21, BH_22, BH_23, BH_24)
```

#### CP2 2017-2023, abv 2024

```{r}
CP2_17 <- read_excel("../input/Annual census data/Demog_CP_17_20170920_SLZ_cleaned.xlsx")

CP2_18 <- read_excel("../input/Annual census data/CP2_AnnualCensus_7.30.2018_9.7.2018_SLZ_cleaned.xlsx")

CP2_19 <- read_excel("../input/Annual census data/CP2_FullCensus_8.7.19_edited9.13.19MM_cleaned.xlsx")

CP2_20 <- read_excel("../input/Annual census data/CP2_annualcensus_2020_finished_cleaned.xlsx")

CP2_21 <- read_excel("../input/Annual census data/CP2_annualcensus_2021_verified_cleaned.xlsx")

CP2_22 <- read_excel("../input/Annual census data/CP2_annualcensus_20220729_verified_YHW_cleaned.xlsx")

CP2_23 <- read_excel("../input/Annual census data/CP2_annualcensus_2023_finished_cleaned.xlsx")

CP2_24 <- read_excel("../input/Annual census data/CP2_annualcensus_2024_verified_cleaned.xlsx")
```

##### Check colnames are the same for all CP2 dataframes and update those that aren't - DONE

```{r}
CP2_col_check <- list(
  CP2_17 = colnames(CP2_17),
  CP2_18 = colnames(CP2_18),
  CP2_19 = colnames(CP2_19),
  CP2_20 = colnames(CP2_20),
  CP2_21 = colnames(CP2_21),
  CP2_22 = colnames(CP2_22),
  CP2_23 = colnames(CP2_23),
  CP2_24 = colnames(CP2_24)
)

CP2_all_col <- unique(unlist(CP2_col_check))

CP2_col_presence <- map_dfr(CP2_all_col, function(col) {
  tibble(
    column = col,
    dataframes = names(CP2_col_check)[map_lgl(CP2_col_check, ~ col %in% .x)]
  )
}, .id = "col_id") %>%
  group_by(column) %>%
  summarise(dataframes = paste(dataframes, collapse = ", "),
            count = n()) %>%
  arrange(count)

print(CP2_col_presence)

#update CP2_19 and CP2_24 to have matching colnames
CP2_19 <- CP2_19 %>%
  rename(Height_cm = Height) %>%
  rename(Num_fruits = Num_fruit)

CP2_24 <- CP2_24 %>%
  rename(Num_repro_branches=Num_repro_branch)
```

##### Combine all CP2 dataframes

```{r}
CP2 <- bind_rows(CP2_17, CP2_18, CP2_19, CP2_20, CP2_21, CP2_22, CP2_23, CP2_24)
```

#### TM2 2016-2023, abv 2024

```{r}
TM2_16 <- read_excel("../input/Annual census data/Demog_TM2_spring_2016_foranalysis_cleaned.xlsx")

TM2_17 <- read_excel("../input/Annual census data/TM2_annualcensus_2017_jrg_fixederrors_foranalysis_cleaned.xlsx")

TM2_18 <- read_excel("../input/Annual census data/TM2_annualcensus_data_2018_foranalysis_cleaned.xlsx") #Check col names

TM2_19 <- read_excel("../input/Annual census data/TM2_AnnualCensus_2019_all_edited6.30.21JRG_cleaned.xlsx")

TM2_20 <- read_excel("../input/Annual census data/TM2_AnnualCensus_2020_foranalysis_cleaned.xlsx") #Check col names

TM2_21 <- read_excel("../input/Annual census data/TM2_annualcensus_2021_finished_cleaned.xlsx")

TM2_22 <- read_excel("../input/Annual census data/TM2_annual_census_2022_complete_YHW_Verified_cleaned.xlsx")

TM2_23 <- read_excel("../input/Annual census data/TM2_annualcensus_2023_finished_cleaned.xlsx")

TM2_24 <- read_excel("../input/Annual census data/TM2_annualcensus_2024_verified_cleaned.xlsx")
```

##### Check colnames are the same for all TM2 dataframes and update those that aren't - Done

```{r}
TM2_col_check <- list(
  TM2_16 = colnames(TM2_16),
  TM2_17 = colnames(TM2_17),
  TM2_18 = colnames(TM2_18),
  TM2_19 = colnames(TM2_19),
  TM2_20 = colnames(TM2_20),
  TM2_21 = colnames(TM2_21),
  TM2_22 = colnames(TM2_22),
  TM2_23 = colnames(TM2_23),
  TM2_24 = colnames(TM2_24)
)

TM2_all_col <- unique(unlist(TM2_col_check))

TM2_col_presence <- map_dfr(TM2_all_col, function(col) {
  tibble(
    column = col,
    dataframes = names(TM2_col_check)[map_lgl(TM2_col_check, ~ col %in% .x)]
  )
}, .id = "col_id") %>%
  group_by(column) %>%
  summarise(dataframes = paste(dataframes, collapse = ", "),
            count = n()) %>%
  arrange(count)

print(TM2_col_presence)


#update TM2_18, TM2_19, and TM2_20 to have matching colnames

#Pick_color...25 and Sword_color...24 to Num_repro_branches and Num_total_branches but need numbers for those cols
TM2_18 <- TM2_18 %>%
  #TM2_18 has 2 Pick_color columns but Pick_color...13 is the column filled with colors, Pick_color...25 is filled with NAs
  rename(Pick_color=Pick_color...13) %>% 
  #TM2_28 has 2 Sword_color columns but Sword_color...14 is the column filled with colors, Sword_color...24 is filled with NAs 
  rename(Sword_color=Sword_color...14) %>%
  #Changing pick_color...25 and Sword_color...24 to Num_repro_branches and Num_total_branches - verified using original data set that these cols were added in the initial cleaning and should be filled with NAs
  rename(Num_repro_branches=Pick_color...25) %>%
  rename(Num_total_branches=Sword_color...24)
  
TM2_19 <- TM2_19 %>%
  rename(Leaf_thickness_mm=Leaf_thickness)

TM2_20 <- TM2_20 %>%
  #TM2_20 has 2 Category_herbivory columns both filled with NAs so I'm renaming Category_herbivory...23
  rename(Category_herbivory=Category_herbivory...23) %>% 
  rename(Num_total_branches=Total_branches) %>%
  #Changing Category_herbivory...31 to Leaf_thickness_mm - verified using origninal data set that these cols were added in the initial cleaning and should be filled with NAs
  rename(Leaf_thickness_mm=Category_herbivory...31)
```

#### WL1 2017-2023, abv 2024

```{r}
WL1_17 <- read_excel("../input/Annual census data/Demog_WL1_17_20170920_SLZ_cleaned.xlsx")

WL1_18 <- read_excel("../input/Annual census data/WL1_AnnualCensus_7.06.2018_MM_cleaned.xlsx")

WL1_19 <- read_excel("../input/Annual census data/WL1_FullCensus_8.7.19_edited9.13.19MM_cleaned.xlsx")

WL1_20 <- read_excel("../input/Annual census data/WL1_annualcensus_2020_finished_cleaned.xlsx") 

WL1_21 <- read_excel("../input/Annual census data/WL1_annualcensus_2021_verified_cleaned.xlsx")

WL1_22 <- read_excel("../input/Annual census data/WL1_annualcensus_20220708_YHW_verified_cleaned.xlsx")

WL1_23 <- read_excel("../input/Annual census data/WL1_annualcensus_2023_finished_cleaned.xlsx")

WL1_24 <- read_excel("../input/Annual census data/WL1_annualcensus_2024_verified_cleaned.xlsx")
```

##### Check colnames are the same for all WL1 dataframes and update those that aren't - DONE

```{r}
WL1_col_check <- list(
  WL1_17 = colnames(WL1_17),
  WL1_18 = colnames(WL1_18),
  WL1_19 = colnames(WL1_19),
  WL1_20 = colnames(WL1_20),
  WL1_21 = colnames(WL1_21),
  WL1_22 = colnames(WL1_22),
  WL1_23 = colnames(WL1_23),
  WL1_24 = colnames(WL1_24)
)

WL1_all_col <- unique(unlist(WL1_col_check))

WL1_col_presence <- map_dfr(WL1_all_col, function(col) {
  tibble(
    column = col,
    dataframes = names(WL1_col_check)[map_lgl(WL1_col_check, ~ col %in% .x)]
  )
}, .id = "col_id") %>%
  group_by(column) %>%
  summarise(dataframes = paste(dataframes, collapse = ", "),
            count = n()) %>%
  arrange(count)

print(WL1_col_presence)

#update WL1_18, WL1_20, WL1_21, WL1_22, WL1_23, WL1_24
WL1_20 <- WL1_20 %>%
  rename(Band_num=Band_number)

WL1_18 <- WL1_18 %>%
  rename(Canopy_width_cm=Canopy_width)

WL1_22 <- WL1_22 %>%
  rename(Stem_diam_mm=Stem_Diam_mm) %>%
  rename(Xcoord=X_coord) %>%
  rename(Ycoord=Y_coord)

WL1_24 <- WL1_24 %>%
  rename(Year_marked='Year _marked') %>%
  rename(Phenology=Pheno)

WL1_23 <- WL1_23 %>%
  rename(Phenology=Pheno)

WL1_21 <- WL1_21 %>%
  rename(Xcoord=X_coord) %>%
  rename(Ycoord=Y_coord)
```

#### WL2 2016-2024

```{r}
WL2_16 <- read_excel("../input/Annual census data/Demog_WL2_summer_2016_20170622_AJC_cleaned.xlsx")

WL2_17 <- read_excel("../input/Annual census data/Demog_WL2_17_20170920_SLZ_cleaned.xlsx")

WL2_18 <- read_excel("../input/Annual census data/WL2_AnnualCensus_7.25.18_20180819_SLZ_cleaned.xlsx")

WL2_19 <- read_excel("../input/Annual census data/WL2_FullCensus_8.7.19_edited9.13.19MM_cleaned.xlsx")

WL2_20 <- read_excel("../input/Annual census data/WL2_annualcensus_2020_finished_cleaned.xlsx") #Check col names

WL2_21 <- read_excel("../input/Annual census data/WL2_annualcensus_2021_verified_cleaned.xlsx")

WL2_22 <- read_excel("../input/Annual census data/WL2_Annual_Census_2022_Verified_cleaned.xlsx")

WL2_23 <- read_excel("../input/Annual census data/WL2_annualcensus_2023_finished_cleaned.xlsx")

WL2_24 <- read_excel("../input/Annual census data/WL2_annualcensus_2024_verified_cleaned.xlsx")
```

##### Check colnames are the same for all WL2 dataframes and update those that aren't - DONE

```{r}
WL2_col_check <- list(
  WL2_16 = colnames(WL2_16),
  WL2_17 = colnames(WL2_17),
  WL2_18 = colnames(WL2_18),
  WL2_19 = colnames(WL2_19),
  WL2_20 = colnames(WL2_20),
  WL2_21 = colnames(WL2_21),
  WL2_22 = colnames(WL2_22),
  WL2_23 = colnames(WL2_23),
  WL2_24 = colnames(WL2_24)
)

WL2_all_col <- unique(unlist(WL2_col_check))

WL2_col_presence <- map_dfr(WL2_all_col, function(col) {
  tibble(
    column = col,
    dataframes = names(WL2_col_check)[map_lgl(WL2_col_check, ~ col %in% .x)]
  )
}, .id = "col_id") %>%
  group_by(column) %>%
  summarise(dataframes = paste(dataframes, collapse = ", "),
            count = n()) %>%
  arrange(count)

print(WL2_col_presence)

#update WL2_16, WL2_20, WL2_22
WL2_16 <- WL2_16 %>%
  rename(Height_cm=Plant_height_cm)

WL2_20 <- WL2_20 %>%
  rename(Band_num=Band_number) %>%
  rename(Num_leaves=Num_leaves...34) %>% #WL2_20 has 2 Num_leaves columns, both filled with NA so I will rename Num_leaves...34 
  select(-Num_leaves...35)

WL2_22 <- WL2_22 %>%
  rename(Phenology=Pheno)
```

#### WL3 2017-2023, abv 2024

```{r}
WL3_17 <- read_excel("../input/Annual census data/Demog_WL3_17_20170920_SLZ_cleaned.xlsx")

WL3_18 <- read_excel("../input/Annual census data/WL3_annualcensus_data_2018_cleaned.xlsx")

WL3_19 <- read_excel("../input/Annual census data/WL3_FullCensus_8.7.19_edited9.13.19MM_cleaned.xlsx")

WL3_20 <- read_excel("../input/Annual census data/WL3_annualcensus_2020_finished_cleaned.xlsx") 

WL3_21 <- read_excel("../input/Annual census data/WL3_annualcensus_2021_verified_cleaned.xlsx")

WL3_22 <- read_excel("../input/Annual census data/WL3_annualcensus_20220727_complete_YHW_verified_cleaned.xlsx")

WL3_23 <- read_excel("../input/Annual census data/WL3_annualcensus_2023_finished_cleaned.xlsx")

WL3_24 <- read_excel("../input/Annual census data/WL3_annualcensus_2024_verified_cleaned.xlsx")
```

##### Check colnames are the same for all WL3 dataframes and update those that aren't - DONE

```{r}
WL3_col_check <- list(
  WL3_17 = colnames(WL3_17),
  WL3_18 = colnames(WL3_18),
  WL3_19 = colnames(WL3_19),
  WL3_20 = colnames(WL3_20),
  WL3_21 = colnames(WL3_21),
  WL3_22 = colnames(WL3_22),
  WL3_23 = colnames(WL3_23),
  WL3_24 = colnames(WL3_24)
)

WL3_all_col <- unique(unlist(WL3_col_check))

WL3_col_presence <- map_dfr(WL3_all_col, function(col) {
  tibble(
    column = col,
    dataframes = names(WL3_col_check)[map_lgl(WL3_col_check, ~ col %in% .x)]
  )
}, .id = "col_id") %>%
  group_by(column) %>%
  summarise(dataframes = paste(dataframes, collapse = ", "),
            count = n()) %>%
  arrange(count)

print(WL3_col_presence)

#update WL3_18, WL3_20, WL2_24
WL3_18 <- WL3_18 %>%
  rename(Longest_leaf_cm=Longest_leaf)

WL3_20 <- WL3_20 %>%
  rename(Band_num=Band_number)

WL3_24 <- WL3_24 %>%
  rename(Num_repro_branches=Num_repro_branch)
```

#### YOSE1 2019-2023, abv 2024

```{r}
YOSE1_19 <- read_excel("../input/Annual census data/Yose1_AnnualCensus_7.29.19_edited9.10.19MM_cleaned.xlsx")

YOSE1_20 <- read_excel("../input/Annual census data/YOSE1_annualcensus_2020_finished_not_checked_cleaned.xlsx")

YOSE1_21 <- read_excel("../input/Annual census data/Yose1_annualcensus_2021_finished_cleaned.xlsx")

YOSE1_22 <- read_excel("../input/Annual census data/YOSE1_annualcensus_2022_completed_verified_cleaned.xlsx")

YOSE1_23 <- read_excel("../input/Annual census data/YOSE1_annualcensus_2023_finished_cleaned.xlsx")

YOSE1_24 <- read_excel("../input/Annual census data/YOSE1_annualcensus_2024_verified_cleaned.xlsx")

```

##### Check colnames are the same for all YOSE1 dataframes and update those that aren't - DONE

```{r}
YOSE1_col_check <- list(
  YOSE1_19 = colnames(YOSE1_19),
  YOSE1_20 = colnames(YOSE1_20),
  YOSE1_21 = colnames(YOSE1_21),
  YOSE1_22 = colnames(YOSE1_22),
  YOSE1_23 = colnames(YOSE1_23),
  YOSE1_24 = colnames(YOSE1_24)
)

YOSE1_all_col <- unique(unlist(YOSE1_col_check))

YOSE1_col_presence <- map_dfr(YOSE1_all_col, function(col) {
  tibble(
    column = col,
    dataframes = names(YOSE1_col_check)[map_lgl(YOSE1_col_check, ~ col %in% .x)]
  )
}, .id = "col_id") %>%
  group_by(column) %>%
  summarise(dataframes = paste(dataframes, collapse = ", "),
            count = n()) %>%
  arrange(count)

print(YOSE1_col_presence)

#update YOSE1_19
YOSE1_19 <- YOSE1_19 %>%
  rename(Num_fruits=Num_fruit)
```

#### YOSE7 2018-2023, abv 2024

```{r}
YOSE7_18 <- read_excel("../input/Annual census data/Yose8_2018_annualcensus_data_cleaned.xlsx")

YOSE7_19 <- read_excel("../input/Annual census data/Yose8_AnnualCensus_7.30.19_edited9.10.19MM_cleaned.xlsx")

YOSE7_20 <- read_excel("../input/Annual census data/YOSE7_annualcensus_2020_finished_cleaned.xlsx")

YOSE7_21 <- read_excel("../input/Annual census data/YOSE7_annualcensus_2021_finished_cleaned.xlsx") #Check col names

YOSE7_22 <- read_excel("../input/Annual census data/YOSE7_annualcensus_2022_completed_verified_cleaned.xlsx")

YOSE7_23 <- read_excel("../input/Annual census data/YOSE7_annualcensus_2023_finished_cleaned.xlsx")

YOSE7_24 <- read_excel("../input/Annual census data/YOSE7_annualcensus_2024_verified_cleaned.xlsx")
```

##### Check colnames are the same for all YOSE7 dataframes and update those that aren't - NEED TO COME BACK TO

```{r}
YOSE7_col_check <- list(
  YOSE7_18 = colnames(YOSE7_18),
  YOSE7_19 = colnames(YOSE7_19),
  YOSE7_20 = colnames(YOSE7_20),
  YOSE7_21 = colnames(YOSE7_21),
  YOSE7_22 = colnames(YOSE7_22),
  YOSE7_23 = colnames(YOSE7_23),
  YOSE7_24 = colnames(YOSE7_24)
)

YOSE7_all_col <- unique(unlist(YOSE7_col_check))

YOSE7_col_presence <- map_dfr(YOSE7_all_col, function(col) {
  tibble(
    column = col,
    dataframes = names(YOSE7_col_check)[map_lgl(YOSE7_col_check, ~ col %in% .x)]
  )
}, .id = "col_id") %>%
  group_by(column) %>%
  summarise(dataframes = paste(dataframes, collapse = ", "),
            count = n()) %>%
  arrange(count)

print(YOSE7_col_presence)

#update YOSE7_21
#NEED TO CHANGE Leaf_thickness_mm...35 to Year_marked but need data
YOSE7_21 <- YOSE7_21 %>%
  rename(Num_total_branches=Total_num_branches) %>%
  rename(Leaf_thickness_mm=Leaf_thickness_mm...28) #YOSE7_21 has 2 Leaf_thickness_mm columns, both filled with NAs
```

#### YOSE8 2018-2023, abv 2024

```{r}
YOSE8_18 <- read_excel("../input/Annual census data/Yose8_2018_annualcensus_data_cleaned.xlsx")

YOSE8_19 <- read_excel("../input/Annual census data/Yose8_AnnualCensus_7.30.19_edited9.10.19MM_cleaned.xlsx")

YOSE8_20 <- read_excel("../input/Annual census data/YOSE8_annualcensus_2020_finished_cleaned.xlsx")

YOSE8_21 <- read_excel("../input/Annual census data/YOSE8_annualcensus_2021_cleaned.xlsx") #Check col names

YOSE8_22 <- read_excel("../input/Annual census data/YOSE8_annualcensus_2022_completed_verified_cleaned.xlsx")

YOSE8_23 <- read_excel("../input/Annual census data/YOSE8_annualcensus_2023_finished_cleaned.xlsx")

YOSE8_24 <- read_excel("../input/Annual census data/YOSE8_annualcensus_2024_verified_cleaned.xlsx")
```

##### Check colnames are the same for all YOSE7 dataframes and update those that aren't - DONE

```{r}
YOSE8_col_check <- list(
  YOSE8_18 = colnames(YOSE8_18),
  YOSE8_19 = colnames(YOSE8_19),
  YOSE8_20 = colnames(YOSE8_20),
  YOSE8_21 = colnames(YOSE8_21),
  YOSE8_22 = colnames(YOSE8_22),
  YOSE8_23 = colnames(YOSE8_23),
  YOSE8_24 = colnames(YOSE8_24)
)

YOSE8_all_col <- unique(unlist(YOSE8_col_check))

YOSE8_col_presence <- map_dfr(YOSE8_all_col, function(col) {
  tibble(
    column = col,
    dataframes = names(YOSE8_col_check)[map_lgl(YOSE8_col_check, ~ col %in% .x)]
  )
}, .id = "col_id") %>%
  group_by(column) %>%
  summarise(dataframes = paste(dataframes, collapse = ", "),
            count = n()) %>%
  arrange(count)

print(YOSE8_col_presence)

#update YOSE8_24
YOSE8_24 <- YOSE8_24 %>%
  rename(Num_repro_branches=Num_repro_branch)
```
