---
title: "WL1_Survival_Across_Years"
output: html_document
date: "2026-05-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(lubridate)
```
Initial test with WL1 2023-2024
```{r}
WL1_all <- read_csv("../input/Cleaned/WL1_cleaned.csv") %>%
  select(-'...1') 

WL1_24 <- WL1_all %>%
  filter(Census_year == 2024) %>%
  mutate(plant_id = paste(Site, Band_color, Band_num, sep = "_"),
         alive = as.integer(!is.na(Phenology) & Phenology != "X"))
  

WL1_23 <- WL1_all %>%
  filter(Census_year == 2023) %>%
  mutate(plant_id = paste(Site, Band_color, Band_num, sep = "_"),
         alive = as.integer(!is.na(Phenology) & Phenology != "X"))
```
```{r}
WL1_23_24_survival <- WL1_23 %>%
  left_join(WL1_24 %>% select(plant_id, alive_2024 = alive, Phenology_2024 = Phenology), by = "plant_id") %>%
  mutate(alive_2024 = if_else(is.na(alive_2024), FALSE, alive_2024),
         survived_23_24 = alive & alive_2024)
```

Try to generalize across all years at WL1
Need to start by only including plants that have the band number and color - not sure if this is the best way to go about it but it is what works for now
Will need to think about how this will work for years without band color and number
```{r}
WL1_survival <- WL1_all %>%
 filter(!is.na(Band_color), !is.na(Band_num)) %>% #only include individuals with a band num and color
  mutate(plant_id = paste(Site, Transect_plot, Quad, Band_color, Band_num, sep = "_"), #create a unique ID for each individual
         alive = as.integer(!is.na(Phenology) & Phenology != "X")) %>% #add a column for binary alive = 1 or dead = 0
  arrange(plant_id, Census_year) %>% #arrange individuals by census year
  group_by(plant_id) %>%
  mutate(age = row_number(), #add an age - so the first row with the unique plant id will be 1 and ever followign row would increase age by 1
         next_year = Census_year + 1, #add next year column
         alive_next_year = lead(alive), #check alive column from the following year
         survived_to_next = as.integer(next_year == Census_year + 1 & alive == 1 & alive_next_year == 1)) %>% #1 if alive = 1 in census year and alive in census year + 1
  ungroup()

#This seems to work
#alive_next_year seems to be NA if the plant was not relocated in the following year or if the plant died in the current census year
```


Test if this works for WL2
```{r}
WL2_16_23 <- read_csv("../input/Cleaned/WL2_cleaned_16_23.csv")
WL2_24 <- read_csv("../input/Cleaned/WL2_cleaned_24.csv")
WL2_all <- bind_rows(WL2_16_23, WL2_24  %>% mutate(Quad = as.character(Quad), Num_flowers = as.character(Num_flowers)))
```

```{r}
WL2_survival <- WL2_all %>%
 filter(!is.na(Band_color), !is.na(Band_num)) %>%
  mutate(plant_id = paste(Site, Transect_plot, Quad, Band_color, Band_num, sep = "_"),
         alive = as.integer(!is.na(Phenology) & Phenology != "X")) %>%
  arrange(plant_id, Census_year) %>%
  group_by(plant_id) %>%
  mutate(age = row_number(),
         next_year = Census_year + 1,
         alive_next_year = lead(alive),
         survived_to_next = as.integer(next_year == Census_year + 1 & alive == 1 & alive_next_year == 1)) %>%
  ungroup()

#Noticed some individuals have a 0 for survived to next year = 0 for 2025, unsure why these are getting assigned a 0 and not NA
```

Testing for WL3
```{r}
WL3_all <- read_csv("..//input/Cleaned/WL3_cleaned_all.csv")

WL3_survival <- WL3_all %>%
 filter(!is.na(Band_color), !is.na(Band_num)) %>%
  mutate(plant_id = paste(Site, Transect_plot, Quad, Band_color, Band_num, sep = "_"),
         alive = as.integer(!is.na(Phenology) & Phenology != "X")) %>%
  arrange(plant_id, Census_year) %>%
  group_by(plant_id) %>%
mutate(age = row_number(),
       next_year = Census_year + 1,
       alive_next_year = lead(alive),
         survived_to_next = as.integer(next_year == Census_year + 1 & alive == 1 & alive_next_year == 1)) %>%
  ungroup()
```

Testing for BH
```{r}
BH_all <- read_csv("../input/Cleaned/BH_cleaned_all.csv")

BH_survival <- BH_all %>%
 filter(!is.na(Band_color), !is.na(Band_num)) %>%
  mutate(plant_id = paste(Site, Transect_plot, Quad, Band_color, Band_num, sep = "_"),
         alive = as.integer(!is.na(Phenology) & Phenology != "X")) %>%
  arrange(plant_id, Census_year) %>%
  group_by(plant_id) %>%
mutate(age = row_number(),
       next_year = Census_year + 1,
       alive_next_year = lead(alive),
         survived_to_next = as.integer(next_year == Census_year + 1 & alive == 1 & alive_next_year == 1)) %>%
  ungroup()
```

Testing for CP2
```{r}
CP2_all <- read_csv("../input/Cleaned/CP2_cleaned_all.csv")

CP2_survival <- CP2_all %>%
 filter(!is.na(Band_color), !is.na(Band_num)) %>%
  mutate(plant_id = paste(Site,Transect_plot, Quad,  Band_color, Band_num, sep = "_"),
         alive = as.integer(!is.na(Phenology) & Phenology != "X")) %>%
  arrange(plant_id, Census_year) %>%
  group_by(plant_id) %>%
mutate(age = row_number(),
       next_year = Census_year + 1,
       alive_next_year = lead(alive),
         survived_to_next = as.integer(next_year == Census_year + 1 & alive == 1 & alive_next_year == 1)) %>%
  ungroup()
```

Testing for TM2
```{r}
TM2_all <- read_csv("../input/Cleaned/TM2_cleaned_all.csv")

TM2_survival <- TM2_all %>%
 filter(!is.na(Band_color), !is.na(Band_num)) %>%
  mutate(plant_id = paste(Site,Transect_plot, Quad, Band_color, Band_num, sep = "_"),
         alive = as.integer(!is.na(Phenology) & Phenology != "X")) %>%
  arrange(plant_id, Census_year) %>%
  group_by(plant_id) %>%
mutate(age = row_number(),
       next_year = Census_year + 1,
       alive_next_year = lead(alive),
         survived_to_next = as.integer(next_year == Census_year + 1 & alive == 1 & alive_next_year == 1)) %>%
  ungroup()
```
Preliminary visualization of survival differences across pops
summarize across years and calc sem?
```{r}
sem <- function(x, na.rm=FALSE) {           #for calculating standard error
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} 

WL1_survival <- WL1_survival %>%
  group_by(plant_id) %>%
  mutate(original_census_year = min(Census_year)) %>% #create a column so we know when the plant was first censused
  ungroup()

#The fates of 1st year plants observed in 2024 haven't been uploaded yet so we exclude them for now
max_year_WL1 <- max(WL1_survival$Census_year)

#remove plants that first appeared in 2024 and were not observed beyond age 1
survival_by_year_WL1 <- WL1_survival %>%
  group_by(plant_id, original_census_year) %>%
  summarise(max_age = max(age),
            survived_to_age2 = as.integer(any(age == 2 & alive == 1)),
            .groups = "drop") %>%
  filter(!(original_census_year == max_year_WL1 & max_age == 1))
  
#summarize by year
WL1_survival_by_year <- survival_by_year_WL1 %>%
  group_by(original_census_year) %>%
  summarize(survival = mean(survived_to_age2),
            survival_se = sem(survived_to_age2),
            n=n(),
            .groups = "drop")

#plot survival differences across years at WL1
ggplot(WL1_survival_by_year, aes(x=original_census_year, survival)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin=survival - survival_se,
                    ymax = survival + survival_se),
                width = 0.15) +
  scale_y_continuous(limits = c(0, 1)) +
  labs(
    x = "Original census year",
    y = "Proportion surviving to age 2"
  ) +
  theme_classic()
```
Survival differences across years at WL2
```{r}
WL2_survival <- WL2_survival %>%
  group_by(plant_id) %>%
  mutate(original_census_year = min(Census_year)) %>% #create a column so we know when the plant was first censused
  ungroup()

#The fates of 1st year plants observed in 2024 haven't been uploaded yet so we exclude them for now
max_year_WL2 <- max(WL2_survival$Census_year)

#remove plants that first appeared in 2024 and were not observed beyond age 1
survival_by_year_WL2 <- WL2_survival %>%
  group_by(plant_id, original_census_year) %>%
  summarise(max_age = max(age),
            survived_to_age2 = as.integer(any(age == 2 & alive == 1)),
            .groups = "drop") %>%
  filter(!(original_census_year == max_year_WL2 & max_age == 1))
  
#summarize by year
WL2_survival_by_year <- survival_by_year_WL2 %>%
  group_by(original_census_year) %>%
  summarize(survival = mean(survived_to_age2),
            survival_se = sem(survived_to_age2),
            n=n(),
            .groups = "drop")

#plot survival differences across years at WL2
ggplot(WL2_survival_by_year, aes(x=original_census_year, survival)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin=survival - survival_se,
                    ymax = survival + survival_se),
                width = 0.15) +
  scale_y_continuous(limits = c(0, 1)) +
  labs(
    x = "Original census year",
    y = "Proportion surviving to age 2"
  ) +
  theme_classic()
```
Survival differences across years at WL3
```{r}
WL3_survival <- WL3_survival %>%
  group_by(plant_id) %>%
  mutate(original_census_year = min(Census_year)) %>% #create a column so we know when the plant was first censused
  ungroup()

#The fates of 1st year plants observed in 2024 haven't been uploaded yet so we exclude them for now
max_year_WL3 <- max(WL3_survival$Census_year)

#remove plants that first appeared in 2024 and were not observed beyond age 1
survival_by_year_WL3 <- WL3_survival %>%
  group_by(plant_id, original_census_year) %>%
  summarise(max_age = max(age),
            survived_to_age2 = as.integer(any(age == 2 & alive == 1)),
            .groups = "drop") %>%
  filter(!(original_census_year == max_year_WL3 & max_age == 1))
  
#summarize by year
WL3_survival_by_year <- survival_by_year_WL3 %>%
  group_by(original_census_year) %>%
  summarize(survival = mean(survived_to_age2),
            survival_se = sem(survived_to_age2),
            n=n(),
            .groups = "drop")

#plot survival differences across years at WL3
ggplot(WL3_survival_by_year, aes(x=original_census_year, survival)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin=survival - survival_se,
                    ymax = survival + survival_se),
                width = 0.15) +
  scale_y_continuous(limits = c(0, 1)) +
  labs(
    x = "Original census year",
    y = "Proportion surviving to age 2"
  ) +
  theme_classic()
```
Survival differences across years at BH
```{r}
BH_survival <- BH_survival %>%
  group_by(plant_id) %>%
  mutate(original_census_year = min(Census_year)) %>% #create a column so we know when the plant was first censused
  ungroup()

#The fates of 1st year plants observed in 2024 haven't been uploaded yet so we exclude them for now
max_year_BH <- max(BH_survival$Census_year)

#remove plants that first appeared in 2024 and were not observed beyond age 1
survival_by_year_BH <- BH_survival %>%
  group_by(plant_id, original_census_year) %>%
  summarise(max_age = max(age),
            survived_to_age2 = as.integer(any(age == 2 & alive == 1)),
            .groups = "drop") %>%
  filter(!(original_census_year == max_year_BH & max_age == 1))
  
#summarize by year
BH_survival_by_year <- survival_by_year_BH %>%
  group_by(original_census_year) %>%
  summarize(survival = mean(survived_to_age2),
            survival_se = sem(survived_to_age2),
            n=n(),
            .groups = "drop")

#plot survival differences across years at WL1
ggplot(BH_survival_by_year, aes(x=original_census_year, survival)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin=survival - survival_se,
                    ymax = survival + survival_se),
                width = 0.15) +
  scale_y_continuous(limits = c(0, 1)) +
  labs(
    x = "Original census year",
    y = "Proportion surviving to age 2"
  ) +
  theme_classic()
```
Survival differences across years at CP2
```{r}
CP2_survival <- CP2_survival %>%
  group_by(plant_id) %>%
  mutate(original_census_year = min(Census_year)) %>% #create a column so we know when the plant was first censused
  ungroup()

#The fates of 1st year plants observed in 2024 haven't been uploaded yet so we exclude them for now
max_year_CP2 <- max(WL2_survival$Census_year)

#remove plants that first appeared in 2024 and were not observed beyond age 1
survival_by_year_CP2 <- CP2_survival %>%
  group_by(plant_id, original_census_year) %>%
  summarise(max_age = max(age),
            survived_to_age2 = as.integer(any(age == 2 & alive == 1)),
            .groups = "drop") %>%
  filter(!(original_census_year == max_year_CP2 & max_age == 1))
  
#summarize by year
CP2_survival_by_year <- survival_by_year_CP2 %>%
  group_by(original_census_year) %>%
  summarize(survival = mean(survived_to_age2),
            survival_se = sem(survived_to_age2),
            n=n(),
            .groups = "drop")

#plot survival differences across years at WL1
ggplot(CP2_survival_by_year, aes(x=original_census_year, survival)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin=survival - survival_se,
                    ymax = survival + survival_se),
                width = 0.15) +
  scale_y_continuous(limits = c(0, 1)) +
  labs(
    x = "Original census year",
    y = "Proportion surviving to age 2"
  ) +
  theme_classic()
```

Survival differences across years at TM2
```{r}
TM2_survival <- TM2_survival %>%
  group_by(plant_id) %>%
  mutate(original_census_year = min(Census_year)) %>% #create a column so we know when the plant was first censused
  ungroup()

#The fates of 1st year plants observed in 2024 haven't been uploaded yet so we exclude them for now
max_year_TM2 <- max(TM2_survival$Census_year)

#remove plants that first appeared in 2024 and were not observed beyond age 1
survival_by_year_TM2 <- TM2_survival %>%
  group_by(plant_id, original_census_year) %>%
  summarise(max_age = max(age),
            survived_to_age2 = as.integer(any(age == 2 & alive == 1)),
            .groups = "drop") %>%
  filter(!(original_census_year == max_year_TM2 & max_age == 1))
  
#summarize by year
TM2_survival_by_year <- survival_by_year_TM2 %>%
  group_by(original_census_year) %>%
  summarize(survival = mean(survived_to_age2),
            survival_se = sem(survived_to_age2),
            n=n(),
            .groups = "drop")

#plot survival differences across years at WL1
ggplot(TM2_survival_by_year, aes(x=original_census_year, survival)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin=survival - survival_se,
                    ymax = survival + survival_se),
                width = 0.15) +
  scale_y_continuous(limits = c(0, 1)) +
  labs(
    x = "Original census year",
    y = "Proportion surviving to age 2"
  ) +
  theme_classic()
```

